---
title: "10x and SmartSeq Integration"
author: "KQ"
date: "2025-04-16"
output: html_document
editor_options: 
  chunk_output_type: console
---

#We have 10x for CRC176B, CRC177B, and CRC177T as wellas SmartSeq Libraries for these.

## Initial Set Up for SmartSeq
```{r}
setwd("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq")
#/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeqExp2_QC
```
```{r}
set.seed(34)
```
```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
library(viridis)
library(tibble)
```
# Process through default steps for CRC177 Tumor1
```{r}
counts<-read.delim(file="250306_smartseq3.bbrd.q10.filt.srt.counts.tsv")
CRC177Tumor1 <- CreateSeuratObject(counts = counts, project = "SmartDeep_CRC177Tumor1", min.features = 100)
CRC177Tumor1 <- NormalizeData(object = CRC177Tumor1)
CRC177Tumor1 <- FindVariableFeatures(object = CRC177Tumor1)
CRC177Tumor1 <- ScaleData(object = CRC177Tumor1)

dim(CRC177Tumor1) #27096    92
#It probably means that for 3 cells we have less than 100 features. 

CRC177Tumor1 <- RunPCA(object = CRC177Tumor1, npcs = 91) #One less than Column Value

CRC177Tumor1 <- FindNeighbors(object = CRC177Tumor1, dims = 1:91)
CRC177Tumor1 <- FindClusters(object = CRC177Tumor1)

#Run UMAP
CRC177Tumor1 <- RunUMAP(object = CRC177Tumor1, dims = 1:10)

#Plot UMAP
DimPlot(CRC177Tumor1, reduction = "umap") + plot_annotation(title = "CRC177Tumor1")

#Plot metadata
VlnPlot(CRC177Tumor1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177Tumor1")

# Add CellType metadata to the Seurat object
CRC177Tumor1$CellType <- "CRC177Tumor1"

# Verify that the metadata has been added
head(CRC177Tumor1@meta.data)
```
```{r}
# Add cell names as a column
CRC177Tumor1@meta.data <- rownames_to_column(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Read the well cell type data
well_celltype_data_CRC177Tumor1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_Tumor_Plate3.csv")

# Merge the data frames
CRC177Tumor1@meta.data <- merge(CRC177Tumor1@meta.data, well_celltype_data_CRC177Tumor1, by.x = "cell_names_CRC177Tumor1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177Tumor1@meta.data <- column_to_rownames(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Check the result
head(CRC177Tumor1@meta.data)
```
```{r}
library(tibble)

# Add cell names as a column
CRC177Tumor1@meta.data <- rownames_to_column(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Read the well cell type data
well_celltype_data_CRC177Tumor1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_Tumor_Plate3.csv")

# Merge the data frames
CRC177Tumor1@meta.data <- merge(CRC177Tumor1@meta.data, well_celltype_data_CRC177Tumor1, by.x = "cell_names_CRC177Tumor1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177Tumor1@meta.data <- column_to_rownames(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Check the result
head(CRC177Tumor1@meta.data)

```

# Process through default steps for CRC177 PMBC1
```{r}
counts2<-read.delim(file="250306_smartseq1.bbrd.q10.filt.srt.counts.tsv")
CRC177PBMC1 <- CreateSeuratObject(counts = counts2, project = "SmartDeep_CRC177PBMC1", min.features = 100)
CRC177PBMC1 <- NormalizeData(object = CRC177PBMC1)
CRC177PBMC1 <- FindVariableFeatures(object = CRC177PBMC1)
CRC177PBMC1 <- ScaleData(object = CRC177PBMC1)

dim(CRC177PBMC1) #23805    93

CRC177PBMC1 <- RunPCA(object = CRC177PBMC1, npcs = 92) #One less than Column Value

CRC177PBMC1 <- FindNeighbors(object = CRC177PBMC1, dims = 1:92)
CRC177PBMC1 <- FindClusters(object = CRC177PBMC1)

#Run UMAP
CRC177PBMC1 <- RunUMAP(object = CRC177PBMC1, dims = 1:10)

#Plot UMAP
DimPlot(CRC177PBMC1, reduction = "umap") + plot_annotation(title = "CRC177PBMC1")

#Plot metadata
VlnPlot(CRC177PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177PBMC1")

# Add CellType metadata to the Seurat object
CRC177PBMC1$CellType <- "CRC177PBMC1"

# Verify that the metadata has been added
head(CRC177PBMC1@meta.data)
```
```{r}
# Check the Start
head(CRC177PBMC1@meta.data)

# Add cell names as a column
CRC177PBMC1@meta.data <- rownames_to_column(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Read the well cell type data
well_celltype_data_CRC177PBMC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate1.csv")

# Merge the data frames
CRC177PBMC1@meta.data <- merge(CRC177PBMC1@meta.data, well_celltype_data_CRC177PBMC1, by.x = "cell_names_CRC177PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC1@meta.data <- column_to_rownames(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Check the result
head(CRC177PBMC1@meta.data)
```
```{r}
# Check the Start
head(CRC177PBMC1@meta.data)

# Add cell names as a column
CRC177PBMC1@meta.data <- rownames_to_column(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Read the well cell type data
well_celltype_data_CRC177PBMC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate1.csv")

# Merge the data frames
CRC177PBMC1@meta.data <- merge(CRC177PBMC1@meta.data, well_celltype_data_CRC177PBMC1, by.x = "cell_names_CRC177PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC1@meta.data <- column_to_rownames(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Check the result
head(CRC177PBMC1@meta.data)
```

# Process through default steps for CRC177 PMBC2
```{r}
counts3<-read.delim(file="250306_smartseq2.bbrd.q10.filt.srt.counts.tsv")
CRC177PBMC2 <- CreateSeuratObject(counts = counts3, project = "SmartDeep_CRC177PBMC2", min.features = 100)
CRC177PBMC2 <- NormalizeData(object = CRC177PBMC2)
CRC177PBMC2 <- FindVariableFeatures(object = CRC177PBMC2)
CRC177PBMC2 <- ScaleData(object = CRC177PBMC2)

dim(CRC177PBMC2) #25962    88

CRC177PBMC2 <- RunPCA(object = CRC177PBMC2, npcs = 87) #One less than Column Value

CRC177PBMC2 <- FindNeighbors(object = CRC177PBMC2, dims = 1:87)
CRC177PBMC2 <- FindClusters(object = CRC177PBMC2)

#Run UMAP
CRC177PBMC2 <- RunUMAP(object = CRC177PBMC2, dims = 1:10)

#Plot UMAP
DimPlot(CRC177PBMC2, reduction = "umap") + plot_annotation(title = "CRC177PBMC2")

#Plot metadata
VlnPlot(CRC177PBMC2, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177PBMC2")

# Add CellType metadata to the Seurat object
CRC177PBMC2$CellType <- "CRC177PBMC2"

# Verify that the metadata has been added
head(CRC177PBMC2@meta.data)
```

```{r}
# Check the Start
head(CRC177PBMC2@meta.data)

# Add cell names as a column
CRC177PBMC2@meta.data <- rownames_to_column(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Read the well cell type data
well_celltype_data_CRC177PBMC2 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC177PBMC2@meta.data <- merge(CRC177PBMC2@meta.data, well_celltype_data_CRC177PBMC2, by.x = "cell_names_CRC177PBMC2", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC2@meta.data <- column_to_rownames(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Check the result
head(CRC177PBMC2@meta.data)
```
```{r}
# Check the Start
head(CRC177PBMC2@meta.data)

# Add cell names as a column
CRC177PBMC2@meta.data <- rownames_to_column(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Read the well cell type data
well_celltype_data_CRC177PBMC2 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC177PBMC2@meta.data <- merge(CRC177PBMC2@meta.data, well_celltype_data_CRC177PBMC2, by.x = "cell_names_CRC177PBMC2", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC2@meta.data <- column_to_rownames(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Check the result
head(CRC177PBMC2@meta.data)
```

# Process through default steps for CRC176 PBMC Plate 1
```{r}
counts5<-read.delim(file="250226_smartseq2.bbrd.q10.filt.srt.counts.tsv")
CRC176PBMC1 <- CreateSeuratObject(counts = counts5, project = "SmartDeep_CRC176PBMC1", min.features = 100)
CRC176PBMC1 <- NormalizeData(object = CRC176PBMC1)
CRC176PBMC1 <- FindVariableFeatures(object = CRC176PBMC1)
CRC176PBMC1 <- ScaleData(object = CRC176PBMC1)

dim(CRC176PBMC1) #21694    91

CRC176PBMC1 <- RunPCA(object = CRC176PBMC1, npcs = 90) #One less than Column Value

CRC176PBMC1 <- FindNeighbors(object = CRC176PBMC1, dims = 1:90)
CRC176PBMC1 <- FindClusters(object = CRC176PBMC1)

#Run UMAP
CRC176PBMC1 <- RunUMAP(object = CRC176PBMC1, dims = 1:10)
 
#Plot UMAP
DimPlot(CRC176PBMC1, reduction = "umap") + plot_annotation(title = "CRC176PBMC1")

#Plot metadata
VlnPlot(CRC176PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC176PBMC1")

# Add CellType metadata to the Seurat object
CRC176PBMC1$CellType <- "CRC176PBMC1"

# Verify that the metadata has been added
head(CRC176PBMC1@meta.data)
```

```{r}
# Check the Start
head(CRC176PBMC1@meta.data)

# Add cell names as a column
CRC176PBMC1@meta.data <- rownames_to_column(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Read the well cell type data
well_celltype_data_CRC176PMBC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC176PBMC1@meta.data <- merge(CRC176PBMC1@meta.data, well_celltype_data_CRC176PMBC1, by.x = "cell_names_CRC176PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC176PBMC1@meta.data <- column_to_rownames(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Check the result
head(CRC176PBMC1@meta.data)
```
```{r}
# Check the Start
head(CRC176PBMC1@meta.data)

# Add cell names as a column
CRC176PBMC1@meta.data <- rownames_to_column(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Read the well cell type data
well_celltype_data_CRC176PMBC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC176PBMC1@meta.data <- merge(CRC176PBMC1@meta.data, well_celltype_data_CRC176PMBC1, by.x = "cell_names_CRC176PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC176PBMC1@meta.data <- column_to_rownames(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Check the result
head(CRC176PBMC1@meta.data)
```

## Initial Set Up for 10x
```{r}
setwd("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq")
#/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeqExp2_QC
```
```{r}
#Set Seed
set.seed(34)
```
```{r}
#Load Required Libraries
library(Seurat)
library(ggplot2)
library(patchwork)
library(dplyr)
```
# Process through default steps for 10x_CRC177T
```{r}
# Load the dataset
hybrid.data_10xCRC177T <- Read10X(data.dir = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC177T_filtered_feature_bc_matrix/")
# Initialize the Seurat object with the raw (non-normalized data).
hybrid.data_10xCRC177T <- CreateSeuratObject(counts = hybrid.data_10xCRC177T, project = "hybrid10x_CRC177T", min.cells = 3, min.features = 200)
hybrid.data_10xCRC177T
```
```{r}
#QC and selecting cells for analysis
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
hybrid.data_10xCRC177T[["percent.mt"]] <- PercentageFeatureSet(hybrid.data_10xCRC177T, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(hybrid.data_10xCRC177T, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(hybrid.data_10xCRC177T, "percent.mt")

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot1 <- FeatureScatter(hybrid.data_10xCRC177T, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(hybrid.data_10xCRC177T, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 
plot2


hybrid.data_10xCRC177T <- subset(hybrid.data_10xCRC177T, subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & percent.mt < 5)
```
```{r}
#Normalizing Data. employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
hybrid.data_10xCRC177T <- NormalizeData(hybrid.data_10xCRC177T, normalization.method = "LogNormalize", scale.factor = 10000)
```
```{r}
#finding highly variable features 
hybrid.data_10xCRC177T <- FindVariableFeatures(hybrid.data_10xCRC177T, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(hybrid.data_10xCRC177T), 10)
top10

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(hybrid.data_10xCRC177T)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 
plot2
```
```{r}
#scaling the data 
all.genes <- rownames(hybrid.data_10xCRC177T)
hybrid.data_10xCRC177T <- ScaleData(hybrid.data_10xCRC177T, features = all.genes)
```
```{r}
#perform linear dimensionality reduction
hybrid.data_10xCRC177T <- RunPCA(hybrid.data_10xCRC177T, features = VariableFeatures(object = hybrid.data_10xCRC177T))

# Examine and visualize PCA results a few different ways
print(hybrid.data_10xCRC177T[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(hybrid.data_10xCRC177T, dims = 1:5, reduction = "pca")
VizDimLoadings(hybrid.data_10xCRC177T, dims = 1:2, reduction = "pca")
VizDimLoadings(hybrid.data_10xCRC177T, dims = 3:4, reduction = "pca")
VizDimLoadings(hybrid.data_10xCRC177T, dims = 4:5, reduction = "pca")

DimPlot(hybrid.data_10xCRC177T, reduction = "pca")
DimHeatmap(hybrid.data_10xCRC177T, dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(hybrid.data_10xCRC177T, dims = 1:5, cells = 500, balanced = TRUE)
```
```{r}
#determine dimensionality of the dataset 
hybrid.data_10xCRC177T <- JackStraw(hybrid.data_10xCRC177T, num.replicate = 100)
hybrid.data_10xCRC177T <- ScoreJackStraw(hybrid.data_10xCRC177T, dims = 1:20)
JackStrawPlot(hybrid.data_10xCRC177T, dims = 1:15)
ElbowPlot(hybrid.data_10xCRC177T)
```
```{r}
#cluster the cells
hybrid.data_10xCRC177T <- FindNeighbors(hybrid.data_10xCRC177T, dims = 1:10)
hybrid.data_10xCRC177T <- FindClusters(hybrid.data_10xCRC177T, resolution = 0.75)
head(Idents(hybrid.data_10xCRC177T), 5)
```
```{r}
#run linear dimensionality reduction UMAP
hybrid.data_10xCRC177T <- RunUMAP(hybrid.data_10xCRC177T, dims = 1:10)
DimPlot(hybrid.data_10xCRC177T, reduction = "umap")

DimPlot(hybrid.data_10xCRC177T, group.by = 'orig.ident', pt.size = 0.1)

saveRDS(hybrid.data_10xCRC177T, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/hybridfusionseq10xCRC177T.rds")
```
```{r}
#finding differentially expressed features 
# install limma for a more efficient implementation of the Wilcoxon Rank Sum Test
#install.packages('BiocManager')
#BiocManager::install('limma')

#LOOK AT PAx5

# find all markers for each individual cluster 0-7
cluster0.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 0, min.pct = 0.25)
head(cluster0.markers, n = 20)
write.csv(cluster0.markers, file = "cluster0_markers")
View(cluster0.markers)

cluster1.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 1, min.pct = 0.25)
write.csv(cluster1.markers, file = "cluster1_markers")
head(cluster1.markers, n = 20)

cluster2.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 2, min.pct = 0.25)
write.csv(cluster2.markers, file = "cluster2_markers")
head(cluster2.markers, n = 20)

cluster3.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 3, min.pct = 0.25)
write.csv(cluster3.markers, file = "cluster3_markers")
head(cluster3.markers, n = 20)

cluster4.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 4, min.pct = 0.25)
write.csv(cluster4.markers, file = "cluster4_markers")
head(cluster4.markers, n = 20)

cluster5.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 5, min.pct = 0.25)
write.csv(cluster5.markers, file = "cluster5_markers")
head(cluster5.markers, n = 20)

cluster6.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 6, min.pct = 0.25)
write.csv(cluster6.markers, file = "cluster6_markers")
head(cluster6.markers, n = 20)

cluster7.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 7, min.pct = 0.25)
write.csv(cluster7.markers, file = "cluster7_markers")
head(cluster7.markers, n = 20)

cluster8.markers <- FindMarkers(hybrid.data_10xCRC177T, ident.1 = 8, min.pct = 0.25)
write.csv(cluster8.markers, file = "cluster8_markers")
head(cluster8.markers, n = 20)

# find markers for every cluster compared to all remaining cells, report only the positive ones
hybrid.data_10xCRC177T.markers <- FindAllMarkers(hybrid.data_10xCRC177T, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
hybrid.data_10xCRC177T.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(hybrid.data_10xCRC177T.markers, file = "hybrid_markers")

hybrid.data_10xCRC177T.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(hybrid.data_10xCRC177T, features = top10$gene) + NoLegend()


saveRDS(hybrid.data_10xCRC177T, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/hybridfusionseq10xCRC177T.rds")
```
```{r}
#specific plots for genes of interest
FeaturePlot(hybrid.data_10xCRC177T, features = c("RUNX1", "SPI1", "GRN", "PF4", "THBS1", "LGALS3"))
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Lgals3", "Runx1", "Spi1", "Grn", "Pf4", "Thbs1"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Thbs1"))
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Ly6c2", "Vcan", "Ccr2", "Nlrp3", "Fcna"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Ki67", "Pcna"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Muc1", "Muc5ac", "Epcam", "Egfr", "Krt8", "Krt5", "Krt14", "Krt19", "Krt9", "Krt7", "Krt18", "Krt10", pt.size = 0.01))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Vim", "Acta2"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Ptprc", "Cd68", "Cd14","Fcgr3"))
# #note Fcgr3 = CD16 and Ptprc = CD45
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Cd44", "Cdx2", "Cdx1", "Tead4", "Alcam"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Mapk14", "Cdkn2a", "Mapk1", "Cdkn1a", "Yap1", "Trp53"))
# #note Mapk14 = p38, Cdkn2a = p16, mapk1 = Erk, cdkn1a = p21, 
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Pdcd1", "Cd274"))
# #note Pdcd1 = PD1 and CD274 = PDL1
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Cd47"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Hur"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Ctsk", "Ctsb", "Ctsl", "Ctsd", "Ctss", "Ctsc", "Ctsg", "Ctse", "Ctsz", "Ctsa", "Ctsh", "Ctsq", "Ctsj"))
# 
# FeaturePlot(hybrid.data_10xCRC177T, features = c("Gap1"))

```
# Process through default steps for 10x_CRC177B
```{r}
# Load the dataset
hybrid.data_10xCRC177B <- Read10X(data.dir = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC177B_filtered_feature_bc_matrix/")

# Initialize the Seurat object with the raw (non-normalized data).
hybrid_10xCRC177B <- CreateSeuratObject(counts = hybrid.data_10xCRC177B, project = "hybrid10x_CRC177B", min.cells = 3, min.features = 200)

hybrid_10xCRC177B

```
```{r}
#QC and selecting cells for analysis
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
hybrid_10xCRC177B[["percent.mt"]] <- PercentageFeatureSet(hybrid_10xCRC177B, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(hybrid_10xCRC177B, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(hybrid_10xCRC177B, "percent.mt")

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot1 <- FeatureScatter(hybrid_10xCRC177B, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(hybrid_10xCRC177B, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 
plot2


hybrid_10xCRC177B <- subset(hybrid_10xCRC177B, subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & percent.mt < 5)
```

```{r}
#Normalizing Data. employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
hybrid_10xCRC177B <- NormalizeData(hybrid_10xCRC177B, normalization.method = "LogNormalize", scale.factor = 10000)
```
 
```{r}
#finding highly variable features
hybrid_10xCRC177B <- FindVariableFeatures(hybrid_10xCRC177B, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(hybrid_10xCRC177B), 10)
top10

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(hybrid_10xCRC177B)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 
plot2
```


```{r}
#scaling the data 
all.genes <- rownames(hybrid_10xCRC177B)
hybrid_10xCRC177B <- ScaleData(hybrid_10xCRC177B, features = all.genes)
```


```{r}
#perform linear dimensionality reduction
hybrid_10xCRC177B <- RunPCA(hybrid_10xCRC177B, features = VariableFeatures(object = hybrid_10xCRC177B))

# Examine and visualize PCA results a few different ways
print(hybrid_10xCRC177B[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(hybrid_10xCRC177B, dims = 1:5, reduction = "pca")
VizDimLoadings(hybrid_10xCRC177B, dims = 1:2, reduction = "pca")
VizDimLoadings(hybrid_10xCRC177B, dims = 3:4, reduction = "pca")
VizDimLoadings(hybrid_10xCRC177B, dims = 4:5, reduction = "pca")

DimPlot(hybrid_10xCRC177B, reduction = "pca")
DimHeatmap(hybrid_10xCRC177B, dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(hybrid_10xCRC177B, dims = 1:5, cells = 500, balanced = TRUE)
```


```{r}
#determine dimensionality of the dataset 
hybrid_10xCRC177B <- JackStraw(hybrid_10xCRC177B, num.replicate = 100)
hybrid_10xCRC177B <- ScoreJackStraw(hybrid_10xCRC177B, dims = 1:20)
JackStrawPlot(hybrid_10xCRC177B, dims = 1:15)
ElbowPlot(hybrid_10xCRC177B)
```

```{r}
#cluster the cells
hybrid_10xCRC177B <- FindNeighbors(hybrid_10xCRC177B, dims = 1:10)
hybrid_10xCRC177B <- FindClusters(hybrid_10xCRC177B, resolution = 0.75)
head(Idents(hybrid_10xCRC177B), 5)
```


```{r}
#run linear dimensionality reduction UMAP
hybrid_10xCRC177B <- RunUMAP(hybrid_10xCRC177B, dims = 1:10)
DimPlot(hybrid_10xCRC177B, reduction = "umap")

DimPlot(hybrid_10xCRC177B, group.by = 'orig.ident', pt.size = 0.1)

saveRDS(hybrid_10xCRC177B, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/hybridfusionseq10xCRC177Bn.rds")
```


```{r}
#finding differentially expressed features 
# install limma for a more efficient implementation of the Wilcoxon Rank Sum Test
#install.packages('BiocManager')
#BiocManager::install('limma')
# find all markers for each individual cluster 0-7
cluster0.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 0, min.pct = 0.25)
head(cluster0.markers, n = 20)
write.csv(cluster0.markers, file = "cluster0_markers")
View(cluster0.markers)

cluster1.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 1, min.pct = 0.25)
write.csv(cluster1.markers, file = "cluster1_markers")
head(cluster1.markers, n = 20)

cluster2.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 2, min.pct = 0.25)
write.csv(cluster2.markers, file = "cluster2_markers")
head(cluster2.markers, n = 20)

cluster3.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 3, min.pct = 0.25)
write.csv(cluster3.markers, file = "cluster3_markers")
head(cluster3.markers, n = 20)

cluster4.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 4, min.pct = 0.25)
write.csv(cluster4.markers, file = "cluster4_markers")
head(cluster4.markers, n = 20)

cluster5.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 5, min.pct = 0.25)
write.csv(cluster5.markers, file = "cluster5_markers")
head(cluster5.markers, n = 20)

cluster6.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 6, min.pct = 0.25)
write.csv(cluster6.markers, file = "cluster6_markers")
head(cluster6.markers, n = 20)

cluster7.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 7, min.pct = 0.25)
write.csv(cluster7.markers, file = "cluster7_markers")
head(cluster7.markers, n = 20)

cluster8.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 8, min.pct = 0.25)
write.csv(cluster8.markers, file = "cluster8_markers")
head(cluster7.markers, n = 20)

cluster9.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 9, min.pct = 0.25)
write.csv(cluster9.markers, file = "cluster9_markers")
head(cluster9.markers, n = 20)

cluster10.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 10, min.pct = 0.25)
write.csv(cluster10.markers, file = "cluster10_markers")
head(cluster10.markers, n = 20)

cluster11.markers <- FindMarkers(hybrid_10xCRC177B, ident.1 = 11, min.pct = 0.25)
write.csv(cluster11.markers, file = "cluster11_markers")
head(cluster11.markers, n = 20)


# find markers for every cluster compared to all remaining cells, report only the positive ones
hybrid_10xCRC177B.markers <- FindAllMarkers(hybrid_10xCRC177B, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
hybrid_10xCRC177B.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(hybrid_10xCRC177B.markers, file = "hybrid_markers")

hybrid_10xCRC177B.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(hybrid_10xCRC177B, features = top10$gene) + NoLegend()


saveRDS(hybrid_10xCRC177B, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/hybridfusionseq10xCRC177Bn.rds")
```
```{r}
#specific plots for genes of interest
FeaturePlot(hybrid_10xCRC177B, features = c("RUNX1", "SPI1", "GRN", "PF4", "THBS1", "LGALS3"))
# FeaturePlot(hybrid_10xCRC177B, features = c("Lgals3", "Runx1", "Spi1", "Grn", "Pf4", "Thbs1"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Thbs1"))
# FeaturePlot(hybrid_10xCRC177B, features = c("Ly6c2", "Vcan", "Ccr2", "Nlrp3", "Fcna"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Ki67", "Pcna"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Muc1", "Muc5ac", "Epcam", "Egfr", "Krt8", "Krt5", "Krt14", "Krt19", "Krt9", "Krt7", "Krt18", "Krt10", pt.size = 0.01))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Vim", "Acta2"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Ptprc", "Cd68", "Cd14","Fcgr3"))
# #note Fcgr3 = CD16 and Ptprc = CD45
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Cd44", "Cdx2", "Cdx1", "Tead4", "Alcam"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Mapk14", "Cdkn2a", "Mapk1", "Cdkn1a", "Yap1", "Trp53"))
# #note Mapk14 = p38, Cdkn2a = p16, mapk1 = Erk, cdkn1a = p21, 
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Pdcd1", "Cd274"))
# #note Pdcd1 = PD1 and CD274 = PDL1
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Cd47"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Hur"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Ctsk", "Ctsb", "Ctsl", "Ctsd", "Ctss", "Ctsc", "Ctsg", "Ctse", "Ctsz", "Ctsa", "Ctsh", "Ctsq", "Ctsj"))
# 
# FeaturePlot(hybrid_10xCRC177B, features = c("Gap1"))
```
# Process through default steps for 10x_CRC176B
```{r}
# Load the dataset
hybrid.data_10xCRC176B <- Read10X(data.dir = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/")

# Initialize the Seurat object with the raw (non-normalized data).
hybrid_10xCRC176B <- CreateSeuratObject(counts = hybrid.data_10xCRC176B, project = "hybrid10x_CRC176B", min.cells = 3, min.features = 200)

hybrid_10xCRC176B

```

```{r}
#QC and selecting cells for analysis
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
hybrid_10xCRC176B[["percent.mt"]] <- PercentageFeatureSet(hybrid_10xCRC176B, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(hybrid_10xCRC176B, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(hybrid_10xCRC176B, "percent.mt")

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot1 <- FeatureScatter(hybrid_10xCRC176B, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(hybrid_10xCRC176B, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 
plot2


hybrid_10xCRC176B <- subset(hybrid_10xCRC176B, subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & percent.mt < 5)
```

```{r}
#Normalizing Data. employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
hybrid_10xCRC176B <- NormalizeData(hybrid_10xCRC176B, normalization.method = "LogNormalize", scale.factor = 10000)
```

```{r}
#finding highly variable features 
hybrid_10xCRC176B <- FindVariableFeatures(hybrid_10xCRC176B, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(hybrid_10xCRC176B), 10)
top10

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(hybrid_10xCRC176B)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 
plot2
```

```{r}
#scaling the data 
all.genes <- rownames(hybrid_10xCRC176B)
hybrid_10xCRC176B <- ScaleData(hybrid_10xCRC176B, features = all.genes)
```

```{r}
#perform linear dimensionality reduction
hybrid_10xCRC176B <- RunPCA(hybrid_10xCRC176B, features = VariableFeatures(object = hybrid_10xCRC176B))

# Examine and visualize PCA results a few different ways
print(hybrid_10xCRC176B[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(hybrid_10xCRC176B, dims = 1:5, reduction = "pca")
VizDimLoadings(hybrid_10xCRC176B, dims = 1:2, reduction = "pca")
VizDimLoadings(hybrid_10xCRC176B, dims = 3:4, reduction = "pca")
VizDimLoadings(hybrid_10xCRC176B, dims = 4:5, reduction = "pca")

DimPlot(hybrid_10xCRC176B, reduction = "pca")
DimHeatmap(hybrid_10xCRC176B, dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(hybrid_10xCRC176B, dims = 1:5, cells = 500, balanced = TRUE)
```

```{r}
#determine dimensionality of the dataset 
hybrid_10xCRC176B <- JackStraw(hybrid_10xCRC176B, num.replicate = 100)
hybrid_10xCRC176B <- ScoreJackStraw(hybrid_10xCRC176B, dims = 1:20)
JackStrawPlot(hybrid_10xCRC176B, dims = 1:15)
ElbowPlot(hybrid_10xCRC176B)
```

```{r}
#cluster the cells
hybrid_10xCRC176B <- FindNeighbors(hybrid_10xCRC176B, dims = 1:10)
hybrid_10xCRC176B <- FindClusters(hybrid_10xCRC176B, resolution = 0.75)
head(Idents(hybrid_10xCRC176B), 5)
```

```{r}
#run linear dimensionality reduction UMAP
hybrid_10xCRC176B <- RunUMAP(hybrid_10xCRC176B, dims = 1:10)
DimPlot(hybrid_10xCRC176B, reduction = "umap")

DimPlot(hybrid_10xCRC176B, group.by = 'orig.ident', pt.size = 0.1)

saveRDS(hybrid_10xCRC176B, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/hybridfusionseq10xCRC176Bn.rds")
```

```{r}
#finding differentially expressed features 
# install limma for a more efficient implementation of the Wilcoxon Rank Sum Test
#install.packages('BiocManager')
#BiocManager::install('limma')

# find all markers for each individual cluster 0-7
cluster0.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 0, min.pct = 0.25)
head(cluster0.markers, n = 20)
write.csv(cluster0.markers, file = "cluster0_markers")
View(cluster0.markers)

cluster1.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 1, min.pct = 0.25)
write.csv(cluster1.markers, file = "cluster1_markers")
head(cluster1.markers, n = 20)

cluster2.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 2, min.pct = 0.25)
write.csv(cluster2.markers, file = "cluster2_markers")
head(cluster2.markers, n = 20)

cluster3.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 3, min.pct = 0.25)
write.csv(cluster3.markers, file = "cluster3_markers")
head(cluster3.markers, n = 20)

cluster4.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 4, min.pct = 0.25)
write.csv(cluster4.markers, file = "cluster4_markers")
head(cluster4.markers, n = 20)

cluster5.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 5, min.pct = 0.25)
write.csv(cluster5.markers, file = "cluster5_markers")
head(cluster5.markers, n = 20)

cluster6.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 6, min.pct = 0.25)
write.csv(cluster6.markers, file = "cluster6_markers")
head(cluster6.markers, n = 20)

cluster7.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 7, min.pct = 0.25)
write.csv(cluster7.markers, file = "cluster7_markers")
head(cluster7.markers, n = 20)

cluster8.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 8, min.pct = 0.25)
write.csv(cluster8.markers, file = "cluster8_markers")
head(cluster8.markers, n = 20)

cluster9.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 9, min.pct = 0.25)
write.csv(cluster9.markers, file = "cluster9_markers")
head(cluster9.markers, n = 20)

cluster10.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 10, min.pct = 0.25)
write.csv(cluster10.markers, file = "cluster10_markers")
head(cluster10.markers, n = 20)

cluster11.markers <- FindMarkers(hybrid_10xCRC176B, ident.1 = 11, min.pct = 0.25)
write.csv(cluster11.markers, file = "cluster11_markers")
head(cluster11.markers, n = 20)


# find markers for every cluster compared to all remaining cells, report only the positive ones
hybrid_10xCRC176B.markers <- FindAllMarkers(hybrid_10xCRC176B, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
hybrid_10xCRC176B.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(hybrid_10xCRC176B.markers, file = "hybrid_markers")

hybrid_10xCRC176B.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(hybrid_10xCRC176B, features = top10$gene) + NoLegend()


saveRDS(hybrid_10xCRC176B, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/10xCompliment/CRC176B_filtered_feature_bc_matrix/hybridfusionseq10xCRC176Bn.rds")
```

```{r}
#specific plots for genes of interest
FeaturePlot(hybrid_10xCRC176B, features = c("RUNX1", "SPI1", "GRN", "PF4", "THBS1", "LGALS3"))
# FeaturePlot(hybrid_10xCRC176B, features = c("Lgals3", "Runx1", "Spi1", "Grn", "Pf4", "Thbs1"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Thbs1"))
# FeaturePlot(hybrid_10xCRC176B, features = c("Ly6c2", "Vcan", "Ccr2", "Nlrp3", "Fcna"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Ki67", "Pcna"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Muc1", "Muc5ac", "Epcam", "Egfr", "Krt8", "Krt5", "Krt14", "Krt19", "Krt9", "Krt7", "Krt18", "Krt10", pt.size = 0.01))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Vim", "Acta2"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Ptprc", "Cd68", "Cd14","Fcgr3"))
# #note Fcgr3 = CD16 and Ptprc = CD45
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Cd44", "Cdx2", "Cdx1", "Tead4", "Alcam"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Mapk14", "Cdkn2a", "Mapk1", "Cdkn1a", "Yap1", "Trp53"))
# #note Mapk14 = p38, Cdkn2a = p16, mapk1 = Erk, cdkn1a = p21, 
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Pdcd1", "Cd274"))
# #note Pdcd1 = PD1 and CD274 = PDL1
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Cd47"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Hur"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Ctsk", "Ctsb", "Ctsl", "Ctsd", "Ctss", "Ctsc", "Ctsg", "Ctse", "Ctsz", "Ctsa", "Ctsh", "Ctsq", "Ctsj"))
# 
# FeaturePlot(hybrid_10xCRC176B, features = c("Gap1"))

```

# Select Integration Features
```{r}
#Assuming you have your Seurat objects in a list
Smart10x_List <- list(hybrid_10xCRC176B, hybrid_10xCRC177B, hybrid.data_10xCRC177T,CRC177Tumor1,CRC177PBMC1,CRC177PBMC2,CRC176PBMC1)
```

```{r}
# Normalize and find variable features for each dataset
Smart10x_List <- lapply(Smart10x_List, function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x, selection.method = "vst", nfeatures = 2000)
  return(x)
})
```

```{r}
#checking for overlap between data sets
# Compute the average expression for each dataset
avg_exp_list <- lapply(Smart10x_List, function(x) rowMeans(GetAssayData(x, slot = "counts")))

# Find shared genes across all experiments
shared_genes <- Reduce(intersect, lapply(avg_exp_list, names))

# Create a correlation matrix for shared genes across experiments
cor_matrix <- cor(do.call(cbind, lapply(avg_exp_list, function(x) x[shared_genes])))

# Print correlation matrix
print(cor_matrix)
#Okay - so we're not seeing a ton of correlations as is... mostly between modalities.
```


```{r}

pca1 <- Embeddings(pca_list[[1]], "pca")[,1:5]  # Extract first 5 PCs
pca2 <- Embeddings(pca_list[[2]], "pca")[,1:5]

# Assuming cell names are stored in rownames
common_cells <- intersect(rownames(pca1), rownames(pca2))

# Subset PCA matrices to include only common cells
pca1_subset <- pca1[common_cells, ]
pca2_subset <- pca2[common_cells, ]

# Compute correlation matrix
cor_pca <- cor(pca1_subset, pca2_subset)

# Print correlation matrix
print(cor_pca)

#UMAP
p_list <- lapply(Smart10x_List, function(x) DimPlot(x, reduction = "umap"))
cowplot::plot_grid(plotlist = p_list)

```

```{r}
# Find Integration Anchors
anchors <- FindIntegrationAnchors(object.list = Smart10x_List, anchor.features = 2000, dims = 1:30)

```

```{r}
# Integrate Data
integrated_data <- IntegrateData(anchorset = anchors, dims = 1:30, k.weight = 4)
#length(anchors@anchors) gave a result of 5, which informed my k.weight selection. The default was 92. 

```
```{r}
DefaultAssay(integrated_data) <- "integrated"
# Scale the integrated data
integrated_data <- ScaleData(integrated_data, verbose = FALSE)

# Run PCA to reduce dimensionality
integrated_data <- RunPCA(integrated_data, npcs = 10, verbose = FALSE)
# Check optimal dimensions
#ElbowPlot(integrated_data) #The outcome was 10. Maybe 9.
```
```{r}
# Optionally, find the neighborhoods and clusters
integrated_data <- FindNeighbors(integrated_data, reduction = "pca", dims = 1:10)
integrated_data <- FindClusters(integrated_data, resolution = 0.4)  # Adjust resolution as needed

# Run UMAP for a non-linear dimension reduction and visualization
integrated_data <- RunUMAP(integrated_data, reduction = "pca", dims = 1:10)

# Visualize the integrated data using a UMAP plot colored by cluster identity
DimPlot(integrated_data, reduction = "umap")
```

```{r}
library(cluster)
# Experiment with different resolutions
resolutions <- c(0.2, 0.3, 0.4, 0.5, 1.0)
for (res in resolutions) {
  integrated_data <- FindClusters(integrated_data, resolution = res)
  print(paste("Resolution:", res))
  print(table(Idents(integrated_data)))
  
  # Compute silhouette score
  sil_scores <- silhouette(as.numeric(Idents(integrated_data)), dist(Embeddings(integrated_data, reduction = "pca")[, 1:10]))
  avg_sil_score <- mean(sil_scores[, 3])
  print(paste("Average silhouette score:", avg_sil_score))
}
# Takeaway is that even at 0.3 and 0.4 is the peak?
```


```{r}
# Extract UMAP embeddings from your integrated Seurat object
umap_coords <- Embeddings(integrated_data, reduction = "umap")
umap_df <- as.data.frame(umap_coords)

# Add cluster identities (assuming clusters are stored in 'seurat_clusters')
umap_df$Cluster <- integrated_data$seurat_clusters  
umap_df$Cluster <- as.factor(umap_df$Cluster)  # ensure clusters are treated as discrete factors
umap_df$orig.ident <- as.factor(integrated_data$orig.ident)  # Original dataset source

# Load ggplot2 and viridis libraries
library(ggplot2)
library(viridis)

# Define a vector of shapes. Adjust the length as necessary; here we choose 10 different shapes.
# Common numeric shape codes: 15 (square), 16 (circle), 17 (triangle), 18 (diamond), etc.
shapes_vector <- c(16, 17, 18, 21, 24, 25, 23, 7, 10, 12)  

# Create the ggplot using the UMAP coordinates
integratedplot <- ggplot(umap_df, aes(x = umap_1, y = umap_2, 
                           color = Cluster, shape = orig.ident)) +
  geom_point(size = 1, alpha = 0.5) +  # Adjust point size and transparency (alpha)
  scale_color_viridis_d() +                                  # Use discrete Viridis palette for colors
  scale_shape_manual(values = shapes_vector[1:length(levels(umap_df$Cluster))]) +
  labs(title = "UMAP Plot: Viridis Colors with Custom Shapes",
       x = "UMAP 1",
       y = "UMAP 2") +
  theme_minimal()

# Display the plot
print(integratedplot)

```

```{r}
#UMAP by WellCellType


# Check metadata columns in each Seurat object
head(integrated_data@meta.data)


# Plot UMAP with clusters associated with shapes
wellcell_plot <- DimPlot(integrated_data, reduction = "umap", shape.by = "WellCellType.x", pt.size = 2, alpha = 0.5) +
  scale_color_viridis(discrete = TRUE) + plot_annotation(title = "0.75 by Well Cell Type")


# Apply the viridis color scale and change the title
wellcell_plot + scale_color_viridis_d() + ggtitle("Well Cell Types for Integrated Assays") + 
  labs(x = NULL, y = NULL) 

```
```{r}
# Extract UMAP coordinates and cluster information into a data frame
umap_coords <- Embeddings(integrated_data, reduction = "umap")
umap_df <- as.data.frame(umap_coords)
umap_df$WellCellType <- integrated_data$WellCellType.x

# Create a separate data frame for NA values
na_df <- umap_df[is.na(umap_df$WellCellType), ]

# Plot UMAP with clusters associated with shapes
wellcell_plot <- DimPlot(integrated_data, reduction = "umap", shape.by = "WellCellType.x", pt.size = 2, alpha = 1) +
  scale_color_viridis(discrete = TRUE) +
  plot_annotation(title = "0.75 by Well Cell Type") +
  scale_color_viridis_d() +
  ggtitle("Well Cell Types for Integrated Assays") +
  labs(x = NULL, y = NULL)

# Overlay points for cells with NA values
wellcell_plot <- wellcell_plot + 
  geom_point(data = na_df, aes(x = umap_1, y = umap_2), shape = 0, size = 2, alpha = 0.1)

# Print the plot
print(wellcell_plot)

```


```{r}
#install.packages("devtools")  # Install devtools if needed
#devtools::install_github("immunogenomics/harmony")  # Install Harmony
#install.packages(c("ggtree", "patchwork", "ggfun", "fgsea", "shadowtext",
                   # "scatterpie", "scattermore", "qvalue", "ggridges", "viridis",
                   # "aplot", "ggforce", "DOSE", "ggrepel", "enrichplot",
                   # "ggraph", "cowplot", "sctransform", "Seurat", "plotly", "ggplotify"))
#.rs.restartR()
```

```{r}
library(harmony)
integrated_data_harmonized <- RunHarmony(integrated_data, group.by.vars = "orig.ident", reduction.save = "pca_harmony", theta = 10)

# Extract UMAP embeddings from your integrated Seurat object
umap_coords_h <- Embeddings(integrated_data_harmonized, reduction = "umap")
umap_h_df <- as.data.frame(umap_coords_h)

# Add cluster identities (assuming clusters are stored in 'seurat_clusters')
umap_h_df$Cluster <- integrated_data_harmonized$seurat_clusters  
umap_h_df$Cluster <- as.factor(umap_h_df$Cluster)  # ensure clusters are treated as discrete factors
umap_h_df$orig.ident <- as.factor(integrated_data_harmonized$orig.ident)  # Original dataset source

# Load ggplot2 and viridis libraries
library(ggplot2)
library(viridis)

# Define a vector of shapes. Adjust the length as necessary; here we choose 10 different shapes.
# Common numeric shape codes: 15 (square), 16 (circle), 17 (triangle), 18 (diamond), etc.
shapes_vector <- c(16, 17, 18, 21, 24, 25, 23, 7, 10, 12)  

# Create the ggplot using the UMAP coordinates
integratedplot_h <- ggplot(umap_h_df, aes(x = umap_1, y = umap_2, 
                           color = Cluster, shape = orig.ident)) +
  geom_point(size = 1, alpha = 0.5) +  # Adjust point size and transparency (alpha)
  scale_color_viridis_d() +                                  # Use discrete Viridis palette for colors
  scale_shape_manual(values = shapes_vector[1:length(levels(umap_h_df$Cluster))]) +
  labs(title = "UMAP Plot: Viridis Colors with Custom Shapes after Harmony 10",
       x = "UMAP 1",
       y = "UMAP 2") +
  theme_minimal()

# Display the plot
print(integratedplot_h)

```

```{r}
library(harmony)
integrated_data_harmonized <- RunHarmony(integrated_data, group.by.vars = "orig.ident", reduction.save = "pca_harmony", theta = 10)

# Extract UMAP embeddings from your integrated Seurat object
umap_coords_h <- Embeddings(integrated_data_harmonized, reduction = "umap")
umap_h_df <- as.data.frame(umap_coords_h)

# Add cluster identities (assuming clusters are stored in 'seurat_clusters')
umap_h_df$Cluster <- integrated_data_harmonized$seurat_clusters  
umap_h_df$Cluster <- as.factor(umap_h_df$Cluster)  # ensure clusters are treated as discrete factors
umap_h_df$orig.ident <- as.factor(integrated_data_harmonized$orig.ident)  # Original dataset source

# Load ggplot2 and viridis libraries
library(ggplot2)
library(viridis)

# Define a vector of shapes. Adjust the length as necessary; here we choose 10 different shapes.
# Common numeric shape codes: 15 (square), 16 (circle), 17 (triangle), 18 (diamond), etc.
shapes_vector <- c(16, 17, 18, 21, 24, 25, 23, 7, 10, 12)  

# Create the ggplot using the UMAP coordinates
integratedplot_h <- ggplot(umap_h_df, aes(x = umap_1, y = umap_2, 
                           color = orig.ident)) +
  geom_point(size = 1, alpha = 0.5) +  # Adjust point size and transparency (alpha)
  scale_color_viridis_d() +                                  # Use discrete Viridis palette for colors
  scale_shape_manual(values = shapes_vector[1:length(levels(umap_h_df$Cluster))]) +
  labs(title = "UMAP Plot: After Harmony 10",
       x = "UMAP 1",
       y = "UMAP 2") +
  theme_minimal()

# Display the plot
print(integratedplot_h)

```

```{r}
library(harmony)
integrated_data_harmonized <- RunHarmony(integrated_data, group.by.vars = "orig.ident", reduction.save = "pca_harmony", theta = 10)

# Extract UMAP embeddings from your integrated Seurat object
umap_coords_h <- Embeddings(integrated_data_harmonized, reduction = "umap")
umap_h_df <- as.data.frame(umap_coords_h)

# Add cluster identities (assuming clusters are stored in 'seurat_clusters')
umap_h_df$Cluster <- integrated_data_harmonized$seurat_clusters  
umap_h_df$Cluster <- as.factor(umap_h_df$Cluster)  # ensure clusters are treated as discrete factors
umap_h_df$orig.ident <- as.factor(integrated_data_harmonized$orig.ident)  # Original dataset source


# Check if WellCellType exists and add it if available
if ("WellCellType" %in% colnames(integrated_data_harmonized)) {umap_h_df$WellCellType <- as.factor(integrated_data_harmonized$WellCellType) # WellCellType metadata
} else {umap_h_df$WellCellType <- NA # Assign NA if WellCellType is not available
}

# Load ggplot2 and viridis libraries
#library(ggplot2)
#library(viridis)

# Define a vector of shapes. Adjust the length as necessary; here we choose 10 different shapes.
# Common numeric shape codes: 15 (square), 16 (circle), 17 (triangle), 18 (diamond), etc.
shapes_vector <- c(16, 17, 18, 21, 24, 25, 23, 7, 10, 12) 

# Create the ggplot using the UMAP coordinates
integratedplot_h <- ggplot(umap_h_df, aes(x = umap_1, y = umap_2, 
                           color = orig.ident, shape = WellCellType)) +
  geom_point(size = 1, alpha = 0.5) +  # Adjust point size and transparency (alpha)
  scale_color_viridis_d() +                                  # Use discrete Viridis palette for colors
  scale_shape_manual(values = shapes_vector[1:length(levels(umap_h_df$Cluster))]) +
  labs(title = "UMAP Plot: After Harmony 10",
       x = "UMAP 1",
       y = "UMAP 2") +
  theme_minimal()

# Display the plot
print(integratedplot_h)
```

```{r}
DefaultAssay(integrated_data_harmonized) <- "RNA"
#specific plots for genes of interest
expr_matrix <- GetAssay(integrated_data_harmonized, assay = "RNA", layer = "counts")[gene_list, ]

```
```{r}
#EPCAM, CDH1, PTPRC, CD68, ITGAM, RUNX1, SPI1, THBS1
FeaturePlot(
  integrated_data_harmonized, 
  features = c("THBS1"),
  cols = c("lightgray", "blue"),  # Adjust color gradient
  min.cutoff = 0,  # Removes extreme low-expression noise
  max.cutoff = 10,  # Caps extreme high values
  pt.size = 1  # Increase point size for visibility
)
```