---
title: "Adding Contorl Cell Meta Data"
author: "KQ"
date: "2025-04-15"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial Set Up
```{r}
setwd("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq")
#/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeqExp2_QC
```
```{r}
set.seed(34)
```
```{r}
library(Seurat)
```
```{r}
library(ggplot2)
library(patchwork)
```
# Process through default steps for CRC177 Tumor1
```{r}
counts<-read.delim(file="250306_smartseq3.bbrd.q10.filt.srt.counts.tsv")
CRC177Tumor1 <- CreateSeuratObject(counts = counts, project = "SmartDeep_CRC177Tumor1", min.features = 100)
CRC177Tumor1 <- NormalizeData(object = CRC177Tumor1)
CRC177Tumor1 <- FindVariableFeatures(object = CRC177Tumor1)
CRC177Tumor1 <- ScaleData(object = CRC177Tumor1)

dim(CRC177Tumor1) #27096    92
#It probably means that for 3 cells we have less than 100 features. 

CRC177Tumor1 <- RunPCA(object = CRC177Tumor1, npcs = 91) #One less than Column Value

CRC177Tumor1 <- FindNeighbors(object = CRC177Tumor1, dims = 1:91)
CRC177Tumor1 <- FindClusters(object = CRC177Tumor1)

#Run UMAP
CRC177Tumor1 <- RunUMAP(object = CRC177Tumor1, dims = 1:10)

#Plot UMAP
DimPlot(CRC177Tumor1, reduction = "umap") + plot_annotation(title = "CRC177Tumor1")

#Plot metadata
VlnPlot(CRC177Tumor1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177Tumor1")

# Add CellType metadata to the Seurat object
CRC177Tumor1$CellType <- "CRC177Tumor1"

# Verify that the metadata has been added
head(CRC177Tumor1@meta.data)
```
##Adding Metadata
```{r}
library(tibble)

# Add cell names as a column
CRC177Tumor1@meta.data <- rownames_to_column(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Read the well cell type data
well_celltype_data_CRC177Tumor1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_Tumor_Plate3.csv")

# Merge the data frames
CRC177Tumor1@meta.data <- merge(CRC177Tumor1@meta.data, well_celltype_data_CRC177Tumor1, by.x = "cell_names_CRC177Tumor1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177Tumor1@meta.data <- column_to_rownames(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Check the result
head(CRC177Tumor1@meta.data)

```

# Process through default steps for CRC177 PMBC1
```{r}
counts2<-read.delim(file="250306_smartseq1.bbrd.q10.filt.srt.counts.tsv")
CRC177PBMC1 <- CreateSeuratObject(counts = counts2, project = "SmartDeep_CRC177PBMC1", min.features = 100)
CRC177PBMC1 <- NormalizeData(object = CRC177PBMC1)
CRC177PBMC1 <- FindVariableFeatures(object = CRC177PBMC1)
CRC177PBMC1 <- ScaleData(object = CRC177PBMC1)

dim(CRC177PBMC1) #23805    93

CRC177PBMC1 <- RunPCA(object = CRC177PBMC1, npcs = 92) #One less than Column Value

CRC177PBMC1 <- FindNeighbors(object = CRC177PBMC1, dims = 1:92)
CRC177PBMC1 <- FindClusters(object = CRC177PBMC1)

#Run UMAP
CRC177PBMC1 <- RunUMAP(object = CRC177PBMC1, dims = 1:10)

#Plot UMAP
DimPlot(CRC177PBMC1, reduction = "umap") + plot_annotation(title = "CRC177Tumor1")

#Plot metadata
VlnPlot(CRC177PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177Tumor1")

# Add CellType metadata to the Seurat object
CRC177PBMC1$CellType <- "CRC177PBMC1"

# Verify that the metadata has been added
head(CRC177PBMC1@meta.data)
```

```{r}
# Check the Start
head(CRC177PBMC1@meta.data)

# Add cell names as a column
CRC177PBMC1@meta.data <- rownames_to_column(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Read the well cell type data
well_celltype_data_CRC177PBMC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate1.csv")

# Merge the data frames
CRC177PBMC1@meta.data <- merge(CRC177PBMC1@meta.data, well_celltype_data_CRC177PBMC1, by.x = "cell_names_CRC177PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC1@meta.data <- column_to_rownames(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Check the result
head(CRC177PBMC1@meta.data)
```
# Process through default steps for CRC177 PMBC2
```{r}
counts3<-read.delim(file="250306_smartseq2.bbrd.q10.filt.srt.counts.tsv")
CRC177PBMC2 <- CreateSeuratObject(counts = counts3, project = "SmartDeep_CRC177PBMC2", min.features = 100)
CRC177PBMC2 <- NormalizeData(object = CRC177PBMC2)
CRC177PBMC2 <- FindVariableFeatures(object = CRC177PBMC2)
CRC177PBMC2 <- ScaleData(object = CRC177PBMC2)

dim(CRC177PBMC2) #25962    88

CRC177PBMC2 <- RunPCA(object = CRC177PBMC2, npcs = 87) #One less than Column Value

CRC177PBMC2 <- FindNeighbors(object = CRC177PBMC2, dims = 1:87)
CRC177PBMC2 <- FindClusters(object = CRC177PBMC2)

#Run UMAP
CRC177PBMC2 <- RunUMAP(object = CRC177PBMC2, dims = 1:10)

#Plot UMAP
DimPlot(CRC177PBMC2, reduction = "umap") + plot_annotation(title = "CRC177PBMC2")

#Plot metadata
VlnPlot(CRC177PBMC2, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177PBMC2")

# Add CellType metadata to the Seurat object
CRC177PBMC2$CellType <- "CRC177PBMC2"

# Verify that the metadata has been added
head(CRC177PBMC2@meta.data)
```

```{r}
# Check the Start
head(CRC177PBMC2@meta.data)

# Add cell names as a column
CRC177PBMC2@meta.data <- rownames_to_column(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Read the well cell type data
well_celltype_data_CRC177PBMC2 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC177PBMC2@meta.data <- merge(CRC177PBMC2@meta.data, well_celltype_data_CRC177PBMC2, by.x = "cell_names_CRC177PBMC2", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC2@meta.data <- column_to_rownames(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Check the result
head(CRC177PBMC2@meta.data)
```
# Process through default steps for CRC179 PBMC Plate 1
```{r}
counts4<-read.delim(file="250226_smartseq1.bbrd.q10.filt.srt.counts.tsv")
CRC179PBMC1 <- CreateSeuratObject(counts = counts4, project = "SmartDeep_CRC179PBMC1", min.features = 100)
CRC179PBMC1 <- NormalizeData(object = CRC179PBMC1)
CRC179PBMC1 <- FindVariableFeatures(object = CRC179PBMC1)
CRC179PBMC1 <- ScaleData(object = CRC179PBMC1)

dim(CRC179PBMC1) #21322    90

CRC179PBMC1 <- RunPCA(object = CRC179PBMC1, npcs = 89) #One less than Column Value

CRC179PBMC1 <- FindNeighbors(object = CRC179PBMC1, dims = 1:89)
CRC179PBMC1 <- FindClusters(object = CRC179PBMC1)

#Run UMAP
CRC179PBMC1 <- RunUMAP(object = CRC179PBMC1, dims = 1:10)

#Plot UMAP
DimPlot(CRC179PBMC1, reduction = "umap") + plot_annotation(title = "CRC179PBMC1")

#Plot metadata
VlnPlot(CRC179PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC179PBMC1")

# Add CellType metadata to the Seurat object
CRC179PBMC1$CellType <- "CRC179PBMC1"

# Verify that the metadata has been added
head(CRC179PBMC1@meta.data)
```

```{r}
# Check the Start
head(CRC179PBMC1@meta.data)

# Add cell names as a column
CRC179PBMC1@meta.data <- rownames_to_column(CRC179PBMC1@meta.data, var = "cell_names_CRC179PBMC1")

# Read the well cell type data
well_celltype_data_CRC179PMBC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate1.csv")

# Merge the data frames
CRC179PBMC1@meta.data <- merge(CRC179PBMC1@meta.data, well_celltype_data_CRC179PMBC1, by.x = "cell_names_CRC179PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC179PBMC1@meta.data <- column_to_rownames(CRC179PBMC1@meta.data, var = "cell_names_CRC179PBMC1")

# Check the result
head(CRC179PBMC1@meta.data)
```

# Process through default steps for CRC176 PBMC Plate 1
```{r}
counts5<-read.delim(file="250226_smartseq2.bbrd.q10.filt.srt.counts.tsv")
CRC176PBMC1 <- CreateSeuratObject(counts = counts5, project = "SmartDeep_CRC176PBMC1", min.features = 100)
CRC176PBMC1 <- NormalizeData(object = CRC176PBMC1)
CRC176PBMC1 <- FindVariableFeatures(object = CRC176PBMC1)
CRC176PBMC1 <- ScaleData(object = CRC176PBMC1)

dim(CRC176PBMC1) #21694    91

CRC176PBMC1 <- RunPCA(object = CRC176PBMC1, npcs = 90) #One less than Column Value

CRC176PBMC1 <- FindNeighbors(object = CRC176PBMC1, dims = 1:90)
CRC176PBMC1 <- FindClusters(object = CRC176PBMC1)

#Run UMAP
CRC176PBMC1 <- RunUMAP(object = CRC176PBMC1, dims = 1:10)
 
#Plot UMAP
DimPlot(CRC176PBMC1, reduction = "umap") + plot_annotation(title = "CRC176PBMC1")

#Plot metadata
VlnPlot(CRC176PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC176PBMC1")

# Add CellType metadata to the Seurat object
CRC176PBMC1$CellType <- "CRC176PBMC1"

# Verify that the metadata has been added
head(CRC176PBMC1@meta.data)
```

```{r}
# Check the Start
head(CRC176PBMC1@meta.data)

# Add cell names as a column
CRC176PBMC1@meta.data <- rownames_to_column(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Read the well cell type data
well_celltype_data_CRC176PMBC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC176PBMC1@meta.data <- merge(CRC176PBMC1@meta.data, well_celltype_data_CRC176PMBC1, by.x = "cell_names_CRC176PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC176PBMC1@meta.data <- column_to_rownames(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Check the result
head(CRC176PBMC1@meta.data)
```


# Process through default steps for CRC176 Tumor Plate 1
```{r}
counts6<-read.delim(file="250226_smartseq3.bbrd.q10.filt.srt.counts.tsv")
CRC176Tumor1 <- CreateSeuratObject(counts = counts6, project = "SmartDeep_CRC176Tumor1", min.features = 100)
CRC176Tumor1 <- NormalizeData(object = CRC176Tumor1)
CRC176Tumor1 <- FindVariableFeatures(object = CRC176Tumor1)
CRC176Tumor1 <- ScaleData(object = CRC176Tumor1)

dim(CRC176Tumor1) #34747    85

CRC176Tumor1 <- RunPCA(object = CRC176Tumor1, npcs = 84) #One less than Column Value

CRC176Tumor1 <- FindNeighbors(object = CRC176Tumor1, dims = 1:84)
CRC176Tumor1 <- FindClusters(object = CRC176Tumor1)

#Run UMAP
CRC176Tumor1 <- RunUMAP(object = CRC176Tumor1, dims = 1:10)

#Plot UMAP
DimPlot(CRC176Tumor1, reduction = "umap") + plot_annotation(title = "CRC176Tumor1")

#Plot metadata
VlnPlot(CRC176Tumor1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC176Tumor1")

# Add CellType metadata to the Seurat object
CRC176Tumor1$CellType <- "CRC176Tumor1"

# Verify that the metadata has been added
head(CRC176Tumor1@meta.data)
```
##Adding Metadata
```{r}
# Check the Start
head(CRC176Tumor1@meta.data)

# Add cell names as a column
CRC176Tumor1@meta.data <- rownames_to_column(CRC176Tumor1@meta.data, var = "cell_names_CRC176Tumor1")

# Read the well cell type data
well_celltype_data_CRC176Tumor1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_Tumor_Plate3.csv")

# Merge the data frames
CRC176Tumor1@meta.data <- merge(CRC176Tumor1@meta.data, well_celltype_data_CRC176Tumor1, by.x = "cell_names_CRC176Tumor1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC176Tumor1@meta.data <- column_to_rownames(CRC176Tumor1@meta.data, var = "cell_names_CRC176Tumor1")

# Check the result
head(CRC176Tumor1@meta.data)
```

##PreMerge Trouble Shooting
```{r}
# Check metadata columns in each Seurat object
head(CRC177Tumor1@meta.data)
head(CRC177PBMC1@meta.data)
head(CRC177PBMC2@meta.data)
head(CRC176PBMC1@meta.data)
head(CRC176Tumor1@meta.data)
head(CRC179PBMC1@meta.data)
```
```{r}
sum(is.na(CRC177PBMC2$CellType))
sum(is.na(CRC177PBMC2$WellCellType))
sum(is.na(CRC177PBMC1$CellType))
sum(is.na(CRC177PBMC1$WellCellType))
sum(is.na(CRC176PBMC1$CellType))
sum(is.na(CRC176PBMC1$WellCellType))
sum(is.na(CRC179PBMC1$CellType))
sum(is.na(CRC179PBMC1$WellCellType))
sum(is.na(CRC177Tumor1$CellType))
sum(is.na(CRC177Tumor1$WellCellType))
sum(is.na(CRC176Tumor1$CellType))
sum(is.na(CRC176Tumor1$WellCellType))
```
```{r}
any(duplicated(colnames(CRC177Tumor1)))
any(duplicated(colnames(CRC177PBMC1)))
any(duplicated(colnames(CRC176PBMC1)))
any(duplicated(colnames(CRC177PBMC2)))
any(duplicated(colnames(CRC179PBMC1)))
any(duplicated(colnames(CRC176Tumor1)))
```

```{r}
library(dplyr)

CRC177Tumor1@meta.data <- CRC177Tumor1@meta.data %>% select(-cell_names_CRC177Tumor1)
CRC177PBMC1@meta.data <- CRC177PBMC1@meta.data %>% select(-cell_names_CRC177PBMC1)
CRC177PBMC2@meta.data <- CRC177PBMC2@meta.data %>% select(-cell_names_CRC177PBMC2)
CRC176PBMC1@meta.data <- CRC176PBMC1@meta.data %>% select(-cell_names_CRC176PBMC1)
CRC176Tumor1@meta.data <- CRC176Tumor1@meta.data %>% select(-cell_names_CRC176Tumor1)
CRC179PBMC1@meta.data <- CRC179PBMC1@meta.data %>% select(-cell_names_CRC179PBMC1)

```

##Attempt to Merge
```{r}
# Merge the Seurat objects with unique cell names
merged_data_all <- merge(CRC177Tumor1, 
                     y = list(CRC177PBMC1, CRC177PBMC2, CRC176PBMC1, CRC176Tumor1, CRC179PBMC1), 
                     add.cell.ids = c("CRC177Tumor1", "CRC177PBMC1", "CRC177PBMC2", "CRC176PBMC1", "CRC176Tumor1", "CRC179PMBC1"))

# Verify the merged object
print(merged_data_all)
```

```{r}
merged_data_all <- NormalizeData(object = merged_data_all)
merged_data_all <- FindVariableFeatures(object = merged_data_all)
merged_data_all <- ScaleData(object = merged_data_all)
```

```{r}
dim(merged_data_all) #40287   539
# Check the dimensions of the scaled data matrix
```

```{r}
# Run PCA
# Run PCA with a reduced number of PCs
merged_data_all <- RunPCA(merged_data_all, npcs = 50)  # Adjust npcs as needed

# Verify the PCA results
print(merged_data_all)
```

```{r}
# Check the variance explained by each PC
ElbowPlot(merged_data_all)
```

```{r}
# Run PCA with the optimal number of PCs
merged_data_all <- RunPCA(merged_data_all, npcs = 10)

# Verify the PCA results
print(merged_data_all)
# Check the variance explained by each PC
ElbowPlot(merged_data_all)
```

```{r}
merged_data_all <- FindNeighbors(object = merged_data_all, dims = 1:10)
merged_data_all <- FindClusters(object = merged_data_all)
```

```{r}
merged_data_all <- RunUMAP(object = merged_data_all, dims = 1:10)
```
```{r}
library(viridis)
```
Plot UMAP

```{r}
merged_data_all_7 <- FindClusters(object = merged_data_all, resolution = 0.75)  # Adjust the resolution value as needed
merged_data_all_7 <- RunUMAP(object = merged_data_all_7, dims = 1:10)
DimPlot(merged_data_all_7, reduction = "umap") +
  scale_color_viridis(discrete = TRUE) + plot_annotation(title = "0.75")
```
```{r}
saveRDS(merged_data_all_7, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/SmartSeqMerged.rds")
```

```{r}
DimPlot(merged_data_all, reduction = "umap") +
  scale_color_viridis(discrete = TRUE)
```
UMAP by CellType

```{r}
DimPlot(merged_data_all_7, reduction = "umap", group.by = "CellType") +
  scale_color_viridis(discrete = TRUE) + plot_annotation(title = "0.75 by Cell Type")
```
UMAP by WellCellType
```{r}
# Plot UMAP with clusters associated with shapes
wellcell_plot <- DimPlot(merged_data_all_7, reduction = "umap", group.by = "WellCellType", shape.by = "CellType") +
  scale_color_viridis(discrete = TRUE) + plot_annotation(title = "0.75 by Well Cell Type")


# Apply the viridis color scale and change the title
wellcell_plot + scale_color_viridis_d() + ggtitle("Well Cell Types by Cell Origin") + 
  labs(x = NULL, y = NULL) 


```

Plot metadata
```{r}
VlnPlot(merged_data_all, features = c("nFeature_RNA","nCount_RNA")) 
```

```{r}
VlnPlot(merged_data_all_7, features = c("nFeature_RNA", "nCount_RNA"), group.by = "CellType") +
  scale_fill_viridis_d()
```

```{r}
# Load necessary library
library(ggplot2)
# Create UMAP plot and extract colors
umap_plot <- DimPlot(merged_data_all, reduction = "umap", group.by = "CellType")
umap_colors <- scales::hue_pal()(length(unique(merged_data_all$CellType)))

# Extract metadata
metadata <- merged_data_all@meta.data

# Create the histogram with matching colors
ggplot(metadata, aes(x = CellType, fill = CellType)) +
  geom_bar() +
  scale_fill_manual(values = umap_colors) +
  theme_minimal() +
  labs(title = "Number of Unique Cells for Each Cell Origin",
       x = "CellType",
       y = "Count")
```

```{r}
# Plot UMAP with clusters associated with shapes
#DimPlot(merged_data_all_7, reduction = "umap", group.by = "CellType", shape.by = "seurat_clusters")
```

```{r}
# Plot UMAP with clusters associated with shapes
umap_plot <- DimPlot(merged_data_all_7, reduction = "umap", shape.by = "CellType", group.by = "seurat_clusters") +
  scale_color_viridis(discrete = TRUE)
# NOT ENOUGH SHAPES
# Apply the viridis color scale and change the title
umap_plot + scale_color_viridis_d() + ggtitle("Clusters by Cell Origin") + 
  labs(x = NULL, y = NULL) 
```

```{r}
library(viridis)
library(ggplot2)

# Create UMAP plot with viridis colors
umap_plot <- DimPlot(merged_data_all_7, reduction = "umap", group.by = "CellType") +
  scale_color_viridis_d()

# Extract the colors used in the UMAP plot
umap_colors <- viridis::viridis_pal()(length(unique(merged_data_all_7$CellType)))

# Extract metadata
metadata <- merged_data_all_7@meta.data

# Create the histogram with matching viridis colors
ggplot(metadata, aes(x = CellType, fill = CellType)) +
  geom_bar() +
  scale_fill_manual(values = umap_colors) +
  theme_minimal() +
  labs(title = "Number of Unique Cells for Each Cell Origin",
       x = "CellType",
       y = "Count") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

#Go Terms

```{r}
# Load necessary libraries
library(Seurat)
library(clusterProfiler)
library(org.Hs.eg.db)
library(viridis)

# Assuming you have a Seurat object named 'merged_data_all'

# Join data layers
merged_data_all_joined <- JoinLayers(merged_data_all_7)
# Identify DEGs for each cluster
degs_list <- list()
for (i in 0:4) {
  degs_list[[i + 1]] <- FindMarkers(merged_data_all_joined, ident.1 = i)
}
```
```{r}
# Perform GO term enrichment analysis for each cluster
ego_list <- list()
for (i in 0:4) {
  gene_list <- rownames(degs_list[[i + 1]])
  ego_list[[i + 1]] <- enrichGO(gene = gene_list, 
                                OrgDb = org.Hs.eg.db, 
                                keyType = "SYMBOL", 
                                ont = "BP",  # Biological Process
                                pAdjustMethod = "BH", 
                                pvalueCutoff = 0.01, 
                                qvalueCutoff = 0.05)
}
```
```{r}
# Combine the results into a single data frame for visualization
library(dplyr)
combined_results <- bind_rows(lapply(1:5, function(i) {
  ego_list[[i]]@result %>% mutate(Cluster = i - 1)
}))
```
```{r}
# Filter to include only the top 20 GO terms based on adjusted p-value
top_terms <- combined_results %>%
  group_by(Cluster) %>%
  top_n(-20, wt = p.adjust) %>%
  ungroup()

# Visualize the results using a dot plot with viridis color scale
library(ggplot2)
p <- ggplot(top_terms, aes(x = Cluster, y = Description, size = Count, color = p.adjust)) +
  geom_point() +
  theme_minimal() +
  scale_color_viridis_c() +
  labs(title = "Top 20 GO Term Enrichment Across Clusters",
       x = "Cluster",
       y = "GO Term",
       size = "Gene Count",
       color = "Adjusted p-value")

# Save the plot with specific dimensions
ggsave("Top_20_GO_Term_Enrichment.png", plot = p, width = 24, height = 16, dpi = 300)
```
### Investigating Specific Genes of Interest

```{r}
library(Seurat)
library(viridis)
library(ggplot2)
genes_of_interest <- c("KRAS", "TP53","MLH1")
plots <- lapply(genes_of_interest, function(gene) {
  FeaturePlot(merged_data_all_7, features = gene) + scale_color_viridis_c()
})

# Display all plots
cowplot::plot_grid(plotlist = plots)
```
```{r}
library(Seurat)
library(viridis)
library(ggplot2)
genes_of_interest <- c("EGFR","MYC","APC")
plots <- lapply(genes_of_interest, function(gene) {
  FeaturePlot(merged_data_all_7, features = gene) + scale_color_viridis_c()
})

# Display all plots
cowplot::plot_grid(plotlist = plots)
```
```{r}
library(Seurat)
library(viridis)
library(ggplot2)
genes_of_interest <- c("CD14","CD68","CD163")
plots <- lapply(genes_of_interest, function(gene) {
  FeaturePlot(merged_data_all_7, features = gene) + scale_color_viridis_c()
})

# Display all plots
cowplot::plot_grid(plotlist = plots)
```
```{r}
library(Seurat)
library(viridis)
library(ggplot2)
genes_of_interest <- c("ANXA1","IFI44","IFI44L")
plots <- lapply(genes_of_interest, function(gene) {
  FeaturePlot(merged_data_all_7, features = gene) + scale_color_viridis_c()
})

# Display all plots
cowplot::plot_grid(plotlist = plots)
```
```{r}
library(Seurat)
library(viridis)
library(ggplot2)
genes_of_interest <- c("SPI1", "RUNX1", "CDH1", "PTPRC")
plots <- lapply(genes_of_interest, function(gene) {
  FeaturePlot(merged_data_all_7, features = gene) + scale_color_viridis_c()
})

# Display all plots
cowplot::plot_grid(plotlist = plots)
```
