---
title: "ArchRInstall&Attempt1"
author: "KQ"
date: "2024-04-04"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Install Packages and Dependencies
```{r}
#if (!requireNamespace("devtools", quietly = TRUE)) install.packages("devtools")
```
```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE)) install.packages("BiocManager")

#BiocManager::install("chromVAR")

```
```{r}
#devtools::install_github("GreenleafLab/ArchR", ref="master", repos = BiocManager::repositories())
```
```{r}
# install.packages("Matrix")
# Specify the version you want
#version <- "1.6-3" #THIS DID NOT WORK, niether did 1.6-1, 1.6-2
# Install that version from CRAN
#devtools::install_version("Matrix", version = version)
```
```{r}
#IRLBA MUST BE INSTALLED AFTER MATRIX 
# 
# remove.packages("Matrix")
# remove.packages("irlba")
```
```{r} 
# #https://github.com/cole-trapnell-lab/monocle3/issues/690
# install.packages("Matrix",type="source", dependencies=T)
# ```
# ```{r}
# install.packages("irlba",type="source", dependencies=T)
```
```{r}
# Get a data frame of installed packages
#installed_packages <- as.data.frame(installed.packages()[, c("Package", "Version")])

# Print the data frame
#print(installed_packages, row.names = FALSE)
```
```{r}
#if (!requireNamespace("BiocManager", quietly = TRUE))
#    install.packages("BiocManager")
#BiocManager::install("Rsamtools")
```

##Creating the ArchR Project
#Launch ArchR
```{r}
library(ArchR)
#ArchR::installExtraPackages()
library(Rsamtools)
library(irlba)
library(dplyr)
library(stringr)

library(pheatmap)
library(ggplot2)
library(hexbin)
library(magick)
library(grid)
library(ggseqlogo)

library(ggimage)
library(TFBSTools)
library(grid)
```

```{r}
set.seed(1)
```

#Read Reference Genome
```{r}
addArchRGenome("mm10")
```
#Input Files
```{r}

setwd("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/ATAC_Data/ArchRInputData")

```

```{r}
# List of BAM files
bamfiles <- c("221025_iCell8_ATAC.bbrd.q10.filt.bam")
```

```{r}
# Prepend the directory path to each file name
#baminput <- paste0(getwd(), "/", bamfiles)
```

```{r}
PathToSampleNames <- "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/ATAC_Data/ArchRInputData/240715_iCell8_ATAC.bbrd.q10.filt.samplelist.dummy.txt"

sampleNamesInput <- readLines(PathToSampleNames)
```

```{r}
# # Set the path to your .list file
# listFilePath <- "./221025_iCell8_ATAC.bbrd.q10.filt.cellIDslist.txt"
# # Read the file into R and convert it to a character vector
# validBarcodes <- readLines(listFilePath)
```
#Test size and structure of input
```{r}
# class(baminput)
# system(paste("wc -l ", baminput, sep = ""))
# 
# print(sampleNamesInput)
```
#Create ArrowFile
```{r}
ArrowFiles805 <- createArrowFiles(
  inputFiles = bamfiles,
  sampleNames = sampleNamesInput,
  gsubExpression=":.*",
  #validBarcodes = validBarcodes,
  minTSS = 4, #Dont set this too high because you can always increase later
  minFrags = 1000,
  addTileMat = TRUE,
  addGeneScoreMat = TRUE,
  #bcTag = "CB", # This is the tag that contains the cell-specific barcode information in your BAM file
  #bamFlag = list(isMinusStrand = FALSE, isProperPair = TRUE, isDuplicate = FALSE)
)
```
#Check it's a character vector
```{r}
class(ArrowFiles805)
file.exists(ArrowFiles805)
```
#Creating ArchRProject
```{r}
projHybrid805 <- ArchRProject(
  ArrowFiles = ArrowFiles805, 
  outputDirectory = "projHybrid805",
  copyArrows = TRUE #This is recommened so that if you modify the Arrow files you have an original copy for later usage.
)
```
#Save ArchRProject
```{r}
saveArchRProject(ArchRProj = projHybrid805, outputDirectory = "Save-projHybrid805", load = FALSE)
```

##Load ArchRProject

```{r}
#Shortern
proj <- projHybrid805
```

```{r}
# Load your custom annotation data
customAnnotations <- read.table("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/ATAC_Data/ArchRInputData/240418_KQ_bysample.annot.csv", header = FALSE, sep = "\t")

```

```{r}
# Get cellColData
cellColData <- getCellColData(proj)
```
```{r}
# Get cell IDs from cellColData
cellColDataIDs <- rownames(cellColData)
```

```{r}
barcodes <- sapply(strsplit(rownames(cellColData), "#"), "[", 2)
```

```{r}
barcodes_df <- data.frame(barcodes = barcodes)
```

```{r}
merged_df <- merge(barcodes_df, customAnnotations, by.x = "barcodes", by.y = "V1")
```

#Adding the barcodes column to the cellColData
```{r}
# Using the $ operator
cellColData$barcodes <- barcodes
```

#Sanity Check - the first five barcodes match those from the row names!
```{r}
# Get the first five rows of cellColData
first_five_rows <- head(cellColData, 5)

# Get the row names
rownames_first_five <- rownames(first_five_rows)

# Get the barcodes (assuming 'barcodes' is the column name for barcodes)
barcodes_first_five <- first_five_rows$barcodes

# Print the row names and barcodes
print(rownames_first_five)
print(barcodes_first_five)

```

```{r}
# Assuming 'barcodes' is the name of the column in both dataframes
final_df <- merge(cellColData, merged_df[, c("barcodes", "V2")], by = "barcodes", all.x = TRUE)
```

```{r}
# Assuming merged_df is your dataframe
# Rename "V2" to "Cell Type"
#matchmerged_df <- rename(merged_df, `Cell Type` = V2)


# Rename "V2" to "Cell Type"
matchmerged_df <- merged_df %>% rename_with(~ "Cell Type", V2)
colnames(merged_df)


# Assuming 'df' is your data frame and 'barcodes' is the column with the barcodes
matchmerged_df$row_names <- paste("HybridATACExperiment#", matchmerged_df$barcodes, sep="")

# If you want to set these as actual row names in the dataframe
rownames(matchmerged_df) <- matchmerged_df$row_names

```

```{r}
# Assuming 'matchedmerged_df' is your dataframe with new columns and 'proj' is your ArchR project
# Make sure the rownames of df match the cell names in the ArchR project

# Convert dataframe to named a character
matchmerged_char <- as.character(matchmerged_df)

```

```{r}
class(matchmerged_char)
```

```{r}
proj <- addCellColData(
  ArchRProj = proj, # The ArchR project
  data = matchmerged_df$`Cell Type`,  # The new columns
  cells = matchmerged_df$row_names, #THANK YOU KEVIN 
  name = "CellType", # The name of the table in the ArchR project to which the columns will be added
  force = FALSE
)
```

#Save ArchRProject
```{r}
saveArchRProject(ArchRProj = proj, outputDirectory = "Save-projHybrid805.2", load = FALSE)
```

#Access TSS-E Score
```{r}
quantile(proj$TSSEnrichment)
```

##QC and Initial Visualization
#Log10 Unique Fragments vs. TSS
```{r}
logvsTSS <- getCellColData(proj, select = c("log10(nFrags)", "TSSEnrichment"))
logvsTSS
colnames(logvsTSS)
# Convert DFrame to data.frame
logvsTSS_df <- as.data.frame(logvsTSS)
```

```{r}
summary(logvsTSS)
```

#Visualize LogvsTSS
```{r}
# p <- ggPoint(
#     x = logvsTSS[,"log10(nFrags)"], 
#     y = logvsTSS[,"TSSEnrichment"], 
#     colorDensity = TRUE,
#     continuousSet = "sambaNight",
#     xlabel = "Log10 Unique Fragments",
#     ylabel = "TSS Enrichment",
#     xlim = c(log10(500), quantile(logvsTSS[,"log10(nFrags)"], probs = 0.99)),
#     ylim = c(0, quantile(logvsTSS[,"TSSEnrichment"], probs = 0.99))
# ) + geom_hline(yintercept = 4, lty = "dashed") + geom_vline(xintercept = 3, lty = "dashed")

#Attempt to use ggplot2 instead of ggPoint
library(ggplot2)

# p <- ggplot(logvsTSS_df, aes(x = log10(nFrags), y = TSSEnrichment)) +
#     geom_point() +
#     scale_x_continuous(limits = c(log10(500), stats::quantile(logvsTSS_df$log10(nFrags), probs = 0.99))) +
#     scale_y_continuous(limits = c(0, stats::quantile(logvsTSS_df$TSSEnrichment, probs = 0.99))) +
#     geom_hline(yintercept = 4, lty = "dashed") +
#     geom_vline(xintercept = 3, lty = "dashed") +
#     labs(x = "Log10 Unique Fragments", y = "TSS Enrichment")

# Get summary statistics
summary_stats <- summary(logvsTSS_df)

# Calculate 99th percentile for each variable
quantile_nFrags <- quantile(logvsTSS_df$log10.nFrags., probs = 0.99)
quantile_TSSEnrichment <- quantile(logvsTSS_df$TSSEnrichment, probs = 0.99)

p <- ggplot(logvsTSS_df, aes(x = log10.nFrags., y = TSSEnrichment)) +
    geom_point() +
    scale_x_continuous(limits = c(log10(500), quantile_nFrags)) +
    scale_y_continuous(limits = c(0, quantile_TSSEnrichment)) +
    geom_hline(yintercept = 4, lty = "dashed") +
    geom_vline(xintercept = 3, lty = "dashed") +
    labs(x = "Log10 Unique Fragments", y = "TSS Enrichment")

print(p)


plotPDF(p, name = "TSS-vs-Frags.pdf", ArchRProj = proj, addDOC = FALSE)
```
#Trouble Shooting ggplot 
```{r}
#any(is.na(logvsTSS_df))
#summary(logvsTSS_df)
#str(logvsTSS_df)
#This all looked good
```

#Ridge Plot for TSS
```{r}
p1 <- plotGroups(
    ArchRProj = proj, 
    groupBy = "Sample", 
    colorBy = "cellColData", 
    name = "TSSEnrichment",
    plotAs = "ridges"
   )
p1
```
#Violin Plot for TSS
```{r}
p2 <- plotGroups(
    ArchRProj = proj, 
    groupBy = "Sample", 
    colorBy = "cellColData", 
    name = "TSSEnrichment",
    plotAs = "violin",
    alpha = 0.4,
    addBoxPlot = TRUE
   )

p2
#ggsave(p2)
```
#Ridge Plot per Sample Log10 Reads
```{r}
p3 <- plotGroups(
    ArchRProj = proj, 
    groupBy = "Sample", 
    colorBy = "cellColData", 
    name = "log10(nFrags)",
    plotAs = "ridges"
   )
p3
#ggsave(p3)
```
#Violin Plot per Sample Log10 Reads
```{r}
p4 <- plotGroups(
    ArchRProj = proj, 
    groupBy = "Sample", 
    colorBy = "cellColData", 
    name = "log10(nFrags)",
    plotAs = "violin",
    alpha = 0.4,
    addBoxPlot = TRUE
   )
p4
#ggsave(p4)
```
#3.4: Size Distribution and TSS E Profiles
```{r}
p5 <- plotFragmentSizes(ArchRProj = proj)
p5
```
#3.4: ...and TSS E Profiles
```{r}
p6 <- plotTSSEnrichment(ArchRProj = proj)
p6
```
#Save ArchRProject
```{r}
saveArchRProject(ArchRProj = proj, outputDirectory = "Save-projHybrid805.2", load = FALSE)
```

```{r}
proj <- loadArchRProject(path = "Save-projHybrid805.2")
```

##Filter doublets (SKIP: Previously Done in BAM Generation)
```{r}
# projHybrid2 <- filterDoublets(projHybrid1)
# projHeme2
# projHemeTmp <- filterDoublets(projHybrid1, filterRatio = 1.5)
```

##Dimensionality Reduction w/ ArchR
```{r}
proj <- addIterativeLSI(
    ArchRProj = proj,
    useMatrix = "TileMatrix", 
    name = "IterativeLSI", 
    iterations = 2, 
    clusterParams = list( #See Seurat::FindClusters
        resolution = c(0.2), 
        sampleCells = 10000, 
        n.start = 10
    ), 
    varFeatures = 25000, 
    dimsToUse = 1:30
)
```
##Clustering*

```{r}
#install.packages("hexbin")
library(hexbin)
```


```{r}
proj <- addClusters(
    input = proj,
    reducedDims = "IterativeLSI",
    method = "Seurat",
    name = "Clusters",
    resolution = 0.8
)
```

```{r}
head(proj$Clusters)
table(proj$Clusters)
```

```{r}
cM <- confusionMatrix(paste0(proj$Clusters), paste0(proj$CellType))

```
#Clustering Heatmap
```{r} 
library(pheatmap)
cM <- cM / Matrix::rowSums(cM)
hm1 <- pheatmap::pheatmap(
    mat = as.matrix(cM),
    color = paletteContinuous("whiteBlue"),
    border_color = "black")

```
# Add UMAP
```{r}
proj <- addUMAP(
    ArchRProj = proj, 
    reducedDims = "IterativeLSI", 
    name = "UMAP", 
    nNeighbors = 30, 
    minDist = 0.5, 
    metric = "cosine"
)
```

```{r}


umap1 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "CellType", embedding = "UMAP")

umap1


```

```{r}
# Define the cluster mapping
cluster_mapping <- c(
  "C2" = "Macrophage",
  "C4" = "Tumor-cluster1",
  "C1" = "Hybrid-M",
  "C3" = "Hybrid-T",
  "C5" = "Tumor-cluster2",
  "C6" = "Tumor-cluster3"
)

# Add the new column to the ArchRProj
proj <- addCellColData(
  ArchRProj = proj,
  data = cluster_mapping[proj$Clusters],
  name = "UpdatedClusters",
  cells = getCellNames(proj),
  force = TRUE
)

customColors <- c(
  "Macrophage" = "cadetblue3",
  "Tumor-cluster1" = "grey45",
  "Tumor-cluster2" = "grey45",
  "Tumor-cluster3" = "grey45",
  "Hybrid-M" = "blue4",
  "Hybrid-T" = "red3")
# Ensure UpdatedClusters is a factor with levels matching customColors
proj$UpdatedClusters <- factor(proj$UpdatedClusters, levels = names(customColors))
# Check the levels of UpdatedClusters
levels(proj$UpdatedClusters)

#umap2 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "CellType", embedding = "UMAP")
umap3 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "UpdatedClusters", embedding = "UMAP", colorList = customColors)
#umap2
umap3

# Extract UMAP coordinates and cluster information
umap_df <- as.data.frame(getEmbedding(proj, embedding = "UMAP"))
umap_df$UpdatedClusters <- proj$UpdatedClusters

# Rename columns for easier plotting
colnames(umap_df) <- c("UMAP_Dimension_1", "UMAP_Dimension_2", "UpdatedClusters")

# Create UMAP plot using ggplot2
ggplot(umap_df, aes(x = UMAP_Dimension_1, y = UMAP_Dimension_2, color = UpdatedClusters)) +
  geom_point(size = 1) +
  scale_color_manual(values = customColors) +
  theme_minimal() +
  labs(title = "UMAP of IterativeLSI colored by Updated Clusters")
```

# Simplifying Clusters
```{r}
#Confirming Current Cluster Assignments
table(proj$UpdatedClusters)
# Combining Tumor Clusters
newClusters <- as.character(proj$UpdatedClusters)
newClusters[newClusters %in% c("Tumor-cluster1", "Tumor-cluster2", "Tumor-cluster3")] <- "Tumor"
proj$Clusters_TuMerged <- newClusters
#Confirming Tumor Merged Cluster Assignments
table(proj$Clusters_TuMerged)
```
#Remaking UMAP w/ Merged Tumor Clusters
```{r}
customColors <- c(
  "Macrophage" = "cadetblue3",
  "Tumor" = "grey45",
  "Hybrid-M" = "blue4",
  "Hybrid-T" = "red3")
# Ensure UpdatedClusters is a factor with levels matching customColors
proj$Clusters_TuMerged <- factor(proj$Clusters_TuMerged, levels = names(customColors))
# Check the levels of UpdatedClusters
levels(proj$Clusters_TuMerged)

#umap2 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "CellType", embedding = "UMAP")
umap3 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "Clusters_TuMerged", embedding = "UMAP", colorList = customColors)
#umap2
umap3

# Extract UMAP coordinates and cluster information
umap_df <- as.data.frame(getEmbedding(proj, embedding = "UMAP"))
umap_df$Clusters_TuMerged <- proj$Clusters_TuMerged

# Rename columns for easier plotting
colnames(umap_df) <- c("UMAP_Dimension_1", "UMAP_Dimension_2", "Clusters_TuMerged")

# Create UMAP plot using ggplot2
ggplot(umap_df, aes(x = UMAP_Dimension_1, y = UMAP_Dimension_2, color = Clusters_TuMerged)) +
  geom_point(size = 1) +
  scale_color_manual(values = customColors) +
  theme_minimal() +
  labs(title = "UMAP of IterativeLSI colored by Simplified Clusters")

library(ggplot2)

ggplot(umap_df, aes(x = UMAP_Dimension_1, y = UMAP_Dimension_2, color = Clusters_TuMerged)) +
  geom_point(size = 1) +
  scale_color_manual(values = customColors) +
  coord_fixed() +
  theme_minimal() +
  theme(
    axis.line = element_line(color = "black"),  # Makes both axes black
    axis.ticks = element_line(color = "black"), # Optional: makes tick marks black too
    axis.text = element_text(color = "black"),  # Optional: makes axis labels black
    axis.title = element_text(color = "black")  # Optional: makes axis titles black
  ) +
  labs(title = "UMAP of IterativeLSI with Black Axes")


```
# TSNE
```{r}
proj <- addTSNE(
    ArchRProj = proj, 
    reducedDims = "IterativeLSI", 
    name = "TSNE", 
    perplexity = 30
)
```
```{r}

tsne1 <- plotEmbedding(ArchRProj = proj, colorBy = "cellColData", name = "CellType", embedding = "TSNE")

tsne1
```

# Save ArchRProject
```{r}
saveArchRProject(ArchRProj = proj, outputDirectory = "Save-projHybrid805.2", load = FALSE)
```

##Marker Genes

# Load ArhcRProject if Needed
```{r}
proj <- loadArchRProject(path = "Save-projHybrid805.2")
```

```{r}
markersGS <- getMarkerFeatures(
    ArchRProj = proj, 
    useMatrix = "GeneScoreMatrix", 
    groupBy = "Clusters",
    bias = c("TSSEnrichment", "log10(nFrags)"),
    testMethod = "wilcoxon"
)

markerList <- getMarkers(markersGS, cutOff = "FDR <= 0.01 & Log2FC >= 1.25")
markerList$C5
```

```{r}
#install.packages("magick")
#library(magick)
```

```{r}
markerGenes  <- c(
    "CCR5", "CD14", "CD68", #Macrophage
    #MC38
    "RUNX1", "RUNX2", "NR1H3", "NTRK1", "BICDL2", "MIR99B", "Mtss1l", "Gng13" #Hybrid Fusion
)

heatmapGS <- plotMarkerHeatmap(
  seMarker = markersGS, 
  cutOff = "FDR <= 0.01 & Log2FC >= 1.25", 
  labelMarkers = markerGenes,
  transpose = TRUE
)
heatmapGS
```

```{r}
ComplexHeatmap::draw(heatmapGS, heatmap_legend_side = "bot", annotation_legend_side = "bot")

```

```{r}
plotPDF(heatmapGS, name = "GeneScores-Marker-Heatmap", width = 8, height = 6, ArchRProj = proj, addDOC = FALSE)
```

```{r}
umap.marker <- plotEmbedding(
    ArchRProj = proj, 
    colorBy = "GeneScoreMatrix", 
    name = markerGenes, 
    embedding = "UMAP",
    quantCut = c(0.01, 0.95),
    imputeWeights = NULL
)

umap.marker
```

```{r}
umap.allgenes <- lapply(umap.marker, function(x){
    x + guides(color = FALSE, fill = FALSE) + 
    theme_ArchR(baseSize = 6.5) +
    theme(plot.margin = unit(c(0, 0, 0, 0), "cm")) +
    theme(
        axis.text.x=element_blank(), 
        axis.ticks.x=element_blank(), 
        axis.text.y=element_blank(), 
        axis.ticks.y=element_blank()
    )
})
do.call(cowplot::plot_grid, c(list(ncol = 3),umap.allgenes))
```

#MAGIC
```{r}
proj <- addImputeWeights(proj)
```

```{r}
magicplot <- plotEmbedding(
    ArchRProj = proj, 
    colorBy = "GeneScoreMatrix", 
    name = markerGenes, 
    embedding = "UMAP",
    imputeWeights = getImputeWeights(proj)
)
magicplot
```

# Save ArchRProject
```{r}
saveArchRProject(ArchRProj = proj, outputDirectory = "Save-projHybrid805.2", load = FALSE)
```

#Track Plotting
```{r}
proj <- loadArchRProject(path = "Save-projHybrid805.2")
```

```{r}
trackplot <- plotBrowserTrack(
    ArchRProj = proj, 
    groupBy = "Clusters", 
    geneSymbol = markerGenes, 
    upstream = 50000,
    downstream = 50000
)
```

```{r}
grid::grid.newpage()
grid::grid.draw(trackplot$Runx1)
```

#Renaming the Clusters was done here... [yes so the "fresh" hybrids that were from the plates we sorted that day without expanding the GFP/RFP double positive cells are more close to macrophages. and the hybrids from 7-1-22 experiment are the ones that had been selected over time in culture and sorted for double positive RFP+GFP over time]
```{r}
# Define the cluster mapping
cluster_mapping <- c(
  "C2" = "Macrophage",
  "C4" = "Tumor-cluster1",
  "C1" = "Hybrid-M",
  "C3" = "Hybrid-T",
  "C5" = "Tumor-cluster2",
  "C6" = "Tumor-cluster3"
)

# Add the new column to the ArchRProj
proj <- addCellColData(
  ArchRProj = proj,
  data = cluster_mapping[proj$Clusters],
  name = "UpdatedClusters",
  cells = getCellNames(proj),
  force = TRUE
)

```
#Trouble Shooting Steps
```{r} 
#table(archrProj$UpdatedClusters) #worked
#head(proj$UpdatedClusters)
#table(proj$UpdatedClusters)
```

```{r}
# Define your custom color palette
customColors <- c(
  "Macrophage" = "cadetblue3",
  "Tumor-cluster1" = "grey45",
  "Tumor-cluster2" = "grey45",
  "Tumor-cluster3" = "grey45",
  "Hybrid-M" = "darkmagenta",
  "Hybrid-T" = "red3"
)
# Plot UMAP by Cell Type with custom colors
celltypecolorsnice <- plotEmbedding(
  ArchRProj = proj,
  colorBy = "cellColData",
  name = "CellType",
  embedding = "UMAP",
  colorList = customColors
)
# Plot UMAP by Updated Clusters
clusterstypecolorsnice<- plotEmbedding(
  ArchRProj = proj,
  colorBy = "cellColData",
  name = "UpdatedClusters",
  embedding = "UMAP"
)

# Display the plots
plotPDF(celltypecolorsnice, clusterstypecolorsnice, name = "UMAP_CellType_UpdatedClusters.pdf", ArchRProj = proj, addDOC = FALSE)


```

#Defining Custome Colors to Match the Manuscript that Ashley's Been Working On
```{r}
library(ArchR)
library(grid)
# Define your custom color palette
customColors <- c(
  "Macrophage" = "cadetblue3",
  "Tumor" = "grey45",
  "Hybrid-M" = "blue4", #blue4
  "Hybrid-T" = "red3")
# Consider looking at Pf4, Spi1, Thbs1
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Runx1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Runx1)
dev.off()  # Close current device if open
dev.new()  # Open a new device
plot(AS1211$Runx1)
# Troubleshooting
# print(AS1211)
# str(AS1211)
#print(AS1211$Runx1)
```
#20251103:AA Supplemental Asks
```{r}
# Define the path to your folder
output_dir <- "20251103SupplementalGeneTracks"
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Epcam",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check CD CD45, CD11b, CD163, CD14 and Epcam, Ecad 
#Try Cd45, Cd11b, Cd163, Cd14, Epcam, Ecad
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Epcam_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Epcam)
dev.off()
```
```{r}
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Cdh1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check CD CD45, CD11b, CD163, CD14 and Epcam, Ecad 
#Try Cd45, Cd11b, Cd163, Cd14, Epcam, Ecad
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Cdh1_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Cdh1)
dev.off()
```
#20251201:AA Supplemental Asks
```{r}
# Define the path to your folder
output_dir <- "20251201SupplementalGeneTracks"
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Cldn1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Cldn1, Satb1, Satb2
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Cldn1_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Cldn1)
dev.off()
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Satb1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Cldn1, Satb1, Satb2
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Satb1_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Satb1)
dev.off()
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Satb2",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Cldn1, Satb1, Satb2
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Satb2_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Satb2)
dev.off()
```
#20251201:AA Supplemental Asks 2.0
```{r}
# Just compare two groups at a time - this usually works
# First: Hybrid-T + Tumor vs Others
proj$TwoGroup <- ifelse(
  proj$Clusters_TuMerged %in% c("Hybrid-T", "Tumor"),
  "HybridT_Tumor",
  "Others"
)

# Get markers without bias correction
markers <- getMarkerFeatures(
  ArchRProj = proj,
  useMatrix = "GeneScoreMatrix",
  groupBy = "TwoGroup",
  useGroups = "HybridT_Tumor",
  bgdGroups = "Others",
  testMethod = "wilcoxon"
)

# Extract significant genes
markerList <- getMarkers(markers, cutOff = "FDR <= 0.05 & Log2FC >= 0.5")
target_genes <- markerList$HybridT_Tumor$name

print(paste("Found", length(target_genes), "candidate genes"))
print(head(target_genes, 25))
# Plot top genes
for(gene in head(target_genes, 5)) {
  p <- plotBrowserTrack(
    ArchRProj = proj,
    groupBy = "Clusters_TuMerged",
    geneSymbol = gene,
    upstream = 50000,
    downstream = 50000,
    pal = customColors
  )
  
  pdf(paste0(output_dir, "/", gene, "_HybridT_Tumor.pdf"), width = 10, height = 8)
  grid::grid.newpage()
  grid::grid.draw(p[[gene]])
  dev.off()
  
  print(paste("Plotted:", gene))
}
```
#20251201:AA Supplemental Asks Part 3.0
```{r}
# Define the path to your folder
output_dir <- "20251201SupplementalGeneTracks"
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Fosl1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Fosl1, Col6a1, Fscn1
# Draw the plot
# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Fosl1_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Fosl1)
dev.off()
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Col6a1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Cldn1, Satb1, Satb2
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Col6a1_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Col6a1)
dev.off()
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Fscn1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Cldn1, Satb1, Satb2, Cyr61
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Fscn1_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Fscn1)
dev.off()
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Cyr61",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
#Check Cldn1, Satb1, Satb2, Cyr61
# Draw the plot

# Create the directory if it doesn't exist
if (!dir.exists(output_dir)) {
  dir.create(output_dir, recursive = TRUE)
}
# Save the plot as PDF
pdf(file.path(output_dir, "Cyr61_track.pdf"), width = 10, height = 8)
grid::grid.newpage()
grid::grid.draw(AS1211$Cyr61)
dev.off()
```
# Tracks for Other Genes of Interest
```{r}
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Spi1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Pf4, Spi1, Thbs1
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Spi1)
```
```{r}
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Pf4",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Pf4, Spi1, Thbs1
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Pf4)
```
```{r}
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Thbs1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Pf4, Spi1, Thbs1
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Thbs1)
```
```{r}
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Grn",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Pf4, Spi1, Thbs1
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Grn)
```
```{r}
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "Clusters_TuMerged",
  geneSymbol = "Lgals3",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Pf4, Spi1, Thbs1
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Lgals3)
```

```{r}
# For Top 10 DGE for the Hybrid-M Cluster
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Apoc2",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Apoc2, Lyz2, Mpeg1, Wldc17, Trobp, C1qa, C1qb, C1qc
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Apoc2)
#Macropage and Hybrid-M look very similar. Hybrid T and the Tumors look similar for Apoc2.
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Lyz2",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Apoc2, Lyz2, Mpeg1, Wldc17, Trobp, C1qa, C1qb, C1qc
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Lyz2)
#Macropage and Hybrid-M look very similar. Hybrid T and the Tumors look similar for Lyz2.
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Mpeg1",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Apoc2, Lyz2, Mpeg1, Wldc17, Trobp, C1qa, C1qb, C1qc
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$Mpeg1)
#Macropage and Hybrid-M look very similar. Hybrid T and the Tumors look similar for Mpeg1 - actually striking for tumor/hybrids-m
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "C1qa",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
# Consider looking at Apoc2, Lyz2, Mpeg1, Wldc17(doesn't work), Trobp(doesn't work), C1qa, C1qb, C1qc
# Draw the plot
grid::grid.newpage()
grid::grid.draw(AS1211$C1qa)
#Macropage and Hybrid-M look very similar. 
```
```{r}
#Top 10 DGE for Hybrid-T: Gsk3a, Hacd3, Ralgapb, Spire1, Tmem128, Eif2ak4, Prim2, Ythdf2, Cnot9
#Unpersausive: Gsk3a, Hacd3, 
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Ralgapb",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
grid::grid.newpage()
grid::grid.draw(AS1211$Ralgapb)
#Macropage and Hybrid-T both have sharper, taller peaks
# Create the track plot with custom colors
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Prim2",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
grid::grid.newpage()
grid::grid.draw(AS1211$Prim2)
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Ythdf2",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
grid::grid.newpage()
grid::grid.draw(AS1211$Ythdf2)
AS1211 <- plotBrowserTrack(
  ArchRProj = proj,
  groupBy = "UpdatedClusters",
  geneSymbol = "Eif2ak4",
  upstream = 50000,
  downstream = 50000,
  pal = customColors)
grid::grid.newpage()
grid::grid.draw(AS1211$Eif2ak4)
```
#Launching ArchRBrowser
```{r}
#ArchRBrowser(projHybridA)
```


##Peak Calling

```{r}
pathToMacs2 <- "/Users/queitsch/opt/anaconda3/bin/MACS2"
```

```{r}
#BiocManager::install("BSgenome.Mmusculus.UCSC.mm10")
```

```{r}

# Error in .getCoverageMetadata(ArchRProj = ArchRProj, groupBy = groupBy,  : No Coverage Metadata found for : Clusters. Please run addGroupCoverages!

# To address this... I'll try to run addGroupCoverages first
proj <- addGroupCoverages(ArchRProj = proj, groupBy = "Clusters_TuMerged")
```

```{r}
proj_motif <- addReproduciblePeakSet(
    ArchRProj = proj, 
    groupBy = "Clusters_TuMerged", 
    pathToMacs2 = pathToMacs2
)
```

```{r}
saveArchRProject(ArchRProj = proj_motif, outputDirectory = "Save-ProjMotif", load = TRUE)
#This worked for 250514: Complete
#Prior two steps take a very long time so be sure to save here!!!
```

```{r}
proj_motif <- loadArchRProject(path = "Save-ProjMotif")
```

```{r}
getPeakSet(proj_motif)
```

```{r}
proj_motif5 <- addPeakMatrix(proj_motif)
```

```{r}
getAvailableMatrices(proj_motif5)
```

##Identifying Marker Peaks with ArchR

```{r}
proj_motif5 <- addCellColData(
  ArchRProj = proj_motif5,
  data = as.character(proj_motif5$Clusters_TuMerged),
  name = "Clusters_TuMerged",
  cells = proj_motif5$cellNames,
  force = TRUE
)

```

```{r}
saveArchRProject(ArchRProj = proj_motif5, outputDirectory = "Save-ProjMotif", load = TRUE)
```

```{r}
proj_motif5 <- loadArchRProject(path = "Save-ProjMotif")
```

```{r}
markersPeaks <- getMarkerFeatures(
    ArchRProj = proj_motif5, 
    useMatrix = "PeakMatrix", 
    groupBy = "Clusters_TuMerged",
    bias = c("TSSEnrichment", "log10(nFrags)"),
    testMethod = "wilcoxon")
```

```{r}
markersPeaks
```

```{r}
markerList <- getMarkers(markersPeaks, cutOff = "FDR <= 0.01 & Log2FC >= 1")
markerList
```

```{r}
markerList$"Hybrid-M"
```

```{r}
markerListGR <- getMarkers(markersPeaks, cutOff = "FDR <= 0.01 & Log2FC >= 1", returnGR = TRUE)
markerListGR
```

```{r}
heatmapPeaks <- plotMarkerHeatmap(
  seMarker = markersPeaks, 
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5",
  transpose = TRUE
)
```

```{r}
draw(heatmapPeaks, heatmap_legend_side = "bot", annotation_legend_side = "bot")
```

```{r}
plotPDF(heatmapPeaks, name = "Peak-Marker-Heatmap", width = 8, height = 6, ArchRProj = proj_motif5, addDOC = FALSE)
```

```{r}
pma <- plotMarkers(seMarker = markersPeaks, name = "Hybrid-M", cutOff = "FDR <= 0.1 & Log2FC >= 1", plotAs = "MA")
pma
```

```{r}
pv <- plotMarkers(seMarker = markersPeaks, name = "Hybrid-M", cutOff = "FDR <= 0.1 & Log2FC >= 1", plotAs = "Volcano")
pv
```
#Optional to generate vectorized PDF
```{r}
#plotPDF(pma, pv, name = "C1-Markers-MA-Volcano", width = 5, height = 5, ArchRProj = projHeme5, addDOC = FALSE)
```

```{r}
p <- plotBrowserTrack(
    ArchRProj = proj_motif5, 
    groupBy = "Clusters_TuMerged", 
    geneSymbol = c("RUNX1"),
    features =  getMarkers(markersPeaks, cutOff = "FDR <= 0.1 & Log2FC >= 1", returnGR = TRUE)["Hybrid-M"],
    upstream = 50000,
    downstream = 50000
)
```

```{r}
grid::grid.newpage()
grid::grid.draw(p$RUNX1)
```

```{r}
plotPDF(p, name = "Plot-Tracks-With-Features", width = 5, height = 5, ArchRProj = proj_motif5, addDOC = FALSE)
```

#Pairwise testing between groups
```{r}
markerTest <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Macrophage"
)
```
# Plot Markers
```{r}
pma <- plotMarkers(seMarker = markerTest, name = "Hybrid-M", cutOff = "FDR <= 0.1 & abs(Log2FC) >= 1", plotAs = "MA")
class(pma)

pmv <- plotMarkers(seMarker = markerTest, name = "Hybrid-M", cutOff = "FDR <= 0.1 & abs(Log2FC) >= 1", plotAs = "Volcano")
class(pmv)

plotPDF(pma, name = "MA-Plot.pdf", ArchRProj = proj_motif5)
plotPDF(pmv, name = "Volcano-Plot.pdf", ArchRProj = proj_motif5)


plotPDF(pma, pmv, name = "Erythroid-vs-Progenitor-Markers-MA-Volcano", width = 5, height = 5, ArchRProj = proj_motif5, addDOC = FALSE)

```
## Motif Enrichment

```{r}
#devtools::install_github("GreenleafLab/chromVARmotifs")
```
# Set Motif Annotations (Crucial)
```{r}
#library(BSgenome.Mmusculus.UCSC.mm10)
proj_motif5 <- addMotifAnnotations(ArchRProj = proj_motif5, motifSet = "cisbp", name = "Motif")
```
# Save Motif Annotations to ArchR Project
```{r}
saveArchRProject(ArchRProj = proj_motif5, outputDirectory = "Save-ProjMotif", load = TRUE)
```
# [TBD]
```{r}
pSet <- getPeakSet(ArchRProj = proj_motif5)
pSet$name <- paste(seqnames(pSet), start(pSet), end(pSet), sep = "_")
```
```{r}
matches <- getMatches(ArchRProj = proj_motif5, name = "Motif")
rownames(matches) <- paste(seqnames(matches), start(matches), end(matches), sep = "_")
matches <- matches[pSet$name]
```
```{r}
motifsUp <- peakAnnoEnrichment(
    seMarker = markerTest,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )
```
# Enrichment Heatmap
```{r}
# plotEnrichHeatmap(motifsUp, n = 20, transpose = TRUE)
# pdf("HybirdM_Tumor_Motif_Enrichment_Heatmap.pdf", width = 8, height = 6)
# plotEnrichHeatmap(enrichMotifs, n = 20, transpose = TRUE)
# dev.off()
```
# Enrichment +
```{r}
df <- data.frame(TF = rownames(motifsUp), mlog10Padj = assay(motifsUp)[,1])
df <- df[order(df$mlog10Padj, decreasing = TRUE),]
df$rank <- seq_len(nrow(df))
```

```{r}
head(df)
```

```{r}
ggUp <- ggplot(df, aes(rank, mlog10Padj, color = mlog10Padj)) + 
  geom_point(size = 1) +
  ggrepel::geom_label_repel(
        data = df[rev(seq_len(30)), ], aes(x = rank, y = mlog10Padj, label = TF), 
        size = 1.5,
        nudge_x = 2,
        color = "black",
        max.overlaps = 30  # Increase max.overlaps
  ) + theme_ArchR() + 
  ylab("-log10(P-adj) Motif Enrichment") + 
  xlab("Rank Sorted TFs Enriched") +
  scale_color_gradientn(colors = paletteContinuous(set = "comet"))

ggUp
#This worked when I typed it into the console and hit enter...
pdf("ggUp_plot.pdf", width = 6, height = 4)
print(ggUp)
dev.off()
```
```{r}
top10 <- df[order(-df$mlog10Padj), ][1:10, ]


dotmotifenrich_HMvM <- ggplot(top10, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj)) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  coord_flip() +  # Makes it horizontal
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Marcophage)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )
pdf("dotmotifenrich_HMvM.pdf", width = 6, height = 4)
print(dotmotifenrich_HMvM)
dev.off()
```
# For Hybrid-M vs. Tumor
```{r}
markerTest_HMvT <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Tumor"
)
motifsUp_HMvT <- peakAnnoEnrichment(
    seMarker = markerTest_HMvT,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )
df_HMvT <- data.frame(TF = rownames(motifsUp_HMvT), mlog10Padj = assay(motifsUp_HMvT)[,1])
df_HMvT <- df_HMvT[order(df_HMvT$mlog10Padj, decreasing = TRUE),]
df_HMvT$rank <- seq_len(nrow(df_HMvT))

top10_HMvT <- df_HMvT[order(-df$mlog10Padj), ][1:10, ]


dotmotifenrich_HMvT <- ggplot(top10_HMvT, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj)) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  coord_flip() +  # Makes it horizontal
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Tumor)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )
pdf("dotmotifenrich_HMvT.pdf", width = 6, height = 4)
print(dotmotifenrich_HMvT)
dev.off()
```
# For Hybrid-M vs. Hybrid-T
```{r}
markerTest_HMvHT <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Hybrid-T"
)
motifsUp_HMvHT <- peakAnnoEnrichment(
    seMarker = markerTest_HMvHT,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )
df_HMvHT <- data.frame(TF = rownames(motifsUp_HMvHT), mlog10Padj = assay(motifsUp_HMvHT)[,1])
df_HMvHT <- df_HMvHT[order(df_HMvHT$mlog10Padj, decreasing = TRUE),]
df_HMvHT$rank <- seq_len(nrow(df_HMvHT))

top10_HMvHT <- df_HMvHT[order(-df$mlog10Padj), ][1:10, ]


dotmotifenrich_HMvHT <- ggplot(top10_HMvHT, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj)) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  coord_flip() +  # Makes it horizontal
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Hybrid-T)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )
pdf("dotmotifenrich_HMvHT.pdf", width = 6, height = 4)
print(dotmotifenrich_HMvHT)
dev.off()
```
# For Hybrid-T vs. Macrophage
```{r}
markerTest_HTvM <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-T",
  bgdGroups = "Macrophage"
)
motifsUp_HTvM <- peakAnnoEnrichment(
    seMarker = markerTest_HTvM,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )
df_HTvM <- data.frame(TF = rownames(motifsUp_HTvM), mlog10Padj = assay(motifsUp_HTvM)[,1])
df_HTvM <- df_HTvM[order(df_HTvM$mlog10Padj, decreasing = TRUE),]
df_HTvM$rank <- seq_len(nrow(df_HTvM))

top10_HTvM <- df_HTvM[order(-df$mlog10Padj), ][1:10, ]


dotmotifenrich_HTvM <- ggplot(top10_HTvM, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj)) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  coord_flip() +  # Makes it horizontal
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top Enriched Motifs (Differential Hybrid-T vs. Macrophage)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )
pdf("dotmotifenrich_HTvM.pdf", width = 6, height = 4)
print(dotmotifenrich_HTvM)
dev.off()
```
# For Hybrid-T vs. Tumor
```{r}
markerTest_HTvT <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-T",
  bgdGroups = "Tumor"
)
motifsUp_HTvT <- peakAnnoEnrichment(
    seMarker = markerTest_HTvT,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )
df_HTvT <- data.frame(TF = rownames(motifsUp_HTvT), mlog10Padj = assay(motifsUp_HTvT)[,1])
df_HTvT <- df_HTvT[order(df_HTvT$mlog10Padj, decreasing = TRUE),]
df_HTvT$rank <- seq_len(nrow(df_HTvT))

top10_HTvT <- df_HTvT[order(-df$mlog10Padj), ][1:10, ]


dotmotifenrich_HTvT <- ggplot(top10_HTvT, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj)) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  coord_flip() +  # Makes it horizontal
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top Enriched Motifs (Differential Hybrid-T vs. Tumor)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )
pdf("dotmotifenrich_HTvT.pdf", width = 6, height = 4)
print(dotmotifenrich_HTvT)
dev.off()
```
# For Hybrid-M vs. Macrophage Updated for Only Enrichment
```{r}
markerTest_HMvM <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Macrophage"
)
# Enriched motifs (Log2FC >= 0.5)
motifsUp_HMvM <- peakAnnoEnrichment(
    seMarker = markerTest_HMvM,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )


# Data frame for enriched motifs
df_up_HMvM <- data.frame(
  TF = rownames(motifsUp_HMvM),
  mlog10Padj = assay(motifsUp_HMvM)[,1],
  Direction = "Enriched"
)
# Remove _ and trailing numbers from TF names
df_up_HMvM$TF <- sub("_\\d+$", "", df_up_HMvM$TF)
df_up_HMvM <- df_up_HMvM[order(df_up_HMvM$mlog10Padj, decreasing = TRUE), ]
top10_up <- head(df_up_HMvM, 10)

# Combine top 5 enriched and depleted
top_motifs_HMvM_E <- rbind(top10_up)


# Plot
dotmotifenrich_HMvM_Enriched_Updated <- ggplot(top_motifs_HMvM_E, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj)) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  coord_flip() +  # Makes it horizontal
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Macrophage)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )

# Save plot
pdf("dotmotifenrich_HMvM_Enriched_Updated.pdf", width = 6, height = 4)
print(dotmotifenrich_HMvM_Enriched_Updated)
dev.off()
```
# For Hybrid-M vs. Macrophage Updated for Depletion
```{r}
markerTest_HMvM <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Macrophage"
)
# Enriched motifs (Log2FC >= 0.5)
motifsUp_HMvM <- peakAnnoEnrichment(
    seMarker = markerTest_HMvM,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )
# Depleted motifs (Log2FC <= -0.5)
motifsDown_HMvM <- peakAnnoEnrichment(
  seMarker = markerTest_HMvM,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC <= -0.5"
)

# Data frame for enriched motifs
df_up_HMvM <- data.frame(
  TF = rownames(motifsUp_HMvM),
  mlog10Padj = assay(motifsUp_HMvM)[,1],
  Direction = "Enriched"
)
# Remove _ and trailing numbers from TF names
df_up_HMvM$TF <- sub("_\\d+$", "", df_up_HMvM$TF)
df_up_HMvM <- df_up_HMvM[order(df_up_HMvM$mlog10Padj, decreasing = TRUE), ]
top5_up <- head(df_up_HMvM, 5)

# Data frame for depleted motifs
df_down_HMvM <- data.frame(
  TF = rownames(motifsDown_HMvM),
  mlog10Padj = assay(motifsDown_HMvM)[,1],
  Direction = "Depleted"
)
# Remove _ and trailing numbers from TF names
df_down_HMvM$TF <- sub("_\\d+$", "", df_down_HMvM$TF)
df_down_HMvM <- df_down_HMvM[order(df_down_HMvM$mlog10Padj, decreasing = TRUE), ]
top5_down <- head(df_down_HMvM, 5)

# Combine top 5 enriched and depleted
top_motifs <- rbind(top5_up, top5_down)


# Plot
dotmotifenrich_HMvM_Updated <- ggplot(top_motifs, aes(x = reorder(TF, mlog10Padj), y = mlog10Padj, color = Direction)) +
  geom_point(aes(size = mlog10Padj)) +
  coord_flip() +  # Horizontal plot
  labs(
    x = "Transcription Factor (Motif)",
    y = "-log10(P-adj)",
    title = "Top 5 Enriched and Depleted Motifs (Hybrid-M vs. Macrophage)"
  ) +
  scale_color_manual(values = c("Enriched" = "red", "Depleted" = "blue")) +
  theme_minimal() +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(hjust = 0.5)
  )

# Save plot
pdf("dotmotifenrich_HMvM_Updated.pdf", width = 6, height = 4)
print(dotmotifenrich_HMvM_Updated)
dev.off()
```

## Plotting Motif Logos 
```{r}
#install.packages("ggseqlogo")
```
```{r}
library(ggseqlogo)
logo_runx1 <- getPeakAnnotation(proj_motif5, "Motif")$motifs[["Runx1_712"]]

PWMatrixToProbMatrix <- function(x){
  if (class(x) != "PWMatrix") stop("x must be a TFBSTools::PWMatrix object")
  m <- (exp(as(x, "matrix"))) * TFBSTools::bg(x)/sum(TFBSTools::bg(x))
  m <- t(t(m)/colSums(m))
  m
}
ppm_runx1 <- PWMatrixToProbMatrix(logo_runx1)
colSums(ppm_runx1) %>% range
logo_plot <- ggseqlogo(ppm_runx1, method = "bits")
ggsave("runx1_logo.png", plot = logo_plot, width = 6, height = 3, dpi = 300)

```

```{r}
#install.packages("ggimage")
```
```{r}
# Load required libraries
library(ArchR)
library(ggplot2)
library(ggimage)
library(TFBSTools)
library(grid)
library(dplyr)

# Create MotifMatrix by computing deviations
proj_motif5 <- addDeviationsMatrix(
  ArchRProj = proj_motif5,
  matrixName = "MotifMatrix"
)
```

```{r}
exists("getPwm", envir = asNamespace("ArchR"))

print(top_motifs_HMvM_E)
# Check if logo files exist
sapply(top_motifs_HMvM_E$logo_path, file.exists)


```
# Plotting Hybrid-M vs. Macrophage Enriched Motifs w/ Logos
```{r}
# Your existing code to get enriched motifs
markerTest_HMvM <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Macrophage"
)

# Enriched motifs (Log2FC >= 0.5)
motifsUp_HMvM <- peakAnnoEnrichment(
    seMarker = markerTest_HMvM,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

# Data frame for enriched motifs
df_up_HMvM <- data.frame(
  TF = rownames(motifsUp_HMvM),
  mlog10Padj = assay(motifsUp_HMvM)[,1],
  Direction = "Enriched"
)

# Remove _ and trailing numbers from TF names
df_up_HMvM$TF <- sub("_\\d+$", "", df_up_HMvM$TF)
df_up_HMvM <- df_up_HMvM[order(df_up_HMvM$mlog10Padj, decreasing = TRUE), ]
top10_up <- head(df_up_HMvM, 10)

# Combine top 10 enriched motifs
top_motifs_HMvM_E <- top10_up

# Directory to save motif logos
dir.create("motif_logos", showWarnings = FALSE)

# Function to generate and save motif logos
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  # Get available motifs
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs) || !motif_name %in% motifs) {
    warning(paste("Motif", motif_name, "not found in MotifMatrix"))
    return(NULL)
  }
  
  # Get motif PWM from annotations
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (!motif_name %in% names(motif_annotations)) {
      stop(paste("Motif", motif_name, "not found in annotations"))
    }
    pwm <- motif_annotations[[motif_name]]@profileMatrix
    pwm
  }, error = function(e) {
    warning(paste("PWM not found for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    return(NULL)
  }
  
  # Convert PWM to PFMatrix
  pwm_matrix <- as.matrix(motif_pwm)
  pfm <- PFMatrix(ID = motif_name, name = motif_name, profileMatrix = pwm_matrix)
  
  # Save logo as PNG
  output_file <- file.path(output_dir, paste0(motif_name, ".png"))
  png(output_file, width = 200, height = 100)
  seqLogo(pfm)
  dev.off()
  
  return(output_file)
}

# Generate logos for top 10 motifs and add image paths to data frame
top_motifs_HMvM_E$logo_path <- sapply(top_motifs_HMvM_E$TF, function(tf) {
  generateMotifLogo(tf, proj_motif5, "motif_logos")
})

# Check if any logos failed to generate
top_motifs_HMvM_E <- top_motifs_HMvM_E[!is.null(top_motifs_HMvM_E$logo_path), ]

# Create and save plot
dotmotifenrich_HMvM_Enriched_Updated <- ggplot(top_motifs_HMvM_E, 
                                              aes(x = mlog10Padj, y = reorder(TF, mlog10Padj))) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  geom_image(aes(y = reorder(TF, mlog10Padj), image = logo_path), size = 0.2, nudge_x = -0.5, by = "height") +
  labs(
    x = "-log10(P-adj)",
    y = "Transcription Factor (Motif)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Macrophage)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10), # Show motif names
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_text(size = 12)
  )

# Save plot
pdf("250515_dotmotifenrich_HMvM_Enriched_WithLogos.pdf", width = 8, height = 6)
print(dotmotifenrich_HMvM_Enriched_Updated)
dev.off()
```

# Trouble Shoot Motif Annotations
```{r}
# Load required libraries
library(ArchR)
library(ggplot2)
library(ggimage)
library(TFBSTools)
library(grid)
library(dplyr)

# Re-add motif annotations with cisbp
proj_motif5 <- addMotifAnnotations(
  ArchRProj = proj_motif5,
  motifSet = "cisbp",
  name = "Motif",
  force = TRUE # Overwrite existing annotations
)

# Re-generate MotifMatrix
proj_motif5 <- addDeviationsMatrix(
  ArchRProj = proj_motif5,
  matrixName = "MotifMatrix",
  force = TRUE # Overwrite existing MotifMatrix
)

motifs <- getFeatures(proj_motif5, useMatrix = "MotifMatrix")
print(head(motifs))
grep("Runx", motifs, value = TRUE)
```
# Is this the code that actually worked??!?!
```{r}
# Load required libraries
library(ArchR)
library(ggplot2)
library(ggimage)
library(ggseqlogo)
library(TFBSTools)
library(grid)
library(dplyr)

# Ensure MotifMatrix exists
if (!"MotifMatrix" %in% getAvailableMatrices(proj_motif5)) {
  cat("Adding motif annotations and generating MotifMatrix...\n")
  proj_motif5 <- addMotifAnnotations(
    ArchRProj = proj_motif5,
    motifSet = "cisbp",
    name = "Motif",
    force = TRUE
  )
  proj_motif5 <- addDeviationsMatrix(
    ArchRProj = proj_motif5,
    matrixName = "MotifMatrix",
    force = TRUE
  )
}

# Verify MotifMatrix motifs
cat("Available motifs in MotifMatrix:\n")
motifs <- getFeatures(proj_motif5, useMatrix = "MotifMatrix")
print(head(motifs))
print(grep("Runx", motifs, value = TRUE))

# Your existing code for marker features and motif enrichment
markerTest_HMvM <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Macrophage"
)

motifsUp_HMvM <- peakAnnoEnrichment(
  seMarker = markerTest_HMvM,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

# Verify enriched motif names
cat("Enriched motifs from peakAnnoEnrichment:\n")
print(head(rownames(motifsUp_HMvM)))

# Data frame for enriched motifs
df_up_HMvM <- data.frame(
  TF = rownames(motifsUp_HMvM),
  mlog10Padj = assay(motifsUp_HMvM)[,1],
  Direction = "Enriched"
)

# Sort and select top 10
df_up_HMvM <- df_up_HMvM[order(df_up_HMvM$mlog10Padj, decreasing = TRUE), ]
top10_up <- head(df_up_HMvM, 10)
top_motifs_HMvM_E <- top10_up

# Verify motif names
cat("Top enriched motifs:\n")
print(top_motifs_HMvM_E$TF)

# Directory for motif logos
dir.create("motif_logos", showWarnings = FALSE)

# Function to convert PWMatrix to probability matrix
PWMatrixToProbMatrix <- function(x){
  if (class(x) != "PWMatrix") stop("x must be a TFBSTools::PWMatrix object")
  m <- (exp(as(x, "matrix"))) * TFBSTools::bg(x)/sum(TFBSTools::bg(x))
  m <- t(t(m)/colSums(m))
  m
}

# Function to generate and save motif logos using ggseqlogo
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  cat("Attempting to process motif:", motif_name, "\n")
  
  # Get available motifs from MotifMatrix
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs)) {
    warning("No motifs found in MotifMatrix")
    return(NULL)
  }
  
  # Check for motif with deviations: or z: prefix
  motif_in_matrix <- motif_name
  if (!motif_name %in% motifs) {
    prefixed_name <- paste0("deviations:", motif_name)
    if (prefixed_name %in% motifs) {
      cat("Found motif with deviations: prefix:", prefixed_name, "\n")
      motif_in_matrix <- prefixed_name
    } else {
      prefixed_name <- paste0("z:", motif_name)
      if (prefixed_name %in% motifs) {
        cat("Found motif with z: prefix:", prefixed_name, "\n")
        motif_in_matrix <- prefixed_name
      } else {
        warning(paste("Motif", motif_name, "not found in MotifMatrix. Available motifs:", paste(head(motifs), collapse = ", ")))
        return(NULL)
      }
    }
  }
  
  # Get motif PWMatrix from annotations (using raw motif name)
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (is.null(motif_annotations)) {
      warning("Motif annotations not found")
      return(NULL)
    }
    if (!motif_name %in% names(motif_annotations)) {
      warning(paste("Motif", motif_name, "not found in annotations. Available annotations:", paste(head(names(motif_annotations)), collapse = ", ")))
      return(NULL)
    }
    pwm <- motif_annotations[[motif_name]] # Get PWMatrix object
    pwm
  }, error = function(e) {
    warning(paste("Failed to retrieve PWMatrix for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    warning(paste("No PWMatrix retrieved for", motif_name))
    return(NULL)
  }
  
  cat("PWMatrix retrieved for", motif_name, "\n")
  
  # Convert PWMatrix to probability matrix
  tryCatch({
    ppm <- PWMatrixToProbMatrix(motif_pwm)
    cat("Probability matrix created for", motif_name, "\n")
    
    # Save logo as PNG using ggseqlogo
    output_file <- file.path(output_dir, paste0(motif_name, "_logo.png"))
    cat("Attempting to save logo to", output_file, "\n")
    
    logo_plot <- ggseqlogo(ppm, method = "bits")
    ggsave(output_file, plot = logo_plot, width = 6, height = 3, dpi = 300)
    
    if (file.exists(output_file)) {
      cat("Logo successfully created for", motif_name, "at", output_file, "\n")
      return(output_file)
    } else {
      warning("Failed to create logo file for", motif_name)
      return(NULL)
    }
  }, error = function(e) {
    warning(paste("Failed to save logo for", motif_name, ":", e$message))
    return(NULL)
  })
}

# Generate logos using original motif names
top_motifs_HMvM_E$logo_path <- sapply(top_motifs_HMvM_E$TF, function(tf) {
  generateMotifLogo(tf, proj_motif5, "motif_logos")
})

# Trim motif names for display *after* logo generation
top_motifs_HMvM_E$TF_display <- sub("_\\d+$", "", top_motifs_HMvM_E$TF)

# Check logo paths and existence
print("Logo paths:")
print(top_motifs_HMvM_E$logo_path)
print("Files exist:")
logo_exists <- sapply(top_motifs_HMvM_E$logo_path, function(x) is.character(x) && nzchar(x) && file.exists(x))
print(logo_exists)
cat("Generated logo files:\n")
print(list.files("motif_logos", pattern = "*.png", full.names = TRUE))

# Create and save plot
dotmotifenrich_HMvM_Enriched_Updated <- ggplot(top_motifs_HMvM_E, 
                                              aes(x = mlog10Padj, y = reorder(TF_display, mlog10Padj))) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  { if (any(logo_exists)) 
      geom_image(aes(x = 1.5, y = reorder(TF_display, mlog10Padj), image = logo_path), 
                 size = 0.1, nudge_y = 0.2, by = "height") } +
  scale_x_continuous(limits = c(1, max(top_motifs_HMvM_E$mlog10Padj) * 1.1)) +
  labs(
    x = "-log10(P-adj)",
    y = "Transcription Factor (Motif)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Macrophage)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_text(size = 12)
  )

# Save plot
pdf("dotmotifenrich_HMvM_Enriched_WithLogos.pdf", width = 8, height = 6)
print(dotmotifenrich_HMvM_Enriched_Updated)
dev.off()
```
# Plotting Hybrid-M vs. Tumor Enriched Motifs w/ Logos
```{r}
# Load required libraries
library(ArchR)
library(ggplot2)
library(ggimage)
library(ggseqlogo)
library(TFBSTools)
library(grid)
library(dplyr)

# Ensure MotifMatrix exists
if (!"MotifMatrix" %in% getAvailableMatrices(proj_motif5)) {
  cat("Adding motif annotations and generating MotifMatrix...\n")
  proj_motif5 <- addMotifAnnotations(
    ArchRProj = proj_motif5,
    motifSet = "cisbp",
    name = "Motif",
    force = TRUE
  )
  proj_motif5 <- addDeviationsMatrix(
    ArchRProj = proj_motif5,
    matrixName = "MotifMatrix",
    force = TRUE
  )
}

# Verify MotifMatrix motifs
cat("Available motifs in MotifMatrix:\n")
motifs <- getFeatures(proj_motif5, useMatrix = "MotifMatrix")
print(head(motifs))
print(grep("Runx", motifs, value = TRUE))

# Your existing code for marker features and motif enrichment
markerTest_HMvT <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M",
  bgdGroups = "Tumor"
)

motifsUp_HMvT <- peakAnnoEnrichment(
  seMarker = markerTest_HMvT,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

# Verify enriched motif names
cat("Enriched motifs from peakAnnoEnrichment:\n")
print(head(rownames(motifsUp_HMvT)))

# Data frame for enriched motifs
df_up_HMvT <- data.frame(
  TF = rownames(motifsUp_HMvT),
  mlog10Padj = assay(motifsUp_HMvT)[,1],
  Direction = "Enriched"
)

# Sort and select top 10
df_up_HMvT <- df_up_HMvT[order(df_up_HMvT$mlog10Padj, decreasing = TRUE), ]
top10_up_HMvT <- head(df_up_HMvT, 10)
top_motifs_HMvT_E <- top10_up_HMvT

# Verify motif names
cat("Top enriched motifs:\n")
print(top_motifs_HMvT_E$TF)

# Directory for motif logos
dir.create("motif_logos", showWarnings = FALSE)

# Function to convert PWMatrix to probability matrix
PWMatrixToProbMatrix <- function(x){
  if (class(x) != "PWMatrix") stop("x must be a TFBSTools::PWMatrix object")
  m <- (exp(as(x, "matrix"))) * TFBSTools::bg(x)/sum(TFBSTools::bg(x))
  m <- t(t(m)/colSums(m))
  m
}

# Function to generate and save motif logos using ggseqlogo
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  cat("Attempting to process motif:", motif_name, "\n")
  
  # Get available motifs from MotifMatrix
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs)) {
    warning("No motifs found in MotifMatrix")
    return(NULL)
  }
  
  # Check for motif with deviations: or z: prefix
  motif_in_matrix <- motif_name
  if (!motif_name %in% motifs) {
    prefixed_name <- paste0("deviations:", motif_name)
    if (prefixed_name %in% motifs) {
      cat("Found motif with deviations: prefix:", prefixed_name, "\n")
      motif_in_matrix <- prefixed_name
    } else {
      prefixed_name <- paste0("z:", motif_name)
      if (prefixed_name %in% motifs) {
        cat("Found motif with z: prefix:", prefixed_name, "\n")
        motif_in_matrix <- prefixed_name
      } else {
        warning(paste("Motif", motif_name, "not found in MotifMatrix. Available motifs:", paste(head(motifs), collapse = ", ")))
        return(NULL)
      }
    }
  }
  
  # Get motif PWMatrix from annotations (using raw motif name)
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (is.null(motif_annotations)) {
      warning("Motif annotations not found")
      return(NULL)
    }
    if (!motif_name %in% names(motif_annotations)) {
      warning(paste("Motif", motif_name, "not found in annotations. Available annotations:", paste(head(names(motif_annotations)), collapse = ", ")))
      return(NULL)
    }
    pwm <- motif_annotations[[motif_name]] # Get PWMatrix object
    pwm
  }, error = function(e) {
    warning(paste("Failed to retrieve PWMatrix for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    warning(paste("No PWMatrix retrieved for", motif_name))
    return(NULL)
  }
  
  cat("PWMatrix retrieved for", motif_name, "\n")
  
  # Convert PWMatrix to probability matrix
  tryCatch({
    ppm <- PWMatrixToProbMatrix(motif_pwm)
    cat("Probability matrix created for", motif_name, "\n")
    
    # Save logo as PNG using ggseqlogo
    output_file <- file.path(output_dir, paste0(motif_name, "_logo.png"))
    cat("Attempting to save logo to", output_file, "\n")
    
    logo_plot <- ggseqlogo(ppm, method = "bits")
    ggsave(output_file, plot = logo_plot, width = 6, height = 3, dpi = 300)
    
    if (file.exists(output_file)) {
      cat("Logo successfully created for", motif_name, "at", output_file, "\n")
      return(output_file)
    } else {
      warning("Failed to create logo file for", motif_name)
      return(NULL)
    }
  }, error = function(e) {
    warning(paste("Failed to save logo for", motif_name, ":", e$message))
    return(NULL)
  })
}

# Generate logos using original motif names
top_motifs_HMvT_E$logo_path <- sapply(top_motifs_HMvT_E$TF, function(tf) {
  generateMotifLogo(tf, proj_motif5, "motif_logos")
})

# Trim motif names for display *after* logo generation
top_motifs_HMvT_E$TF_display <- sub("_\\d+$", "", top_motifs_HMvT_E$TF)

# Check logo paths and existence
print("Logo paths:")
print(top_motifs_HMvT_E$logo_path)
print("Files exist:")
logo_exists <- sapply(top_motifs_HMvT_E$logo_path, function(x) is.character(x) && nzchar(x) && file.exists(x))
print(logo_exists)
cat("Generated logo files:\n")
print(list.files("motif_logos", pattern = "*.png", full.names = TRUE))

# Create and save plot
dotmotifenrich_HMvT_Enriched_Updated <- ggplot(top_motifs_HMvT_E, 
                                              aes(x = mlog10Padj, y = reorder(TF_display, mlog10Padj))) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  { if (any(logo_exists)) 
      geom_image(aes(x = 100, y = reorder(TF_display, mlog10Padj), image = logo_path), 
                 size = 0.1, nudge_y = 0.2, by = "height") } +
  scale_x_continuous(limits = c(1, max(top_motifs_HMvT_E$mlog10Padj) * 1.1)) +
  labs(
    x = "-log10(P-adj)",
    y = "Transcription Factor (Motif)",
    title = "Top Enriched Motifs (Differential Hybrid-M vs. Tumor)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_text(size = 12)
  )

# Save plot
pdf("dotmotifenrich_HMvT_Enriched_WithLogos.pdf", width = 8, height = 6)
print(dotmotifenrich_HMvT_Enriched_Updated)
dev.off()
```
# Plotting Hybrid-T vs. Tumor Enriched Motifs w/ Logos
```{r}
# Load required libraries
library(ArchR)
library(ggplot2)
library(ggimage)
library(ggseqlogo)
library(TFBSTools)
library(grid)
library(dplyr)

# Ensure MotifMatrix exists
if (!"MotifMatrix" %in% getAvailableMatrices(proj_motif5)) {
  cat("Adding motif annotations and generating MotifMatrix...\n")
  proj_motif5 <- addMotifAnnotations(
    ArchRProj = proj_motif5,
    motifSet = "cisbp",
    name = "Motif",
    force = TRUE
  )
  proj_motif5 <- addDeviationsMatrix(
    ArchRProj = proj_motif5,
    matrixName = "MotifMatrix",
    force = TRUE
  )
}

# Verify MotifMatrix motifs
cat("Available motifs in MotifMatrix:\n")
motifs <- getFeatures(proj_motif5, useMatrix = "MotifMatrix")
print(head(motifs))
print(grep("Runx", motifs, value = TRUE))

# Your existing code for marker features and motif enrichment
markerTest_HTvT <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-T",
  bgdGroups = "Tumor"
)

motifsUp_HTvT <- peakAnnoEnrichment(
  seMarker = markerTest_HTvT,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

# Verify enriched motif names
cat("Enriched motifs from peakAnnoEnrichment:\n")
print(head(rownames(motifsUp_HTvT)))

# Data frame for enriched motifs
df_up_HTvT <- data.frame(
  TF = rownames(motifsUp_HTvT),
  mlog10Padj = assay(motifsUp_HTvT)[,1],
  Direction = "Enriched"
)

# Sort and select top 10
df_up_HTvT <- df_up_HTvT[order(df_up_HTvT$mlog10Padj, decreasing = TRUE), ]
top10_up_HTvT <- head(df_up_HTvT, 10)
top_motifs_HTvT_E <- top10_up_HTvT

# Verify motif names
cat("Top enriched motifs:\n")
print(top_motifs_HTvT_E$TF)

# Directory for motif logos
dir.create("motif_logos", showWarnings = FALSE)

# Function to convert PWMatrix to probability matrix
PWMatrixToProbMatrix <- function(x){
  if (class(x) != "PWMatrix") stop("x must be a TFBSTools::PWMatrix object")
  m <- (exp(as(x, "matrix"))) * TFBSTools::bg(x)/sum(TFBSTools::bg(x))
  m <- t(t(m)/colSums(m))
  m
}

# Function to generate and save motif logos using ggseqlogo
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  cat("Attempting to process motif:", motif_name, "\n")
  
  # Get available motifs from MotifMatrix
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs)) {
    warning("No motifs found in MotifMatrix")
    return(NULL)
  }
  
  # Check for motif with deviations: or z: prefix
  motif_in_matrix <- motif_name
  if (!motif_name %in% motifs) {
    prefixed_name <- paste0("deviations:", motif_name)
    if (prefixed_name %in% motifs) {
      cat("Found motif with deviations: prefix:", prefixed_name, "\n")
      motif_in_matrix <- prefixed_name
    } else {
      prefixed_name <- paste0("z:", motif_name)
      if (prefixed_name %in% motifs) {
        cat("Found motif with z: prefix:", prefixed_name, "\n")
        motif_in_matrix <- prefixed_name
      } else {
        warning(paste("Motif", motif_name, "not found in MotifMatrix. Available motifs:", paste(head(motifs), collapse = ", ")))
        return(NULL)
      }
    }
  }
  
  # Get motif PWMatrix from annotations (using raw motif name)
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (is.null(motif_annotations)) {
      warning("Motif annotations not found")
      return(NULL)
    }
    if (!motif_name %in% names(motif_annotations)) {
      warning(paste("Motif", motif_name, "not found in annotations. Available annotations:", paste(head(names(motif_annotations)), collapse = ", ")))
      return(NULL)
    }
    pwm <- motif_annotations[[motif_name]] # Get PWMatrix object
    pwm
  }, error = function(e) {
    warning(paste("Failed to retrieve PWMatrix for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    warning(paste("No PWMatrix retrieved for", motif_name))
    return(NULL)
  }
  
  cat("PWMatrix retrieved for", motif_name, "\n")
  
  # Convert PWMatrix to probability matrix
  tryCatch({
    ppm <- PWMatrixToProbMatrix(motif_pwm)
    cat("Probability matrix created for", motif_name, "\n")
    
    # Save logo as PNG using ggseqlogo
    output_file <- file.path(output_dir, paste0(motif_name, "_logo.png"))
    cat("Attempting to save logo to", output_file, "\n")
    
    logo_plot <- ggseqlogo(ppm, method = "bits")
    ggsave(output_file, plot = logo_plot, width = 6, height = 3, dpi = 300)
    
    if (file.exists(output_file)) {
      cat("Logo successfully created for", motif_name, "at", output_file, "\n")
      return(output_file)
    } else {
      warning("Failed to create logo file for", motif_name)
      return(NULL)
    }
  }, error = function(e) {
    warning(paste("Failed to save logo for", motif_name, ":", e$message))
    return(NULL)
  })
}

# Generate logos using original motif names
top_motifs_HTvT_E$logo_path <- sapply(top_motifs_HTvT_E$TF, function(tf) {
  generateMotifLogo(tf, proj_motif5, "motif_logos")
})

# Trim motif names for display *after* logo generation
top_motifs_HTvT_E$TF_display <- sub("_\\d+$", "", top_motifs_HTvT_E$TF)

# Check logo paths and existence
print("Logo paths:")
print(top_motifs_HTvT_E$logo_path)
print("Files exist:")
logo_exists <- sapply(top_motifs_HTvT_E$logo_path, function(x) is.character(x) && nzchar(x) && file.exists(x))
print(logo_exists)
cat("Generated logo files:\n")
print(list.files("motif_logos", pattern = "*.png", full.names = TRUE))

# Create and save plot
dotmotifenrich_HTvT_Enriched_Updated <- ggplot(top_motifs_HTvT_E, 
                                              aes(x = mlog10Padj, y = reorder(TF_display, mlog10Padj))) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  { if (any(logo_exists)) 
      geom_image(aes(x = 25, y = reorder(TF_display, mlog10Padj), image = logo_path), 
                 size = 0.1, nudge_y = 0.2, by = "height") } +
  scale_x_continuous(limits = c(1, max(top_motifs_HTvT_E$mlog10Padj) * 1.1)) +
  labs(
    x = "-log10(P-adj)",
    y = "Transcription Factor (Motif)",
    title = "Top Enriched Motifs (Differential Hybrid-T vs. Tumor)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_text(size = 12)
  )

# Save plot
pdf("dotmotifenrich_HTvT_Enriched_WithLogos.pdf", width = 8, height = 6)
print(dotmotifenrich_HTvT_Enriched_Updated)
dev.off()
```
# Plotting Hybrid-T vs. Macrophage Enriched Motifs w/ Logos
```{r}
# Load required libraries
library(ArchR)
library(ggplot2)
library(ggimage)
library(ggseqlogo)
library(TFBSTools)
library(grid)
library(dplyr)

# Ensure MotifMatrix exists
if (!"MotifMatrix" %in% getAvailableMatrices(proj_motif5)) {
  cat("Adding motif annotations and generating MotifMatrix...\n")
  proj_motif5 <- addMotifAnnotations(
    ArchRProj = proj_motif5,
    motifSet = "cisbp",
    name = "Motif",
    force = TRUE
  )
  proj_motif5 <- addDeviationsMatrix(
    ArchRProj = proj_motif5,
    matrixName = "MotifMatrix",
    force = TRUE
  )
}

# Verify MotifMatrix motifs
cat("Available motifs in MotifMatrix:\n")
motifs <- getFeatures(proj_motif5, useMatrix = "MotifMatrix")
print(head(motifs))
print(grep("Runx", motifs, value = TRUE))

# Your existing code for marker features and motif enrichment
markerTest_HTvM <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-T",
  bgdGroups = "Macrophage"
)

motifsUp_HTvM <- peakAnnoEnrichment(
  seMarker = markerTest_HTvM,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

# Verify enriched motif names
cat("Enriched motifs from peakAnnoEnrichment:\n")
print(head(rownames(motifsUp_HTvM)))

# Data frame for enriched motifs
df_up_HTvM <- data.frame(
  TF = rownames(motifsUp_HTvM),
  mlog10Padj = assay(motifsUp_HTvM)[,1],
  Direction = "Enriched"
)

# Sort and select top 10
df_up_HTvM <- df_up_HTvM[order(df_up_HTvM$mlog10Padj, decreasing = TRUE), ]
top10_up_HTvM <- head(df_up_HTvM, 10)
top_motifs_HTvM_E <- top10_up_HTvM

# Verify motif names
cat("Top enriched motifs:\n")
print(top_motifs_HTvM_E$TF)

# Directory for motif logos
dir.create("motif_logos", showWarnings = FALSE)

# Function to convert PWMatrix to probability matrix
PWMatrixToProbMatrix <- function(x){
  if (class(x) != "PWMatrix") stop("x must be a TFBSTools::PWMatrix object")
  m <- (exp(as(x, "matrix"))) * TFBSTools::bg(x)/sum(TFBSTools::bg(x))
  m <- t(t(m)/colSums(m))
  m
}

# Function to generate and save motif logos using ggseqlogo
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  cat("Attempting to process motif:", motif_name, "\n")
  
  # Get available motifs from MotifMatrix
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs)) {
    warning("No motifs found in MotifMatrix")
    return(NULL)
  }
  
  # Check for motif with deviations: or z: prefix
  motif_in_matrix <- motif_name
  if (!motif_name %in% motifs) {
    prefixed_name <- paste0("deviations:", motif_name)
    if (prefixed_name %in% motifs) {
      cat("Found motif with deviations: prefix:", prefixed_name, "\n")
      motif_in_matrix <- prefixed_name
    } else {
      prefixed_name <- paste0("z:", motif_name)
      if (prefixed_name %in% motifs) {
        cat("Found motif with z: prefix:", prefixed_name, "\n")
        motif_in_matrix <- prefixed_name
      } else {
        warning(paste("Motif", motif_name, "not found in MotifMatrix. Available motifs:", paste(head(motifs), collapse = ", ")))
        return(NULL)
      }
    }
  }
  
  # Get motif PWMatrix from annotations (using raw motif name)
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (is.null(motif_annotations)) {
      warning("Motif annotations not found")
      return(NULL)
    }
    if (!motif_name %in% names(motif_annotations)) {
      warning(paste("Motif", motif_name, "not found in annotations. Available annotations:", paste(head(names(motif_annotations)), collapse = ", ")))
      return(NULL)
    }
    pwm <- motif_annotations[[motif_name]] # Get PWMatrix object
    pwm
  }, error = function(e) {
    warning(paste("Failed to retrieve PWMatrix for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    warning(paste("No PWMatrix retrieved for", motif_name))
    return(NULL)
  }
  
  cat("PWMatrix retrieved for", motif_name, "\n")
  
  # Convert PWMatrix to probability matrix
  tryCatch({
    ppm <- PWMatrixToProbMatrix(motif_pwm)
    cat("Probability matrix created for", motif_name, "\n")
    
    # Save logo as PNG using ggseqlogo
    output_file <- file.path(output_dir, paste0(motif_name, "_logo.png"))
    cat("Attempting to save logo to", output_file, "\n")
    
    logo_plot <- ggseqlogo(ppm, method = "bits")
    ggsave(output_file, plot = logo_plot, width = 6, height = 3, dpi = 300)
    
    if (file.exists(output_file)) {
      cat("Logo successfully created for", motif_name, "at", output_file, "\n")
      return(output_file)
    } else {
      warning("Failed to create logo file for", motif_name)
      return(NULL)
    }
  }, error = function(e) {
    warning(paste("Failed to save logo for", motif_name, ":", e$message))
    return(NULL)
  })
}

# Generate logos using original motif names
top_motifs_HTvM_E$logo_path <- sapply(top_motifs_HTvM_E$TF, function(tf) {
  generateMotifLogo(tf, proj_motif5, "motif_logos")
})

# Trim motif names for display *after* logo generation
top_motifs_HTvM_E$TF_display <- sub("_\\d+$", "", top_motifs_HTvM_E$TF)

# Check logo paths and existence
print("Logo paths:")
print(top_motifs_HTvM_E$logo_path)
print("Files exist:")
logo_exists <- sapply(top_motifs_HTvM_E$logo_path, function(x) is.character(x) && nzchar(x) && file.exists(x))
print(logo_exists)
cat("Generated logo files:\n")
print(list.files("motif_logos", pattern = "*.png", full.names = TRUE))

# Create and save plot
dotmotifenrich_HTvM_Enriched_Updated <- ggplot(top_motifs_HTvM_E, 
                                              aes(x = mlog10Padj, y = reorder(TF_display, mlog10Padj))) +
  geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
  { if (any(logo_exists)) 
      geom_image(aes(x = 25, y = reorder(TF_display, mlog10Padj), image = logo_path), 
                 size = 0.1, nudge_y = 0.2, by = "height") } +
  scale_x_continuous(limits = c(1, max(top_motifs_HTvM_E$mlog10Padj) * 1.1)) +
  labs(
    x = "-log10(P-adj)",
    y = "Transcription Factor (Motif)",
    title = "Top Enriched Motifs (Differential Hybrid-T vs. Macrophage)"
  ) +
  scale_color_gradientn(colors = c("blue", "purple", "red")) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(hjust = 0.5),
    axis.title.y = element_text(size = 12)
  )

# Save plot
pdf("dotmotifenrich_HTvM_Enriched_WithLogos.pdf", width = 8, height = 6)
print(dotmotifenrich_HTvM_Enriched_Updated)
dev.off()
```

# 250521: Combined plots for Hybrids-M/T vs. All
```{r}
# Hybrid-M vs all
markerTest_HM_vsAll <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-M"
)

# Hybrid-T vs all
markerTest_HT_vsAll <- getMarkerFeatures(
  ArchRProj = proj_motif5, 
  useMatrix = "PeakMatrix",
  groupBy = "Clusters_TuMerged",
  testMethod = "wilcoxon",
  bias = c("TSSEnrichment", "log10(nFrags)"),
  useGroups = "Hybrid-T"
)

# Save Annotated Enrichments
motifsUp_HM <- peakAnnoEnrichment(
  seMarker = markerTest_HM_vsAll,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

motifsUp_HT <- peakAnnoEnrichment(
  seMarker = markerTest_HT_vsAll,
  ArchRProj = proj_motif5,
  peakAnnotation = "Motif",
  cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
)

# Prepare data frames of top 10 motifs for each
getTopMotifsDF <- function(motif_enrichment, label) {
  df <- data.frame(
    TF = rownames(motif_enrichment),
    mlog10Padj = assay(motif_enrichment)[,1],
    Direction = label
  )
  df <- df[order(df$mlog10Padj, decreasing = TRUE), ]
  head(df, 10)
}

top_HM <- getTopMotifsDF(motifsUp_HM, "Hybrid-M")
top_HT <- getTopMotifsDF(motifsUp_HT, "Hybrid-T")

# Function to generate and save motif logos using ggseqlogo
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  cat("Attempting to process motif:", motif_name, "\n")
  
  # Get available motifs from MotifMatrix
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs)) {
    warning("No motifs found in MotifMatrix")
    return(NULL)
  }
  
  # Check for motif with deviations: or z: prefix
  motif_in_matrix <- motif_name
  if (!motif_name %in% motifs) {
    prefixed_name <- paste0("deviations:", motif_name)
    if (prefixed_name %in% motifs) {
      cat("Found motif with deviations: prefix:", prefixed_name, "\n")
      motif_in_matrix <- prefixed_name
    } else {
      prefixed_name <- paste0("z:", motif_name)
      if (prefixed_name %in% motifs) {
        cat("Found motif with z: prefix:", prefixed_name, "\n")
        motif_in_matrix <- prefixed_name
      } else {
        warning(paste("Motif", motif_name, "not found in MotifMatrix. Available motifs:", paste(head(motifs), collapse = ", ")))
        return(NULL)
      }
    }
  }
  
  # Get motif PWMatrix from annotations (using raw motif name)
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (is.null(motif_annotations)) {
      warning("Motif annotations not found")
      return(NULL)
    }
    if (!motif_name %in% names(motif_annotations)) {
      warning(paste("Motif", motif_name, "not found in annotations. Available annotations:", paste(head(names(motif_annotations)), collapse = ", ")))
      return(NULL)
    }
    pwm <- motif_annotations[[motif_name]] # Get PWMatrix object
    pwm
  }, error = function(e) {
    warning(paste("Failed to retrieve PWMatrix for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    warning(paste("No PWMatrix retrieved for", motif_name))
    return(NULL)
  }
  
  cat("PWMatrix retrieved for", motif_name, "\n")
  
  # Convert PWMatrix to probability matrix
  tryCatch({
    ppm <- PWMatrixToProbMatrix(motif_pwm)
    cat("Probability matrix created for", motif_name, "\n")
    
    # Save logo as PNG using ggseqlogo
    output_file <- file.path(output_dir, paste0(motif_name, "_logo.png"))
    cat("Attempting to save logo to", output_file, "\n")
    
    logo_plot <- ggseqlogo(ppm, method = "bits")
    ggsave(output_file, plot = logo_plot, width = 6, height = 3, dpi = 300)
    
    if (file.exists(output_file)) {
      cat("Logo successfully created for", motif_name, "at", output_file, "\n")
      return(output_file)
    } else {
      warning("Failed to create logo file for", motif_name)
      return(NULL)
    }
  }, error = function(e) {
    warning(paste("Failed to save logo for", motif_name, ":", e$message))
    return(NULL)
  })
}


#Apply existing Motif Logo function
top_HM$logo_path <- sapply(top_HM$TF, function(tf) generateMotifLogo(tf, proj_motif5, "motif_logos"))
top_HT$logo_path <- sapply(top_HT$TF, function(tf) generateMotifLogo(tf, proj_motif5, "motif_logos"))

top_HM$TF_display <- sub("_\\d+$", "", top_HM$TF)
top_HT$TF_display <- sub("_\\d+$", "", top_HT$TF)

#Plot it
plotMotifEnrichment <- function(df, title, filename) {
  logo_exists <- sapply(df$logo_path, function(x) is.character(x) && nzchar(x) && file.exists(x))
  
  p <- ggplot(df, aes(x = mlog10Padj, y = reorder(TF_display, mlog10Padj))) +
    geom_point(aes(size = mlog10Padj, color = mlog10Padj)) +
    { if (any(logo_exists)) 
        geom_image(aes(x = 1.5, y = reorder(TF_display, mlog10Padj), image = logo_path), 
                   size = 0.1, nudge_y = 0.2, by = "height") } +
    scale_x_continuous(limits = c(1, max(df$mlog10Padj) * 1.1)) +
    labs(
      x = "-log10(P-adj)",
      y = "Transcription Factor (Motif)",
      title = title
    ) +
    scale_color_gradientn(colors = c("blue", "purple", "red")) +
    theme_minimal() +
    theme(
      axis.text.y = element_text(size = 10),
      axis.text.x = element_text(size = 10),
      plot.title = element_text(hjust = 0.5),
      axis.title.y = element_text(size = 12)
    )
  
  pdf(filename, width = 8, height = 6)
  print(p)
  dev.off()
}
#Generate Plots
plotMotifEnrichment(top_HM, "Top Enriched Motifs: Hybrid-M vs All", "HybridM_vs_All_MotifEnrichment.pdf")
plotMotifEnrichment(top_HT, "Top Enriched Motifs: Hybrid-T vs All", "HybridT_vs_All_MotifEnrichment.pdf")


```




# Generating Only Logos
```{r}
# Load required libraries
library(ArchR)
library(TFBSTools)
library(grid)
library(dplyr)

# Directory for motif logos
dir.create("motif_logos", showWarnings = FALSE)

# Function to generate and save motif logos
generateMotifLogo <- function(motif_name, archr_project, output_dir) {
  cat("Attempting to process motif:", motif_name, "\n")
  
  # Get available motifs from MotifMatrix
  motifs <- tryCatch({
    getFeatures(archr_project, useMatrix = "MotifMatrix")
  }, error = function(e) {
    warning(paste("Failed to retrieve MotifMatrix features:", e$message))
    return(NULL)
  })
  
  if (is.null(motifs)) {
    warning("No motifs found in MotifMatrix")
    return(NULL)
  }
  
  # Check for motif with deviations: or z: prefix
  motif_in_matrix <- motif_name
  if (!motif_name %in% motifs) {
    prefixed_name <- paste0("deviations:", motif_name)
    if (prefixed_name %in% motifs) {
      cat("Found motif with deviations: prefix:", prefixed_name, "\n")
      motif_in_matrix <- prefixed_name
    } else {
      prefixed_name <- paste0("z:", motif_name)
      if (prefixed_name %in% motifs) {
        cat("Found motif with z: prefix:", prefixed_name, "\n")
        motif_in_matrix <- prefixed_name
      } else {
        warning(paste("Motif", motif_name, "not found in MotifMatrix. Available motifs:", paste(head(motifs), collapse = ", ")))
        return(NULL)
      }
    }
  }
  
  # Get motif PWM from annotations (using raw motif name)
  motif_pwm <- tryCatch({
    motif_annotations <- getPeakAnnotation(archr_project, "Motif")$motifs
    if (is.null(motif_annotations)) {
      warning("Motif annotations not found")
      return(NULL)
    }
    if (!motif_name %in% names(motif_annotations)) {
      warning(paste("Motif", motif_name, "not found in annotations. Available annotations:", paste(head(names(motif_annotations)), collapse = ", ")))
      return(NULL)
    }
    pwm <- motif_annotations[[motif_name]]@profileMatrix
    pwm
  }, error = function(e) {
    warning(paste("Failed to retrieve PWM for", motif_name, ":", e$message))
    return(NULL)
  })
  
  if (is.null(motif_pwm)) {
    warning(paste("No PWM retrieved for", motif_name))
    return(NULL)
  }
  
  cat("PWM retrieved for", motif_name, "\n")
  
  # Ensure PWM is a matrix
  pwm_matrix <- as.matrix(motif_pwm)
  
  # Save logo as PNG
  output_file <- file.path(output_dir, paste0(motif_name, "_logo.png"))
  cat("Attempting to save logo to", output_file, "\n")
  
  tryCatch({
    png(output_file, width = 400, height = 200)
    TFBSTools::seqLogo(pwm_matrix) # Explicitly use TFBSTools::seqLogo with raw matrix
    dev.off()
    
    if (file.exists(output_file)) {
      cat("Logo successfully created for", motif_name, "at", output_file, "\n")
      return(output_file)
    } else {
      warning("Failed to create logo file for", motif_name)
      return(NULL)
    }
  }, error = function(e) {
    warning(paste("Failed to save logo for", motif_name, ":", e$message))
    return(NULL)
  })
}

# Generate logos for top 10 enriched motifs
cat("Generating logos for top 10 enriched motifs...\n")
logo_paths <- sapply(top_motifs_HMvM_E$TF, function(tf) {
  generateMotifLogo(tf, proj_motif5, "motif_logos")
})

# Create a data frame to summarize results
logo_results <- data.frame(
  TF = top_motifs_HMvM_E$TF,
  Logo_Path = unlist(logo_paths),
  Logo_Created = sapply(logo_paths, function(x) is.character(x) && nzchar(x) && file.exists(x))
)

# Print results
print("Logo generation results:")
print(logo_results)

# List all generated PNG files
cat("Generated logo files:\n")
print(list.files("motif_logos", pattern = "*.png", full.names = TRUE))
```

```{r}
motifsDo <- peakAnnoEnrichment(
    seMarker = markerTest,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC <= -0.5"
  )

motifsDo
```

```{r}
df <- data.frame(TF = rownames(motifsDo), mlog10Padj = assay(motifsDo)[,1])
df <- df[order(df$mlog10Padj, decreasing = TRUE),]
df$rank <- seq_len(nrow(df))

ggDo <- ggplot(df, aes(rank, mlog10Padj, color = mlog10Padj)) + 
  geom_point(size = 1) +
  ggrepel::geom_label_repel(
        data = df[rev(seq_len(30)), ], aes(x = rank, y = mlog10Padj, label = TF), 
        size = 1.5,
        nudge_x = 2,
        color = "black",
        max.overlaps = 30  # Increase max.overlaps
  ) + theme_ArchR() + 
  ylab("-log10(FDR) Motif Enrichment") +
  xlab("Rank Sorted TFs Enriched") +
  scale_color_gradientn(colors = paletteContinuous(set = "comet"))

ggDo
```

```{r}
plotPDF(ggUp, ggDo, name = "C1-vs-C2-Markers-Motifs-Enriched", width = 10, height = 10, ArchRProj = proj_motif5, addDOC = FALSE)
#no such index at level 2
```

#Motif Enrichment in Marker Peaks

```{r}
enrichMotifs <- peakAnnoEnrichment(
    seMarker = markersPeaks,
    ArchRProj = proj_motif5,
    peakAnnotation = "Motif",
    cutOff = "FDR <= 0.1 & Log2FC >= 0.5"
  )

heatmapEM <- plotEnrichHeatmap(enrichMotifs, n = 10, transpose = TRUE)
#N determiens the number of motifs

ComplexHeatmap::draw(heatmapEM, heatmap_legend_side = "bot", annotation_legend_side = "bot")

plotPDF(heatmapEM, name = "Motifs-Enriched-Marker-Heatmap", width = 8, height = 6, ArchRProj = proj_motif5, addDOC = FALSE)
```


#Motif Footprinting
```{r}
motifPositions <- getPositions(proj_motif5)
```
```{r}
motifPositions
```
```{r}
motifs <- c("RUNX1")
markerMotifs <- unlist(lapply(motifs, function(x) grep(x, names(motifPositions), value = TRUE)))
markerMotifs <- markerMotifs[markerMotifs %ni% "SREBF1_22"]
markerMotifs
```
```{r}
proj_motif5 <- addGroupCoverages(ArchRProj = proj_motif5, groupBy = "Clusters_TuMerged")
```
```{r}
seFoot <- getFootprints(
  ArchRProj = proj_motif5, 
  positions = motifPositions[markerMotifs], 
  groupBy = "Clusters_TuMerged"
)
```
#Subtract Tn5 Bias
```{r}
plotFootprints(
  seFoot = seFoot,
  ArchRProj = proj_motif5, 
  normMethod = "Subtract",
  plotName = "Footprints-Subtract-Bias",
  addDOC = FALSE,
  smoothWindow = 5
)
```
#Divide by Tn5 Bias
```{r}
plotFootprints(
  seFoot = seFoot,
  ArchRProj = proj_motif5, 
  normMethod = "Divide",
  plotName = "Footprints-Divide-Bias",
  addDOC = FALSE,
  smoothWindow = 5
)
```


