---
title: "SMART-Seq Patient Merge"
author: "KQ"
date: "2025-07-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Initial Set Up
```{r}
setwd("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq")
#/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeqExp2_QC
```
# Set Seed
```{r}
set.seed(34)
```
# Load Libraries Required
```{r}
library(Seurat)
library(ggplot2)
library(patchwork)
library(clusterProfiler)
library(org.Hs.eg.db)
library(viridis)
library(ggrepel)
library(dplyr)
library(stringr)
library(RColorBrewer)
library(tibble)
library(enrichplot)
```
## Process Raw Data for Each Library
# Process through default steps for CRC177 Tumor1
```{r}
counts<-read.delim(file="250306_smartseq3.bbrd.q10.filt.srt.counts.tsv")
CRC177Tumor1 <- CreateSeuratObject(counts = counts, project = "SmartDeep_CRC177Tumor1", min.features = 100)
CRC177Tumor1 <- NormalizeData(object = CRC177Tumor1)
CRC177Tumor1 <- FindVariableFeatures(object = CRC177Tumor1)
CRC177Tumor1 <- ScaleData(object = CRC177Tumor1)

dim(CRC177Tumor1) #27096    92
#It probably means that for 3 cells we have less than 100 features. 

CRC177Tumor1 <- RunPCA(object = CRC177Tumor1, npcs = 91) #One less than Column Value

CRC177Tumor1 <- FindNeighbors(object = CRC177Tumor1, dims = 1:91)
CRC177Tumor1 <- FindClusters(object = CRC177Tumor1)

#Run UMAP
CRC177Tumor1 <- RunUMAP(object = CRC177Tumor1, dims = 1:10)

#Plot UMAP
DimPlot(CRC177Tumor1, reduction = "umap") + plot_annotation(title = "CRC177Tumor1")

#Plot metadata
VlnPlot(CRC177Tumor1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177Tumor1")

# Add CellType metadata to the Seurat object
CRC177Tumor1$CellType <- "CRC177Tumor1"

# Verify that the metadata has been added
head(CRC177Tumor1@meta.data)
```
# Adding Metadata
```{r}
library(tibble)

# Add cell names as a column
CRC177Tumor1@meta.data <- rownames_to_column(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Read the well cell type data
well_celltype_data_CRC177Tumor1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_Tumor_Plate3.csv")

# Merge the data frames
CRC177Tumor1@meta.data <- merge(CRC177Tumor1@meta.data, well_celltype_data_CRC177Tumor1, by.x = "cell_names_CRC177Tumor1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177Tumor1@meta.data <- column_to_rownames(CRC177Tumor1@meta.data, var = "cell_names_CRC177Tumor1")

# Check the result
head(CRC177Tumor1@meta.data)
```
##Updating Metadata for CHC vs. THC Analysis
```{r}
# Update metadata for cell type
CRC177Tumor1@meta.data$WellCellType[CRC177Tumor1@meta.data$WellCellType == "Hybrid"] <- "THC"
# Check the result
table(CRC177Tumor1@meta.data$WellCellType)
# Check the overall meta data structure is unchanged
head(CRC177Tumor1@meta.data)
```
```{r}
# Pull out Patient ID
CRC177Tumor1$PatientID <- "CRC177"
# Check the result
table(CRC177Tumor1@meta.data$PatientID)
# Check the overall meta data structure is unchanged
head(CRC177Tumor1@meta.data)
```

# Process through default steps for CRC177 PMBC1
```{r}
counts2<-read.delim(file="250306_smartseq1.bbrd.q10.filt.srt.counts.tsv")
CRC177PBMC1 <- CreateSeuratObject(counts = counts2, project = "SmartDeep_CRC177PBMC1", min.features = 100)
CRC177PBMC1 <- NormalizeData(object = CRC177PBMC1)
CRC177PBMC1 <- FindVariableFeatures(object = CRC177PBMC1)
CRC177PBMC1 <- ScaleData(object = CRC177PBMC1)

dim(CRC177PBMC1) #23805    93

CRC177PBMC1 <- RunPCA(object = CRC177PBMC1, npcs = 92) #One less than Column Value

CRC177PBMC1 <- FindNeighbors(object = CRC177PBMC1, dims = 1:92)
CRC177PBMC1 <- FindClusters(object = CRC177PBMC1)

#Run UMAP
CRC177PBMC1 <- RunUMAP(object = CRC177PBMC1, dims = 1:10)

#Plot UMAP
DimPlot(CRC177PBMC1, reduction = "umap") + plot_annotation(title = "CRC177Tumor1")

#Plot metadata
VlnPlot(CRC177PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177Tumor1")

# Add CellType metadata to the Seurat object
CRC177PBMC1$CellType <- "CRC177PBMC1"

# Verify that the metadata has been added
head(CRC177PBMC1@meta.data)
```
```{r}
# Check the Start
head(CRC177PBMC1@meta.data)

# Add cell names as a column
CRC177PBMC1@meta.data <- rownames_to_column(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Read the well cell type data
well_celltype_data_CRC177PBMC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate1.csv")

# Merge the data frames
CRC177PBMC1@meta.data <- merge(CRC177PBMC1@meta.data, well_celltype_data_CRC177PBMC1, by.x = "cell_names_CRC177PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC1@meta.data <- column_to_rownames(CRC177PBMC1@meta.data, var = "cell_names_CRC177PBMC1")

# Check the result
head(CRC177PBMC1@meta.data)
```
# Updating Metadata for CHC vs. THC Analysis
```{r}
# Update metadata for cell type
CRC177PBMC1@meta.data$WellCellType[CRC177PBMC1@meta.data$WellCellType == "Hybrid"] <- "CHC"
# Check the result
table(CRC177PBMC1@meta.data$WellCellType)
# Check the overall meta data structure is unchanged
head(CRC177PBMC1@meta.data)
```
```{r}
# Pull out Patient ID
CRC177PBMC1$PatientID <- "CRC177"
# Check the result
table(CRC177PBMC1@meta.data$PatientID)
# Check the overall meta data structure is unchanged
head(CRC177PBMC1@meta.data)
```
# Process through default steps for CRC177 PMBC2
```{r}
counts3<-read.delim(file="250306_smartseq2.bbrd.q10.filt.srt.counts.tsv")
CRC177PBMC2 <- CreateSeuratObject(counts = counts3, project = "SmartDeep_CRC177PBMC2", min.features = 100)
CRC177PBMC2 <- NormalizeData(object = CRC177PBMC2)
CRC177PBMC2 <- FindVariableFeatures(object = CRC177PBMC2)
CRC177PBMC2 <- ScaleData(object = CRC177PBMC2)

dim(CRC177PBMC2) #25962    88

CRC177PBMC2 <- RunPCA(object = CRC177PBMC2, npcs = 87) #One less than Column Value

CRC177PBMC2 <- FindNeighbors(object = CRC177PBMC2, dims = 1:87)
CRC177PBMC2 <- FindClusters(object = CRC177PBMC2)

#Run UMAP
CRC177PBMC2 <- RunUMAP(object = CRC177PBMC2, dims = 1:10)

#Plot UMAP
DimPlot(CRC177PBMC2, reduction = "umap") + plot_annotation(title = "CRC177PBMC2")

#Plot metadata
VlnPlot(CRC177PBMC2, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC177PBMC2")

# Add CellType metadata to the Seurat object
CRC177PBMC2$CellType <- "CRC177PBMC2"

# Verify that the metadata has been added
head(CRC177PBMC2@meta.data)
```
```{r}
# Check the Start
head(CRC177PBMC2@meta.data)

# Add cell names as a column
CRC177PBMC2@meta.data <- rownames_to_column(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Read the well cell type data
well_celltype_data_CRC177PBMC2 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC177PBMC2@meta.data <- merge(CRC177PBMC2@meta.data, well_celltype_data_CRC177PBMC2, by.x = "cell_names_CRC177PBMC2", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC177PBMC2@meta.data <- column_to_rownames(CRC177PBMC2@meta.data, var = "cell_names_CRC177PBMC2")

# Check the result
head(CRC177PBMC2@meta.data)
```
# Updating Metadata for CHC vs. THC Analysis
```{r}
# Update metadata for cell type
CRC177PBMC2@meta.data$WellCellType[CRC177PBMC2@meta.data$WellCellType == "Hybrid"] <- "CHC"
# Check the result
table(CRC177PBMC2@meta.data$WellCellType)
# Check the overall meta data structure is unchanged
head(CRC177PBMC2@meta.data)
```
```{r}
# Pull out Patient ID
CRC177PBMC2$PatientID <- "CRC177"
# Check the result
table(CRC177PBMC2@meta.data$PatientID)
# Check the overall meta data structure is unchanged
head(CRC177PBMC2@meta.data)
```

# Process through default steps for CRC179 PBMC Plate 1
```{r}
counts4<-read.delim(file="250226_smartseq1.bbrd.q10.filt.srt.counts.tsv")
CRC179PBMC1 <- CreateSeuratObject(counts = counts4, project = "SmartDeep_CRC179PBMC1", min.features = 100)
CRC179PBMC1 <- NormalizeData(object = CRC179PBMC1)
CRC179PBMC1 <- FindVariableFeatures(object = CRC179PBMC1)
CRC179PBMC1 <- ScaleData(object = CRC179PBMC1)

dim(CRC179PBMC1) #21322    90

CRC179PBMC1 <- RunPCA(object = CRC179PBMC1, npcs = 89) #One less than Column Value

CRC179PBMC1 <- FindNeighbors(object = CRC179PBMC1, dims = 1:89)
CRC179PBMC1 <- FindClusters(object = CRC179PBMC1)

#Run UMAP
CRC179PBMC1 <- RunUMAP(object = CRC179PBMC1, dims = 1:10)

#Plot UMAP
DimPlot(CRC179PBMC1, reduction = "umap") + plot_annotation(title = "CRC179PBMC1")

#Plot metadata
VlnPlot(CRC179PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC179PBMC1")

# Add CellType metadata to the Seurat object
CRC179PBMC1$CellType <- "CRC179PBMC1"

# Verify that the metadata has been added
head(CRC179PBMC1@meta.data)
```
```{r}
# Check the Start
head(CRC179PBMC1@meta.data)

# Add cell names as a column
CRC179PBMC1@meta.data <- rownames_to_column(CRC179PBMC1@meta.data, var = "cell_names_CRC179PBMC1")

# Read the well cell type data
well_celltype_data_CRC179PMBC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate1.csv")

# Merge the data frames
CRC179PBMC1@meta.data <- merge(CRC179PBMC1@meta.data, well_celltype_data_CRC179PMBC1, by.x = "cell_names_CRC179PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC179PBMC1@meta.data <- column_to_rownames(CRC179PBMC1@meta.data, var = "cell_names_CRC179PBMC1")

# Check the result
head(CRC179PBMC1@meta.data)
```
##Updating Metadata for CHC vs. THC Analysis
```{r}
# Update metadata for cell type
CRC179PBMC1@meta.data$WellCellType[CRC179PBMC1@meta.data$WellCellType == "Hybrid"] <- "CHC"
# Check the result
table(CRC179PBMC1@meta.data$WellCellType)
# Check the overall meta data structure is unchanged
head(CRC179PBMC1@meta.data)
```
```{r}
# Pull out Patient ID
CRC179PBMC1$PatientID <- "CRC179"
# Check the result
table(CRC179PBMC1@meta.data$PatientID)
# Check the overall meta data structure is unchanged
head(CRC179PBMC1@meta.data)
```

# Process through default steps for CRC176 PBMC Plate 1
```{r}
counts5<-read.delim(file="250226_smartseq2.bbrd.q10.filt.srt.counts.tsv")
CRC176PBMC1 <- CreateSeuratObject(counts = counts5, project = "SmartDeep_CRC176PBMC1", min.features = 100)
CRC176PBMC1 <- NormalizeData(object = CRC176PBMC1)
CRC176PBMC1 <- FindVariableFeatures(object = CRC176PBMC1)
CRC176PBMC1 <- ScaleData(object = CRC176PBMC1)

dim(CRC176PBMC1) #21694    91

CRC176PBMC1 <- RunPCA(object = CRC176PBMC1, npcs = 90) #One less than Column Value

CRC176PBMC1 <- FindNeighbors(object = CRC176PBMC1, dims = 1:90)
CRC176PBMC1 <- FindClusters(object = CRC176PBMC1)

#Run UMAP
CRC176PBMC1 <- RunUMAP(object = CRC176PBMC1, dims = 1:10)
 
#Plot UMAP
DimPlot(CRC176PBMC1, reduction = "umap") + plot_annotation(title = "CRC176PBMC1")

#Plot metadata
VlnPlot(CRC176PBMC1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC176PBMC1")

# Add CellType metadata to the Seurat object
CRC176PBMC1$CellType <- "CRC176PBMC1"

# Verify that the metadata has been added
head(CRC176PBMC1@meta.data)
```

```{r}
# Check the Start
head(CRC176PBMC1@meta.data)

# Add cell names as a column
CRC176PBMC1@meta.data <- rownames_to_column(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Read the well cell type data
well_celltype_data_CRC176PMBC1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_PMBC_Plate2.csv")

# Merge the data frames
CRC176PBMC1@meta.data <- merge(CRC176PBMC1@meta.data, well_celltype_data_CRC176PMBC1, by.x = "cell_names_CRC176PBMC1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC176PBMC1@meta.data <- column_to_rownames(CRC176PBMC1@meta.data, var = "cell_names_CRC176PBMC1")

# Check the result
head(CRC176PBMC1@meta.data)
```
##Updating Metadata for CHC vs. THC Analysis
```{r}
# Update metadata for cell type
CRC176PBMC1@meta.data$WellCellType[CRC176PBMC1@meta.data$WellCellType == "Hybrid"] <- "CHC"
# Check the result
table(CRC176PBMC1@meta.data$WellCellType)
# Check the overall meta data structure is unchanged
head(CRC176PBMC1@meta.data)
```
```{r}
# Pull out Patient ID
CRC176PBMC1$PatientID <- "CRC176"
# Check the result
table(CRC176PBMC1@meta.data$PatientID)
# Check the overall meta data structure is unchanged
head(CRC176PBMC1@meta.data)
```

# Process through default steps for CRC176 Tumor Plate 1
```{r}
counts6<-read.delim(file="250226_smartseq3.bbrd.q10.filt.srt.counts.tsv")
CRC176Tumor1 <- CreateSeuratObject(counts = counts6, project = "SmartDeep_CRC176Tumor1", min.features = 100)
CRC176Tumor1 <- NormalizeData(object = CRC176Tumor1)
CRC176Tumor1 <- FindVariableFeatures(object = CRC176Tumor1)
CRC176Tumor1 <- ScaleData(object = CRC176Tumor1)

dim(CRC176Tumor1) #34747    85

CRC176Tumor1 <- RunPCA(object = CRC176Tumor1, npcs = 84) #One less than Column Value

CRC176Tumor1 <- FindNeighbors(object = CRC176Tumor1, dims = 1:84)
CRC176Tumor1 <- FindClusters(object = CRC176Tumor1)

#Run UMAP
CRC176Tumor1 <- RunUMAP(object = CRC176Tumor1, dims = 1:10)

#Plot UMAP
DimPlot(CRC176Tumor1, reduction = "umap") + plot_annotation(title = "CRC176Tumor1")

#Plot metadata
VlnPlot(CRC176Tumor1, features = c("nFeature_RNA","nCount_RNA")) + plot_annotation(title = "CRC176Tumor1")

# Add CellType metadata to the Seurat object
CRC176Tumor1$CellType <- "CRC176Tumor1"

# Verify that the metadata has been added
head(CRC176Tumor1@meta.data)
```

```{r}
# Check the Start
head(CRC176Tumor1@meta.data)

# Add cell names as a column
CRC176Tumor1@meta.data <- rownames_to_column(CRC176Tumor1@meta.data, var = "cell_names_CRC176Tumor1")

# Read the well cell type data
well_celltype_data_CRC176Tumor1 <- read.csv("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/WellAnnotationFiles/WellAnnotation_Tumor_Plate3.csv")

# Merge the data frames
CRC176Tumor1@meta.data <- merge(CRC176Tumor1@meta.data, well_celltype_data_CRC176Tumor1, by.x = "cell_names_CRC176Tumor1", by.y = "CellID", all.x = TRUE)

# Convert the cell names column back to row names
CRC176Tumor1@meta.data <- column_to_rownames(CRC176Tumor1@meta.data, var = "cell_names_CRC176Tumor1")

# Check the result
head(CRC176Tumor1@meta.data)
```
##Updating Metadata for CHC vs. THC Analysis
```{r}
# Update metadata for cell type
CRC176Tumor1@meta.data$WellCellType[CRC176Tumor1@meta.data$WellCellType == "Hybrid"] <- "THC"
# Check the result
table(CRC176Tumor1@meta.data$WellCellType)
# Check the overall meta data structure is unchanged
head(CRC176Tumor1@meta.data)
```
```{r}
# Pull out Patient ID
CRC176Tumor1$PatientID <- "CRC176"
# Check the result
table(CRC176Tumor1@meta.data$PatientID)
# Check the overall meta data structure is unchanged
head(CRC176Tumor1@meta.data)
```

# PreMerge Trouble Shooting
```{r}
# Check metadata columns in each Seurat object
head(CRC177Tumor1@meta.data)
head(CRC177PBMC1@meta.data)
head(CRC177PBMC2@meta.data)
head(CRC176PBMC1@meta.data)
head(CRC176Tumor1@meta.data)
head(CRC179PBMC1@meta.data)
```
```{r}
sum(is.na(CRC177PBMC2$CellType))
sum(is.na(CRC177PBMC2$WellCellType))
sum(is.na(CRC177PBMC1$CellType))
sum(is.na(CRC177PBMC1$WellCellType))
sum(is.na(CRC176PBMC1$CellType))
sum(is.na(CRC176PBMC1$WellCellType))
sum(is.na(CRC179PBMC1$CellType))
sum(is.na(CRC179PBMC1$WellCellType))
sum(is.na(CRC177Tumor1$CellType))
sum(is.na(CRC177Tumor1$WellCellType))
sum(is.na(CRC176Tumor1$CellType))
sum(is.na(CRC176Tumor1$WellCellType))
```
```{r}
any(duplicated(colnames(CRC177Tumor1)))
any(duplicated(colnames(CRC177PBMC1)))
any(duplicated(colnames(CRC176PBMC1)))
any(duplicated(colnames(CRC177PBMC2)))
any(duplicated(colnames(CRC179PBMC1)))
any(duplicated(colnames(CRC176Tumor1)))
```

# Attempt to Merge
```{r}
# Merge the Seurat objects with unique cell names
merged_data_all <- merge(CRC177Tumor1, 
                     y = list(CRC177PBMC1, CRC177PBMC2, CRC176PBMC1, CRC176Tumor1, CRC179PBMC1), 
                     add.cell.ids = c("CRC177Tumor1", "CRC177PBMC1", "CRC177PBMC2", "CRC176PBMC1", "CRC176Tumor1", "CRC179PMBC1"))

# Verify the merged object
print(merged_data_all)
```

# Process Merged Seurat Object
```{r}
merged_data_all <- NormalizeData(object = merged_data_all) #Might not be required again
merged_data_all <- FindVariableFeatures(merged_data_all, selection.method = "vst", nfeatures = 2000) 
merged_data_all <- ScaleData(object = merged_data_all)
```
```{r}
dim(merged_data_all) #40287   539
# Check the dimensions of the scaled data matrix
```

# Run PCA
```{r}
# Run PCA
# Run PCA with a reduced number of PCs
merged_data_all <- RunPCA(merged_data_all, npcs = 50)  # Adjust npcs as needed
# Verify the PCA results
print(merged_data_all)
```
```{r}
# Check the variance explained by each PC
ElbowPlot(merged_data_all) #10 looks to be optimal
```
# Determine Clusters
```{r}
merged_data_all <- FindNeighbors(object = merged_data_all, dims = 1:10)
```
#Plot UMAP
```{r}
set.seed(48)
merged_data_all <- FindClusters(merged_data_all, resolution = 0.9) # Adjust the resolution value as needed. I decided on this, because it separates out the Cluster 5 - one of which ends up in CHC3 and one of which is largely composed of THC cells. 
merged_data_all <- RunUMAP(merged_data_all, 
                          dims = 1:10, 
                          seed.use = 48,     # Explicit UMAP seed
                          verbose = TRUE)
DimPlot(merged_data_all, reduction = "umap", pt.size = 2, alpha = 0.8) +
  scale_color_viridis(discrete = TRUE) + plot_annotation(title = "0.9")

set.seed(48)
merged_data_all <- FindClusters(merged_data_all, resolution = 0.75) # Adjust the resolution value as needed
merged_data_all <- RunUMAP(merged_data_all, 
                          dims = 1:10, 
                          seed.use = 48,     # Explicit UMAP seed
                          verbose = TRUE)
DimPlot(merged_data_all, reduction = "umap", pt.size = 2, alpha = 0.8) +
  scale_color_viridis(discrete = TRUE) + plot_annotation(title = "0.75")
```

#Save RDS of Seurat Object
```{r}
saveRDS(merged_data_all, file = "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/SmartSeq_CRC177_DeepSeq/SmartSeqMerged.rds")
```