---
title: "20240319_artificialdoublets2"
author: "Ashley Anderson, Konstantin Queitsch"
date: "2025-04-25"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("scDblFinder")
# 
# install.packages("reticulate")
# install.packages("ggExtra")
```

## Start Up

# Loading Required Packages
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
```
#Setting a Seeds
```{r}
set.seed(34)
```
# Install uncommon package
```{r}
# install presto for faster implementation of wilcoxon rank sum test 
#install.packages('devtools')
#devtools::install_github('immunogenomics/presto')
```

## Artificial Doublet Analysis

# Read in counts matrix & barcode file with cellidentity, make seurat object with cell identity 
```{r}
counts <- read.delim(file="/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/ArtDup Experiment/231201_ArtificialDoublets.bbrd.q10.filt.srt.counts.tsv")

barcode_annotation <- read.table("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/ArtDup Experiment/240319_BarcodesAnnotated.txt", header = TRUE, stringsAsFactors = FALSE)


doublets2 <- CreateSeuratObject(counts = counts, project = "artificial doublets2", min.features = 500)


row.names(barcode_annotation) <- barcode_annotation$Cells

#head(Cells(doublets2))
#head(barcode_annotation$Cells)
#colnames(barcode_annotation)
colnames(barcode_annotation) <- c("Cells", "CellIdentity")

row.names(barcode_annotation) <- barcode_annotation$Cells
doublets2_meta <- AddMetaData(object = doublets2, metadata = barcode_annotation)

View(doublets2_meta)


# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
doublets2_meta[["percent.mt"]] <- PercentageFeatureSet(doublets2, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(doublets2_meta, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
```


#subset based on data quality to remove cells with high MT DNA and low nFeature RNA
```{r}
doublets2_meta <- subset(doublets2_meta, subset = nFeature_RNA > 1000 & percent.mt < 15)
```

#Normalizing Data. employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
```{r}
doublets2_meta <- NormalizeData(doublets2_meta, normalization.method = "LogNormalize", scale.factor = 10000)
```


#finding highly variable features 
```{r}
doublets2_meta <- FindVariableFeatures(doublets2_meta, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(doublets2_meta), 10)
top10

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(doublets2_meta)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 
plot2
```

#scaling the data 
```{r}
all.genes <- rownames(doublets2_meta)
doublets2_meta <- ScaleData(doublets2_meta, features = all.genes)
```

#perform linear dimensionality reduction
```{r}
doublets2_meta <- RunPCA(doublets2_meta, features = VariableFeatures(object = doublets2_meta))

# Examine and visualize PCA results a few different ways
print(doublets2_meta[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(doublets2_meta, dims = 1:5, reduction = "pca")
VizDimLoadings(doublets2_meta, dims = 1:2, reduction = "pca")
VizDimLoadings(doublets2_meta, dims = 3:4, reduction = "pca")
VizDimLoadings(doublets2_meta, dims = 4:5, reduction = "pca")

DimPlot(doublets2_meta, reduction = "pca")
DimHeatmap(doublets2_meta, dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(doublets2_meta, dims = 1:5, cells = 500, balanced = TRUE)
```

#determine dimensionality of the dataset 
```{r}
ElbowPlot(doublets2_meta)
```

#cluster the cells
```{r}
doublets2_meta <- FindNeighbors(doublets2_meta, dims = 1:10)
doublets2_meta <- FindClusters(doublets2_meta, dims = 1:10, resolution = 0.5)
head(Idents(doublets2_meta), 5)
```

#run linear dimensionality reduction UMAP
```{r}
doublets2_meta <- RunUMAP(doublets2_meta, dims = 1:10)
DimPlot(doublets2_meta, reduction = "umap")

DimPlot(doublets2_meta, group.by = 'CellIdentity')
DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("red", "gray","gray", "blue","gray","gray", "gray", "gray"))
DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "green","gray", "gray","gray","gray", "gray", "pink"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "red","gray", "gray","gray","gray", "gray", "gray"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "gray","red", "gray","gray","gray", "gray", "gray"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "red","gray","gray", "gray", "gray"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","red","gray", "gray", "gray"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","gray","red", "gray", "gray"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","gray","gray", "red", "gray"))

DimPlot(doublets2_meta, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","gray","gray", "gray", "red"))



FeaturePlot(doublets2_meta, features = c("Spp1", "Cxcl9", "Il1a"))


```

# Subset data to remove the "Remove" cells
```{r}
#cells-to-include <- doublets2_meta$orig.ident = c("HF", "Mac", "Tum", "Mac-Tum")
cells_to_include <- c("HF", "Mac", "Tum", "Mac-Tum")
doublets2_subset <- subset(doublets2_meta, subset = CellIdentity %in% cells_to_include)

```
# Run UMAP on Subsetted Data
```{r}
doublets2_subset <- RunUMAP(doublets2_subset, dims = 1:10)
DimPlot(doublets2_subset, reduction = "umap")

DimPlot(doublets2_subset, group.by = 'CellIdentity', pt.size = 2)
DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("red", "gray","blue", "grey"), pt.size = 2) 
DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("grey", "green","blue", "grey"), pt.size = 2) 
#DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "green","gray", "gray","gray","gray", "gray", "pink"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "red","gray", "gray","gray","gray", "gray", "gray"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "gray","red", "gray","gray","gray", "gray", "gray"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "red","gray","gray", "gray", "gray"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","red","gray", "gray", "gray"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","gray","red", "gray", "gray"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","gray","gray", "red", "gray"))
# 
# DimPlot(doublets2_subset, group.by = 'CellIdentity', cols = c("gray", "gray","gray", "gray","gray","gray", "gray", "red"))
# 
# 
# 
#FeaturePlot(doublets2_subset, features = c("Spp1", "Cxcl9", "Il1a", "Runx1")) #AA
FeaturePlot(doublets2_subset, features = c("Epcam", "Spi1", "Runx1", "Cldn1")) #Hybrids
FeaturePlot(doublets2_subset, features = c("Epcam", "Cldn1", "Satb2")) #MC38?
FeaturePlot(doublets2_subset, features = c("Cd14", "Ccr5", "Adgre1")) #Macrophage Cluster
#Adgre1 codes for F4/80 receptor which is highly and constitutevly expressed on resident tissue macrophage
#Was previously used in the 2011 Wong Paper.

```

```{r}
library(ggplot2)
library(patchwork)

# Create the FeaturePlots
plot1 <- FeaturePlot(doublets2_subset, features = "Epcam")
plot2 <- FeaturePlot(doublets2_subset, features = "Cldn1")
plot3 <- FeaturePlot(doublets2_subset, features = "Satb2")

# Add a centered title
combined_plot <- plot1 + plot2 + plot3 + plot_annotation(title = "MC38 Canonical Genes")
combined_plot


```

#finding differentially expressed features 
```{r}

doublets2_meta <- JoinLayers(doublets2_meta)

# find all markers for each individual cluster 0-5
cluster0.markers <- FindMarkers(doublets2_meta, ident.1 = 0, min.pct = 0.25)
head(cluster0.markers, n = 20)
write.csv(cluster0.markers, file = "doublets2_meta_cluster0_markers")
View(cluster0.markers)

cluster1.markers <- FindMarkers(doublets2_meta, ident.1 = 1, min.pct = 0.25)
write.csv(cluster1.markers, file = "doublets2_meta_cluster1_markers")
head(cluster1.markers, n = 20)


cluster2.markers <- FindMarkers(doublets2_meta, ident.1 = 2, min.pct = 0.25)
write.csv(cluster2.markers, file = "doublets2_meta_cluster2_markers")
head(cluster2.markers, n = 20)


cluster3.markers <- FindMarkers(doublets2_meta, ident.1 = 3, min.pct = 0.25)
write.csv(cluster3.markers, file = "doublets2_meta_cluster3_markers")
head(cluster3.markers, n = 20)


cluster4.markers <- FindMarkers(doublets2_meta, ident.1 = 4, min.pct = 0.25)
write.csv(cluster4.markers, file = "doublets2_meta_cluster4_markers")
head(cluster4.markers, n = 20)


cluster5.markers <- FindMarkers(doublets2_meta, ident.1 = 5, min.pct = 0.25)
write.csv(cluster5.markers, file = "doublets2_meta_cluster5_markers")
head(cluster5.markers, n = 20)


# find markers for every cluster compared to all remaining cells, report only the positive ones
doublets2_meta.markers <- FindAllMarkers(doublets2_meta, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
doublets2_meta.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(doublets2_meta.markers, file = "doublets2_meta_pos_markers")

doublets2_meta.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(doublets2_meta, features = top10$gene) + NoLegend()


```

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("org.Mm.eg.db") #org.Mm.eg.db for mouse


if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("clusterProfiler")


if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("AnnotationDbi")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("tidyverse")


if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("RColorBrewer")

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("fgsea")
```

```{r}
library(AnnotationDbi)
#library(tidyverse)
library(org.Mm.eg.db)
library(clusterProfiler)
library(fgsea)
library(RColorBrewer)
```

##GSEA analysis hybrid cluster 1 and 4
```{r}

##cluster1

cluster1.markers$gene <- rownames(cluster1.markers)
cluster1.markers <- cluster1.markers %>% arrange(desc(avg_log2FC))
cluster1_foldchange <- cluster1.markers$avg_log2FC
names(cluster1_foldchange) <- cluster1.markers$gene

head(cluster1_foldchange)

cluster1_foldchange.df <- as.data.frame(cluster1_foldchange)

cluster1_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster1_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster1_enslist <- na.omit(cluster1_foldchange.df)

cluster1_enslist <- cluster1_foldchange.df$cluster1_foldchange
names(cluster1_enslist) <- cluster1_foldchange.df$ENSEMBL
cluster1_enslist

View(cluster1_enslist)

gse.cluster1 <- gseGO(cluster1_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster1.df <- as.data.frame(gse.cluster1)

write.csv(gse.cluster1.df, file = "doublets2_meta.gse.cluster1.csv")


##visualization
require(DOSE)
dotplot(gse.cluster1, showCategory=10, split=".sign", font.size = 18) + facet_grid(.~.sign)


gse.cluster1.2 <- pairwise_termsim(gse.cluster1) # calculate pairwise similarities of the enriched terms using Jaccard’s similarity index

treeplot(gse.cluster1.2)

emapplot(gse.cluster1.2)


##cluster4

cluster4.markers$gene <- rownames(cluster4.markers)
cluster4.markers <- cluster4.markers %>% arrange(desc(avg_log2FC))
cluster4_foldchange <- cluster4.markers$avg_log2FC
names(cluster4_foldchange) <- cluster4.markers$gene

head(cluster4_foldchange)

cluster4_foldchange.df <- as.data.frame(cluster4_foldchange)

cluster4_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster4_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster4_enslist <- na.omit(cluster4_foldchange.df)

cluster4_enslist <- cluster4_foldchange.df$cluster4_foldchange
names(cluster4_enslist) <- cluster4_foldchange.df$ENSEMBL
cluster4_enslist

View(cluster4_enslist)

gse.cluster4 <- gseGO(cluster4_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster4.df <- as.data.frame(gse.cluster4)

write.csv(gse.cluster4.df, file = "doublets2_meta.gse.cluster4.csv")


##visualization
require(DOSE)
dotplot(gse.cluster4, showCategory=10, split=".sign", font.size = 18) + facet_grid(.~.sign)


gse.cluster4.2 <- pairwise_termsim(gse.cluster4) # calculate pairwise similarities of the enriched terms using Jaccard’s similarity index

treeplot(gse.cluster4.2)

emapplot(gse.cluster4.2)

```

#DEG hybrid clusters 1 and 4
```{r}

library(ggplot2)
library(ggrepel)

##hybrid cluster1

View(cluster1.markers)

# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
cluster1.markers$diffexpressed <- "NO"
cluster1.markers$diffexpressed[cluster1.markers$avg_log2FC > 2 & cluster1.markers$p_val_adj < 0.05] <- "UP"
cluster1.markers$diffexpressed[cluster1.markers$avg_log2FC < -2 & cluster1.markers$p_val_adj < 0.05] <- "DOWN"


View(cluster1.markers)

# Create a new column "delabel" to de, that will contain the name of the top 30 differentially expressed genes (NA in case they are not)
cluster1.markers$delabel <- ifelse(cluster1.markers$gene %in% head(cluster1.markers[order(cluster1.markers$p_val_adj), "gene"], 30), cluster1.markers$gene, NA)

View(cluster1.markers)

cluster1_volcano <- ggplot(data = cluster1.markers, aes(x = avg_log2FC, y = -log10(p_val_adj), col = diffexpressed, label = delabel)) +
  geom_vline(xintercept = c(-2, 2), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = 75, col = "gray", linetype = 'dashed') +
  geom_point(size = 2) +
  scale_color_manual(values = c("#00AFBB", "grey", "#bb0c00"), labels = c("Downregulated", "Not significant", "Upregulated")) +
  coord_cartesian(ylim = c(0, 150), xlim = c(-8, 8)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = 'Gene Expression', #legend_title,
       x = expression("log"[2]*"FC"), y = expression("-log"[10]*"p-value")) +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customise the breaks in the x axis
  ggtitle('Hybrid Clone4 Cluster1 DEGs ') + 
  geom_text_repel(max.overlaps = Inf)
 

cluster1_volcano


##hybrid cluster4

View(cluster4.markers)

# Add a column to the data frame to specify if they are UP- or DOWN- regulated (log2fc respectively positive or negative)
cluster4.markers$diffexpressed <- "NO"
cluster4.markers$diffexpressed[cluster4.markers$avg_log2FC > 1 & cluster4.markers$p_val_adj < 0.05] <- "UP"
cluster4.markers$diffexpressed[cluster4.markers$avg_log2FC < -1 & cluster4.markers$p_val_adj < 0.05] <- "DOWN"


View(cluster4.markers)

# Create a new column "delabel" to de, that will contain the name of the top 30 differentially expressed genes (NA in case they are not)
cluster4.markers$delabel <- ifelse(cluster4.markers$gene %in% head(cluster4.markers[order(cluster4.markers$p_val_adj), "gene"], 30), cluster4.markers$gene, NA)

View(cluster4.markers)

cluster4_volcano <- ggplot(data = cluster4.markers, aes(x = avg_log2FC, y = -log10(p_val_adj), col = diffexpressed, label = delabel)) +
  geom_vline(xintercept = c(-2, 2), col = "gray", linetype = 'dashed') +
  geom_hline(yintercept = 25, col = "gray", linetype = 'dashed') +
  geom_point(size = 2) +
  scale_color_manual(values = c("#00AFBB", "grey", "#bb0c00"), labels = c("Downregulated", "Not significant", "Upregulated")) +
  coord_cartesian(ylim = c(0, 80), xlim = c(-8, 6)) + # since some genes can have minuslog10padj of inf, we set these limits
  labs(color = 'Gene Expression', #legend_title,
       x = expression("log"[2]*"FC"), y = expression("-log"[10]*"p-value")) +
  scale_x_continuous(breaks = seq(-10, 10, 2)) + # to customize the breaks in the x axis
  ggtitle('Hybrid Clone4 cluster4 DEGs ') + 
  geom_text_repel(max.overlaps = Inf)
 

cluster4_volcano
```
