---
title: "MC38xY01hybrid_scRNAseq"
author: "Ashley Anderson"
date: "2023-10-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#installing seurat
```{r}
# install.packages('Seurat')
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# 
# BiocManager::install("biomaRt")
# install.packages("rmarkdown")
# install.packages("dplyr")
```


#loading packages
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
```

```{r}
#Set working directory
setwd("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF")

# List of directories containing your samples
sample_dirs <- list("/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/HF03", "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/HF04", "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/HF07", "/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/HF11","/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/HFF","/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/HFP","/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/MC38","/Users/queitsch/Documents/AdeyLab/1_Hybrid_Fusion/240918_10xscRNA_HF/Macrophage")

# Load the dataset 
hybrid.data <- lapply(sample_dirs, function(dir) {Read10X(data.dir = dir) %>% CreateSeuratObject()})

```
```{r}
combined_seurat <- Reduce(function(x, y) merge(x, y), hybrid.data)
#Warning: Some cell names are duplicated across objects provided. Renaming to enforce unique cell names.
```
```{r}
# Create a vector of sample names corresponding to each directory
sample_names <- c("HF03", "HF04", "HF07", "HF11", "HFF", "HFP", "MC38", "Macrophage")

# Add sample information to each Seurat object
for (i in seq_along(hybrid.data)) {
  hybrid.data[[i]]$sample <- sample_names[i]
}

# Merge the Seurat objects again
combined_seurat <- Reduce(function(x, y) merge(x, y), hybrid.data)

```
```{r}
# ,,, and integrate to account for batch affects
seurat_list <- SplitObject(combined_seurat, split.by = "sample")
seurat_list <- lapply(seurat_list, function(x) {
  x <- NormalizeData(x)
  x <- FindVariableFeatures(x)
  return(x)
})
```

```{r}

features <- SelectIntegrationFeatures(object.list = seurat_list)

```
```{r}
anchors <- FindIntegrationAnchors(object.list = seurat_list, anchor.features = features)
```
```{r}
integrated_seurat <- IntegrateData(anchorset = anchors)


```


#QC and selecting cells for analysis
```{r}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
hybrid[["percent.mt"]] <- PercentageFeatureSet(hybrid, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(hybrid, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
VlnPlot(hybrid, "percent.mt")

# FeatureScatter is typically used to visualize feature-feature relationships, but can be used for anything calculated by the object, i.e. columns in object metadata, PC scores etc.

plot1 <- FeatureScatter(hybrid, feature1 = "nCount_RNA", feature2 = "percent.mt")
plot2 <- FeatureScatter(hybrid, feature1 = "nCount_RNA", feature2 = "nFeature_RNA")
plot1 
plot2


hybrid <- subset(hybrid, subset = nFeature_RNA > 1000 & nFeature_RNA < 8000 & percent.mt < 5)
```

#Normalizing Data. employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result.
```{r}
hybrid <- NormalizeData(hybrid, normalization.method = "LogNormalize", scale.factor = 10000)
```

#finding highly variable features 
```{r}
hybrid <- FindVariableFeatures(hybrid, selection.method = "vst", nfeatures = 2000)

# Identify the 10 most highly variable genes
top10 <- head(VariableFeatures(hybrid), 10)
top10

# plot variable features with and without labels
plot1 <- VariableFeaturePlot(hybrid)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot1 
plot2
```

#scaling the data 
```{r}
all.genes <- rownames(hybrid)
hybrid <- ScaleData(hybrid, features = all.genes)
```

#perform linear dimensionality reduction
```{r}
hybrid <- RunPCA(hybrid, features = VariableFeatures(object = hybrid))

# Examine and visualize PCA results a few different ways
print(hybrid[["pca"]], dims = 1:5, nfeatures = 5)

VizDimLoadings(hybrid, dims = 1:5, reduction = "pca")
VizDimLoadings(hybrid, dims = 1:2, reduction = "pca")
VizDimLoadings(hybrid, dims = 3:4, reduction = "pca")
VizDimLoadings(hybrid, dims = 4:5, reduction = "pca")

DimPlot(hybrid, reduction = "pca")
DimHeatmap(hybrid, dims = 1, cells = 500, balanced = TRUE)
DimHeatmap(hybrid, dims = 1:5, cells = 500, balanced = TRUE)
```

#determine dimensionality of the dataset 
```{r}
hybrid <- JackStraw(hybrid, num.replicate = 100)
hybrid <- ScoreJackStraw(hybrid, dims = 1:20)
JackStrawPlot(hybrid, dims = 1:15)
ElbowPlot(hybrid)
```

#cluster the cells
```{r}
hybrid <- FindNeighbors(hybrid, dims = 1:10)
hybrid <- FindClusters(hybrid, resolution = 0.5)
head(Idents(hybrid), 5)
```

#run linear dimensionality reduction UMAP
```{r}
hybrid <- RunUMAP(hybrid, dims = 1:10)
DimPlot(hybrid, reduction = "umap")

DimPlot(hybrid, group.by = 'orig.ident', pt.size = 0.1)

saveRDS(hybrid, file = "~/Documents/HybridFusionSeq/hybridfusionseq-onlyhybrids.rds")
```

#finding differentially expressed features 
```{r}
# install limma for a more efficient implementation of the Wilcoxon Rank Sum Test
install.packages('BiocManager')
BiocManager::install('limma')

# find all markers for each individual cluster 0-7
cluster0.markers <- FindMarkers(hybrid, ident.1 = 0, min.pct = 0.25)
head(cluster0.markers, n = 20)
write.csv(cluster0.markers, file = "cluster0_markers")
View(cluster0.markers)

cluster1.markers <- FindMarkers(hybrid, ident.1 = 1, min.pct = 0.25)
write.csv(cluster1.markers, file = "cluster1_markers")
head(cluster1.markers, n = 20)

cluster2.markers <- FindMarkers(hybrid, ident.1 = 2, min.pct = 0.25)
write.csv(cluster2.markers, file = "cluster2_markers")
head(cluster2.markers, n = 20)

cluster3.markers <- FindMarkers(hybrid, ident.1 = 3, min.pct = 0.25)
write.csv(cluster3.markers, file = "cluster3_markers")
head(cluster3.markers, n = 20)

cluster4.markers <- FindMarkers(hybrid, ident.1 = 4, min.pct = 0.25)
write.csv(cluster4.markers, file = "cluster4_markers")
head(cluster4.markers, n = 20)

cluster5.markers <- FindMarkers(hybrid, ident.1 = 5, min.pct = 0.25)
write.csv(cluster5.markers, file = "cluster5_markers")
head(cluster5.markers, n = 20)

cluster6.markers <- FindMarkers(hybrid, ident.1 = 6, min.pct = 0.25)
write.csv(cluster6.markers, file = "cluster6_markers")
head(cluster6.markers, n = 20)

cluster7.markers <- FindMarkers(hybrid, ident.1 = 7, min.pct = 0.25)
write.csv(cluster7.markers, file = "cluster7_markers")
head(cluster7.markers, n = 20)


# find markers for every cluster compared to all remaining cells, report only the positive ones
hybrid.markers <- FindAllMarkers(hybrid, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
hybrid.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(hybrid.markers, file = "hybrid_markers")

hybrid.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(hybrid, features = top10$gene) + NoLegend()


saveRDS(hybrid, file = "~/Documents/HybridFusionSeq/hybridfusionseq-onlyhybrids.rds")
```


#specific plots for genes of interest
```{r}
FeaturePlot(hybrid, features = c("Lgals3", "Runx1", "Spi1", "Grn", "Pf4", "Thbs1"))

FeaturePlot(hybrid, features = c("Thbs1"))
FeaturePlot(hybrid, features = c("Ly6c2", "Vcan", "Ccr2", "Nlrp3", "Fcna"))

FeaturePlot(hybrid, features = c("Ki67", "Pcna"))

FeaturePlot(hybrid, features = c("Muc1", "Muc5ac", "Epcam", "Egfr", "Krt8", "Krt5", "Krt14", "Krt19", "Krt9", "Krt7", "Krt18", "Krt10", pt.size = 0.01))

FeaturePlot(hybrid, features = c("Vim", "Acta2"))

FeaturePlot(hybrid, features = c("Ptprc", "Cd68", "Cd14","Fcgr3"))
#note Fcgr3 = CD16 and Ptprc = CD45

FeaturePlot(hybrid, features = c("Cd44", "Cdx2", "Cdx1", "Tead4", "Alcam"))

FeaturePlot(hybrid, features = c("Mapk14", "Cdkn2a", "Mapk1", "Cdkn1a", "Yap1", "Trp53"))
#note Mapk14 = p38, Cdkn2a = p16, mapk1 = Erk, cdkn1a = p21, 

FeaturePlot(hybrid, features = c("Pdcd1", "Cd274"))
#note Pdcd1 = PD1 and CD274 = PDL1

FeaturePlot(hybrid, features = c("Cd47"))

FeaturePlot(hybrid, features = c("Hur"))

FeaturePlot(hybrid, features = c("Ctsk", "Ctsb", "Ctsl", "Ctsd", "Ctss", "Ctsc", "Ctsg", "Ctse", "Ctsz", "Ctsa", "Ctsh", "Ctsq", "Ctsj"))

FeaturePlot(hybrid, features = c("Gap1"))

```




