---
title: "Hybrid cell identification from published CRC scRNAseq dataset - Tan lab"
author: "Ashley Anderson"
date: "2024-09-05"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# install.packages('Seurat')
# install.packages('Matrix')
# install.packages('ggplot2')
# install.packages('gprofiler2')
# install.packages('ggExtra')
# install.packages('viridis')
# 
# remotes::install_github('chris-mcginnis-ucsf/DoubletFinder')
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("fgsea")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("SingleR")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("celldex")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("scds")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("enrich")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("DOSE")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("clusterProfiler")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("org.Hs.eg.db")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install("hdf5r")
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install('DoubletFinder')
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install('scDblFinder')
# 
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install('SingleCellExperiment', force= TRUE)
# 
```

```{r}
library(Seurat)
options(Seurat.object.assay.version = "v5")
library(patchwork)
library(dplyr)
library(ggplot2)
library(org.Hs.eg.db)
library(clusterProfiler)
library(DOSE)
library(fgsea)
library(enrichplot)
library(gprofiler2)
library(scDblFinder)
library(reticulate)
library(RColorBrewer)
library(ggExtra)
library(cowplot)
library(irlba)
library(Matrix)
library(ggrepel)
library(SingleCellExperiment)
library(scater)
library(scds)
library(future)
library(viridis)
library(hdf5r)
library(DoubletFinder)
library(tibble)
random_seed <- 123456
set.seed(random_seed)
```


```{r}
setwd("F:/CRC-TanLab/data-synapse")
epithelial_cells <- Read10X_h5("Epithelial_Count_matrix.h5")
nonepithelial_cells <- Read10X_h5("NonEpithelial_Count_matrix.h5")

epithelial_cells <- CreateSeuratObject(counts = epithelial_cells, project = "CRCepi.data", min.cells = 3, min.features = 200)
nonepithelial_cells <- CreateSeuratObject(counts = nonepithelial_cells, project = "CRCnonepi.data", min.cells = 3, min.features = 200)


##add in metadata 
epithelial_metadata <- read.csv("Epithelial_metadata.csv")
nonepithelial_metadata <- read.csv("NonEpithelial_metadata.csv")

head(epithelial_metadata)
# Check if the cell IDs match
all(rownames(epithelial_cells@meta.data) %in% epithelial_metadata$cell.ID)  

rownames(epithelial_metadata) <- epithelial_metadata$cell.ID
epithelial_metadata <- epithelial_metadata[rownames(epithelial_cells@meta.data), ]

# Add metadata to Seurat object
epithelial_cells <- AddMetaData(epithelial_cells, epithelial_metadata)
View(epithelial_cells)

head(nonepithelial_metadata)
# Check if the cell IDs match
all(rownames(nonepithelial_cells@meta.data) %in% nonepithelial_metadata$cell.ID)  

rownames(nonepithelial_metadata) <- nonepithelial_metadata$cell.ID
nonepithelial_metadata <- nonepithelial_metadata[rownames(nonepithelial_cells@meta.data), ]

# Add metadata to Seurat object
nonepithelial_cells <- AddMetaData(nonepithelial_cells, nonepithelial_metadata)
View(nonepithelial_cells)


# Visualize QC metrics as a violin plot
metricsplot <- VlnPlot(epithelial_cells, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
metricsplot



#merge your two seurat objects into one 
CRCcombined <- merge(epithelial_cells, y = c(nonepithelial_cells), project = "CRCcombined")
CRCcombined

head(colnames(CRCcombined))
tail(colnames(CRCcombined))
unique(sapply(X = strsplit(colnames(CRCcombined), split = "_"), FUN = "[", 1))
table(CRCcombined$orig.ident)

# Visualize QC metrics as a violin plot
metricsplot1 <- VlnPlot(CRCcombined, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
metricsplot1

##save the object 
##saveRDS(CRCcombined, file = "CRCcombined")
```

```{r}
CRCcombined <- NormalizeData(CRCcombined)
CRCcombined <- FindVariableFeatures(CRCcombined)
CRCcombined <- ScaleData(CRCcombined)
CRCcombined <- RunPCA(CRCcombined)
CRCcombined <- FindNeighbors(CRCcombined, dims = 1:30, reduction = "pca")
CRCcombined <- FindClusters(CRCcombined, resolution = 2, cluster.name = "unintegrated_clusters")
CRCcombined <- RunUMAP(CRCcombined, dims = 1:30, reduction = "pca")
DimPlot(CRCcombined, reduction = "umap", group.by = c("seurat_clusters"))
DimPlot(CRCcombined, reduction = "umap", group.by = c('cell.type'))
DimPlot(CRCcombined, reduction = "umap", group.by = c('dataset'))
```


#evaluate tumor-macrophage hybrids
```{r}

FeaturePlot(CRCcombined, features = c("CD14", "CD163", "CD68", "ITGAM"))

#Assigning hybrid cell identity to all tumor-macrophage positive cells that express EPCAM and/or ECAD and/or panCK and CD14 and/or CD163 and/or CD68 and or Itgam, Where panCK includes all keratins recognized by the AE1/AE3 clone used for IF studies in our lab (Note, KRT6 not found in dataset and KRT10 found to be highly expressed in immune cell populations in this dataset and others in protein atlas and thus was removed, Krt1 not expressed in epithielial population but was increased in mast cel population and thus was removed as well). The AE3 monoclonal antibody recognizes the 65 to 67 triplet, 64, 59, 58, 56, 54 and 52kD proteins also known as cytokeratin 1, 2, 3, 4, 5, 6, 7, and 8 while the AE1 antibody recognizes 56.5, 54', 50, 50', 48, and 40 kDa proteins (also known as CK10, 14, 15, 16 and 19).
FeaturePlot(CRCcombined, features = c('KRT1', 'KRT2', 'KRT3', 'KRT4', 'KRT5', 'KRT7', 'KRT8', 'KRT10', 'KRT14', 'KRT15', 'KRT16', 'KRT19'))
Idents(CRCcombined, WhichCells(object=CRCcombined, expression = (CD14 > 2 | CD68 > 2 | CD163 > 0 | ITGAM > 0) & (EPCAM > 0 | CDH1 > 0 | KRT2 > 0 | KRT3 > 0 | KRT4 > 0 | KRT5 > 0 | KRT7 > 0 | KRT8 > 0 | KRT14 > 0 | KRT15 > 0 | KRT16 > 0 | KRT19 > 0), slot='data')) <-"tumor-macrophage hybrids"
Idents(CRCcombined, WhichCells(object=CRCcombined, expression = (CD14 > 2 | CD68 > 2 | CD163 > 0 | ITGAM > 0) & (EPCAM > 0 | CDH1 > 0 | KRT2 > 0 | KRT3 > 0 | KRT4 > 0 | KRT5 > 0 | KRT7 > 0 | KRT8 > 0 | KRT14 > 0 | KRT15 > 0 | KRT16 > 0 | KRT19 > 0), slot='data', invert = T)) <-"other"

#Add hybrid cell identity to metadata with a column name 'hybridcells'
CRCcombined$hybridcells <- Idents(CRCcombined)
#Count number of hybrid cells 
table(CRCcombined@meta.data$hybridcells)
#Count number of hybrid cells per cell type category
table(CRCcombined@meta.data$hybridcells, CRCcombined@meta.data$cell.type)


DimPlot(CRCcombined, reduction = 'umap', group.by = c('hybridcells'), cols = c("gray", "red"), pt.size = 1)
```


#doublet detection and removal of cells with high doublet scores
```{r}
#identifying doublets and removing cells with high doublet scores from dataset
CRCcombined[['RNA']] <- as(CRCcombined[['RNA']], Class='Assay')
scdoublet <-as.SingleCellExperiment(CRCcombined)
scdoublet = cxds(scdoublet)
scdoublet = bcds(scdoublet)
scdoublet = cxds_bcds_hybrid(scdoublet)

CRCcombined$hybridscore <-scdoublet@colData@listData[["hybrid_score"]]
CRCcombined$cxdsscore <-scdoublet@colData@listData[["cxds_score"]]
CRCcombined$bcdsscore <-scdoublet@colData@listData[["bcds_score"]]

#change hybrid cell identified to hybrid for cell.type but keep metadata of original cell type
CRCcombined$original_cell_type <- CRCcombined$cell.type # Save original cell types
table(CRCcombined@meta.data$cell.type, CRCcombined@meta.data$original_cell_type)

# Identify hybrid cells based on expression criteria
hybrid_cells <- WhichCells(
  object = CRCcombined, 
  expression = (CD14 > 2 | CD68 > 2 | CD163 > 0 | ITGAM > 0) & 
               (EPCAM > 0 | CDH1 > 0 | KRT2 > 0 | KRT3 > 0 | KRT4 > 0 | 
                KRT5 > 0 | KRT7 > 0 | KRT8 > 0 | KRT14 > 0 | KRT15 > 0 | 
                KRT16 > 0 | KRT19 > 0), slot = 'data')

# Add or update cell.type metadata to classify hybrids
CRCcombined$cell.type <- ifelse(colnames(CRCcombined) %in% hybrid_cells,"Hybrid", CRCcombined$cell.type)

# Count number of hybrid cells
table(CRCcombined@meta.data$cell.type)


#violin plot of hybrid scores before filtering for doublet cells.
VlnPlot(CRCcombined, features = c("hybridscore"), pt.size = 0)
VlnPlot(CRCcombined, features = c("hybridscore"), group.by = c('cell.type'), pt.size = 0)

#Create a new Seurat object by keeping cells with hybrid score of less than 0.5
CRCcombined_nd <- subset(CRCcombined, hybridscore < 0.5)

#Count number of hybrid cells after removing doublets
table(CRCcombined_nd@meta.data$hybridcells)
#Count number of hybrid cells per cell type category
table(CRCcombined_nd@meta.data$hybridcells, CRCcombined_nd@meta.data$cell.type)

#violin plot of hybrid scores after filtering for doublet cells 
VlnPlot(CRCcombined_nd, features = c("hybridscore"))
VlnPlot(CRCcombined_nd, features = c("hybridscore"), group.by = c('cell.type'), pt.size = 0)

table(CRCcombined@meta.data$sample.origin)

```


#run scDblFinder to compare to cxds/bcds scores above
```{r}
# Convert full Seurat object to SCE
sce <- as.SingleCellExperiment(CRCcombined)

# Compute density scores on matching cell barcodes
density_scores <- computeDoubletDensity(sce)

# Add scores back
CRCcombined$scDblFinder_density <- density_scores[colnames(CRCcombined)]

# Add cell names from the SCE object
names(density_scores) <- colnames(sce)

# Assign to Seurat object, now the names will match
CRCcombined$scDblFinder_density <- density_scores[colnames(CRCcombined)]

VlnPlot(CRCcombined, features = c("scDblFinder_density"))
VlnPlot(CRCcombined, features = c("scDblFinder_density"), pt.size = 0, group.by = "cell.type")

library(ggbreak)

p <- VlnPlot(
  CRCcombined,
  features = "scDblFinder_density",
  pt.size = 0,
  group.by = "cell.type"
)

# Add a y-axis break at 20
p + scale_y_break(c(5, max(CRCcombined$scDblFinder_density, na.rm = TRUE)), scales = 0.3)

```



#adding patient clinical metadata to evaluate runx1 pathway by site, age, gender, stage, etc
```{r}

clinical_metadata <- read.csv("patient_clinical_information.csv")
head(clinical_metadata)
head(CRCcombined@meta.data)

rownames(clinical_metadata) <- clinical_metadata$patient.ID
colnames(clinical_metadata)

metadata_to_add <- clinical_metadata[, c('Gender', 'Age.at.recruitment', 'MSS.MSI', 'Site', 'Sidedness', 'Group.Stage', 'Stage.TNM', 'KRAS', 'BRAF', 'TP53', 'APC', 'PIK3CA', 'CMS', 'Stage')]

rownames(metadata_to_add) <- clinical_metadata$patient.ID

cell_metadata <- metadata_to_add[CRCcombined_nd@meta.data$patient.ID, ]

CRCcombined_nd <- AddMetaData(object = CRCcombined_nd, metadata = cell_metadata)
head(CRCcombined_nd)
```

```{r}
hyb_only <- subset(
  CRCcombined_nd,
  subset = hybridcells == "tumor-macrophage hybrids" & 
           sample.origin %in% c('Tumor', 'LymphNode', 'Tumor-2', 'Normal')
)

VlnPlot(
  hyb_only,
  features = c('NOC2L', 'UTP4', 'PDCD11', "ZG16"),
  group.by = "sample.origin",
  pt.size = 1,
  cols = c('red3', 'blue3', "green4", 'purple')
)

VlnPlot(
  hyb_only,
  features = c('NOC2L', 'UTP4', 'PDCD11', "ZG16"),
  group.by = "sample.origin",
  pt.size = 0,
  cols = c('red3', 'blue3', "green4", 'purple')
)

hyb_only$hyb_site <- case_when(
  hyb_only$sample.origin %in% c("Tumor", "Tumor-2") ~ "TumorHybrid",
  hyb_only$sample.origin == "LymphNode"             ~ "LNHybrid",
  hyb_only$sample.origin == "Normal"                ~ "NormalHybrid",
  TRUE                                              ~ NA_character_
)

Idents(hyb_only) <- "hyb_site"

# Quick check:
table(hyb_only$hyb_site)

markers_all_hybs <- FindAllMarkers(
  hyb_only,
  only.pos        = TRUE,      # only upregulated in that group
  test.use        = "wilcox",  # default is fine, but you can change
  min.pct         = 0.1,
  logfc.threshold = 0.25
)

# Inspect top markers per group
markers_all_hybs %>% 
  group_by(cluster) %>% 
  top_n(20, wt = avg_log2FC) %>% 
  arrange(cluster, desc(avg_log2FC)) %>% 
  head()
write.csv(markers_all_hybs, file = "allhybridDEGs.csv")

# Tumor hybrids vs LN hybrids
deg_tumor_vs_ln <- FindMarkers(
  hyb_only,
  ident.1        = "TumorHybrid",
  ident.2        = "LNHybrid",
  test.use       = "wilcox",
  min.pct        = 0.1,
  logfc.threshold= 0.25
)
write.csv(deg_tumor_vs_ln, file = "DEG-tumor-v-ln-hybrids.csv")

# Tumor hybrids vs Normal hybrids
deg_tumor_vs_normal <- FindMarkers(
  hyb_only,
  ident.1        = "TumorHybrid",
  ident.2        = "NormalHybrid",
  test.use       = "wilcox",
  min.pct        = 0.1,
  logfc.threshold= 0.25
)
write.csv(deg_tumor_vs_normal, file = "DEG-tumor-v-normal-hybrids.csv")

# LN hybrids vs Normal hybrids
deg_ln_vs_normal <- FindMarkers(
  hyb_only,
  ident.1        = "LNHybrid",
  ident.2        = "NormalHybrid",
  test.use       = "wilcox",
  min.pct        = 0.1,
  logfc.threshold= 0.25
)
write.csv(deg_ln_vs_normal, file = "DEG-ln-v-normal-hybrids.csv")
```


#subsetting data to look only at hybrid cells in tumor/LN samples (keeping normal samples separate)
```{r}

CRCcombined_T <- subset(CRCcombined_nd, subset = sample.origin %in% c('Tumor', 'LymphNode', 'Tumor-2'))

table(CRCcombined_T@meta.data$sample.origin)

#Count number of hybrid cells after removing doublets
table(CRCcombined_T@meta.data$hybridcells)
#Count number of hybrid cells per cell type category
hybridcount_celltype <- table(CRCcombined_T@meta.data$hybridcells, CRCcombined_T@meta.data$cell.type)
write.csv(hybridcount_celltype, file = "hybridcounts_celltype.csv")
#Count number of hybrid cells per CRC stage
table(CRCcombined_T@meta.data$hybridcells, CRCcombined_T@meta.data$Group.Stage)
#Count number of hybrid cells per CRC stage
report <- table(CRCcombined_T@meta.data$hybridcells, CRCcombined_T@meta.data$patient.ID)
write.csv(report, file = "hybrids_patientID.csv")


#used for figure 1 of manuscript 
DimPlot(CRCcombined_T, reduction = 'umap', group.by = c('cell.type'))
DimPlot(CRCcombined_T, reduction = 'umap', group.by = c('hybridcells'), cols = c("gray", "red3"), pt.size = 1)

#code for just looking at hybrid cells in normal samples - for future analyses

# CRCcombined_N <- subset(CRCcombined_nd, subset = sample.origin %in% c('Normal'))
# 
# table(CRCcombined_N@meta.data$sample.origin)
# 
# 
# #Count number of hybrid cells after removing doublets
# table(CRCcombined_N@meta.data$hybridcells)
# #Count number of hybrid cells per cell type category
# table(CRCcombined_N@meta.data$hybridcells, CRCcombined_N@meta.data$cell.type)
# 
# DimPlot(CRCcombined_N, reduction = 'umap', group.by = c('cell.type'))
# DimPlot(CRCcombined_N, reduction = 'umap', group.by = c('hybridcells'), cols = c("gray", "red"), pt.size = 1)


```

#evaluating M1 vs M2 phenotypes in hybrid cells 
```{r}
McDC <- subset(CRCcombined_T, subset = cell.type == "McDC")

#M1
VlnPlot(McDC, features =  c("CD86", "CD40"))
#M2
VlnPlot(McDC, features =  c("CD163", "MRC1"))

```

#evaluating CD163 positive cells
```{r}
# pull metadata
md <- CRCcombined_T@meta.data  

# genes of interest
genes <- c("CD163", "CD14", "CD68", "ITGAM")

# grab expression values for those genes (from counts slot)
expr <- GetAssayData(CRCcombined_T, slot = "counts")[genes, ]

# add to metadata as TRUE/FALSE (whether >0)
for (g in genes) {
  md[[paste0(g, "_pos")]] <- expr[g, ] > 0
}

# make tables
tables <- lapply(genes, function(g) {
  table(md$hybridcells, md[[paste0(g, "_pos")]])
})

names(tables) <- genes
tables
```



#evaluating Runx1 pathway in hybrid cells compared to other cell types
```{r}
library(ggbreak)
VlnPlot(CRCcombined_T, features = c('NOC2L', 'UTP4', 'PDCD11'), group.by = "hybridcells", pt.size = 0, cols = c('tumor-macrophage hybrids' = 'green4', 'other' = 'gray'))

VlnPlot(CRCcombined_T, features = 'NOC2L', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))

VlnPlot(CRCcombined_T, features = 'UTP4', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))

VlnPlot(CRCcombined_T, features = 'PDCD11', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))

VlnPlot(CRCcombined_T, features = c('RUNX1', "SPI1", "THBS1", "LGALS3", "GRN", "PF4"), group.by = "hybridcells", pt.size = 0, cols = c('tumor-macrophage hybrids' = 'green4', 'other' = 'gray'))

VlnPlot(CRCcombined_T, features = "RUNX1", group.by = 'cell.type', pt.size = 0)

VlnPlot(CRCcombined_T, features = 'RUNX1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))
VlnPlot(CRCcombined_T, features = 'THBS1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))
VlnPlot(CRCcombined_T, features = 'SPI1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))
VlnPlot(CRCcombined_T, features = 'GRN', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))
VlnPlot(CRCcombined_T, features = 'PF4', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))
VlnPlot(CRCcombined_T, features = 'LGALS3', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0, cols = c('tumor-macrophage hybrids' = 'gray', 'other' = 'green4'))
VlnPlot(CRCcombined_T, features = 'THBS1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'LGALS3', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'SPI1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'GRN', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
PF4plt <- VlnPlot(CRCcombined_T, features = 'PF4', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
PF4plt + scale_y_break(c(.01, 1), scales = 0.3) 

runx1pathwaygenes <- c("RUNX1", "THBS1", "SPI1", "PF4", "GRN", "LGALS3")

# Ensure identities are set
Idents(CRCcombined_T) <- CRCcombined_T$cell.type

# Initialize a results data frame
results <- data.frame(
  gene = character(),
  t_statistic = numeric(),
  p_value = numeric(),
  mean_Hybrid = numeric(),
  mean_Epithelial = numeric(),
  stringsAsFactors = FALSE
)

#stats for hybrid vs epithelial cells
for (gene in runx1pathwaygenes) {
  expr_df <- data.frame(
    expr = FetchData(CRCcombined_T, gene)[,1],
    cell_type = Idents(CRCcombined_T)
  )
  expr_sub <- expr_df[expr_df$cell_type %in% c("Hybrid", "Epithelial"), ]
  
  # Run t-test, only if both groups have at least two cells
  if (length(unique(expr_sub$cell_type)) == 2 &&
      sum(expr_sub$cell_type == "Hybrid") > 1 &&
      sum(expr_sub$cell_type == "Epithelial") > 1) {
    t_res <- t.test(expr ~ cell_type, data = expr_sub)
    mean_hybrid <- mean(expr_sub$expr[expr_sub$cell_type == "Hybrid"])
    mean_epi <- mean(expr_sub$expr[expr_sub$cell_type == "Epithelial"])
    results <- rbind(results, data.frame(
      gene = gene,
      t_statistic = t_res$statistic,
      p_value = t_res$p.value,
      mean_Hybrid = mean_hybrid,
      mean_Epithelial = mean_epi
    ))
  }
}

print(results)


#stats for hybrid vs McDC cells

# Ensure identities are set
Idents(CRCcombined_T) <- CRCcombined_T$cell.type

# Initialize a results data frame
results <- data.frame(
  gene = character(),
  t_statistic = numeric(),
  p_value = numeric(),
  mean_Hybrid = numeric(),
  mean_McDC = numeric(),
  stringsAsFactors = FALSE
)


for (gene in runx1pathwaygenes) {
  expr_df <- data.frame(
    expr = FetchData(CRCcombined_T, gene)[,1],
    cell_type = Idents(CRCcombined_T)
  )
  expr_sub <- expr_df[expr_df$cell_type %in% c("Hybrid", "McDC"), ]
  
  # Run t-test, only if both groups have at least two cells
  if (length(unique(expr_sub$cell_type)) == 2 &&
      sum(expr_sub$cell_type == "Hybrid") > 1 &&
      sum(expr_sub$cell_type == "McDC") > 1) {
    t_res <- t.test(expr ~ cell_type, data = expr_sub)
    mean_hybrid <- mean(expr_sub$expr[expr_sub$cell_type == "Hybrid"])
    mean_McDC <- mean(expr_sub$expr[expr_sub$cell_type == "McDC"])
    results <- rbind(results, data.frame(
      gene = gene,
      t_statistic = t_res$statistic,
      p_value = t_res$p.value,
      mean_Hybrid = mean_hybrid,
      mean_McDC = mean_McDC
    ))
  }
}

print(results)


VlnPlot(CRCcombined_T, features = 'RUNX1', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'THBS1', group.by = 'sample.origin', split.by = 'hybridcells')
VlnPlot(CRCcombined_T, features = 'LGALS3', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'SPI1', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)


VlnPlot(CRCcombined_T, features = 'CD163', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'CD163', group.by = 'cell.type')
FeaturePlot(CRCcombined_T, features = "CD163")

VlnPlot(CRCcombined_T, features = 'RUNX1', group.by = 'MSS.MSI', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'THBS1', group.by = 'MSS.MSI', split.by = 'hybridcells')
VlnPlot(CRCcombined_T, features = 'LGALS3', group.by = 'MSS.MSI', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T, features = 'SPI1', group.by = 'MSS.MSI', split.by = 'hybridcells', pt.size = 0)


VlnPlot(CRCcombined_T, features = 'RUNX1', group.by = 'Stage', split.by = 'hybridcells', pt.size = 0)


```



```{r}
#hybrid DEG

install.packages('remotes')
remotes::install_github("immunogenomics/presto")
library(presto)

hybrid.markers <- FindMarkers(CRCcombined_T, ident.1 = 'tumor-macrophage hybrids', min.pct = 0.25)
write.csv(hybrid.markers, file = "hybrid_markers.csv")
View(hybrid.markers)

hybridmarkers_df <- hybrid.markers %>% rownames_to_column(var = "gene")

# pick top up and down
top_up   <- hybridmarkers_df %>% 
  filter(avg_log2FC > 0) %>% 
  arrange(desc(avg_log2FC)) %>% 
  slice_head(n = 70) 


top_down <- hybridmarkers_df %>% 
  filter(avg_log2FC < 0) %>% 
  arrange(avg_log2FC) %>% 
  slice_head(n = 20)


genes_to_plot <- c(top_up$gene)


# Scale the DEGs so they will appear in scale data slot, they were not originally scaled in seurat object because it only scales the top HVGs
CRCcombined_T <- ScaleData(
  CRCcombined_T, 
  features = genes_to_plot,    
  verbose  = FALSE
)

#subset for heatmap to show only hybrid, epi, and macrophage cells
CRC.hybrid.epi.mac <- subset(CRCcombined_T, idents = c("Hybrid", "Epithelial", "McDC"))


DoHeatmap(CRC.hybrid.epi.mac, features = genes_to_plot, group.by = "cell.type") 


DotPlot(CRC.hybrid.epi.mac, features = genes_to_plot, group.by = "cell.type") + coord_flip()

DotPlot(CRCcombined_T, features = genes_to_plot, group.by = "cell.type") + coord_flip() +
   theme(axis.text.x = element_text(angle = 45, hjust = 1),
         axis.text.y = element_text(face = "italic"))
  
Idents(CRCcombined_T) <- CRCcombined_T$cell.type  
levels(Idents(CRCcombined_T))        # Active identities
levels(factor(CRCcombined_T$cell.type))  # From metadata
  

#hybrid cells vs epithelial cells 
hybrid.v.epi.markers <- FindMarkers(CRCcombined_T, ident.1 = 'Hybrid', ident.2 = "Epithelial", min.pct = 0.25)
write.csv(hybrid.v.epi.markers, file = "hybrid-v-epi_markers")
View(hybrid.v.epi.markers)


hybrid.v.epi.markers_df <- hybrid.v.epi.markers %>% rownames_to_column(var = "gene")

# pick top up and down
top_up   <- hybrid.v.epi.markers_df %>% 
  filter(avg_log2FC > 0) %>% 
  arrange(desc(avg_log2FC)) %>% 
  slice_head(n = 30) 

genes_to_plot <- c(top_up$gene)


# Scale the DEGs so they will appear in scale data slot, they were not originally scaled in seurat object because it only scales the top HVGs
CRCcombined_T <- ScaleData(
  CRCcombined_T, 
  features = genes_to_plot,    
  verbose  = FALSE
)

DoHeatmap(CRC.hybrid.epi.mac, features = genes_to_plot, group.by = "cell.type")



#hybrid cells vs macrophage cells 
hybrid.v.mac.markers <- FindMarkers(CRCcombined_T, ident.1 = 'Hybrid', ident.2 = "McDC", min.pct = 0.25)
write.csv(hybrid.v.mac.markers, file = "hybrid-v-mac_markers.csv")
View(hybrid.v.mac.markers)


hybrid.v.mac.markers_df <- hybrid.v.mac.markers %>% rownames_to_column(var = "gene")

# pick top up
top_up   <- hybrid.v.mac.markers_df %>% 
  filter(avg_log2FC > 0) %>% 
  arrange(desc(avg_log2FC)) %>% 
  slice_head(n = 30) 

genes_to_plot <- c(top_up$gene)


# Scale the DEGs so they will appear in scale data slot, they were not originally scaled in seurat object because it only scales the top HVGs
CRC.hybrid.epi.mac <- ScaleData(
  CRC.hybrid.epi.mac, 
  features = genes_to_plot,    
  verbose  = FALSE
)


DoHeatmap(CRC.hybrid.epi.mac, features = genes_to_plot, group.by = "cell.type")

```

#GSEA using DEG lists from above
```{r}

library(org.Hs.eg.db)
library(clusterProfiler)
library(DOSE)
library(fgsea)
library(enrichplot)

#GSEA
hybrid.markers$gene <- rownames(hybrid.markers)
hybrid.markers <- hybrid.markers %>% arrange(desc(avg_log2FC))
hybrid_foldchange <- hybrid.markers$avg_log2FC
names(hybrid_foldchange) <- hybrid.markers$gene

head(hybrid_foldchange)

hybrid_foldchange.df <- as.data.frame(hybrid_foldchange)

hybrid_foldchange.df$ENSEMBL <- mapIds(org.Hs.eg.db, keys = rownames(hybrid_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid_enslist <- na.omit(hybrid_foldchange.df)

hybrid_enslist <- hybrid_foldchange.df$hybrid_foldchange
names(hybrid_enslist) <- hybrid_foldchange.df$ENSEMBL
hybrid_enslist

View(hybrid_enslist)

library(future)
future::plan(future::sequential)

gse.hybrid.goBP <- gseGO(hybrid_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = org.Hs.eg.db, eps = 1e-300, maxGSSize = 200)

gse.hybrid.df <- as.data.frame(gse.hybrid.goBP)

write.csv(gse.hybrid.df, file = "gse.hybrid.goBP.csv")



require(DOSE)
dotplothybrid <- dotplot(gse.hybrid.goBP, showCategory=10, split=".sign", label_format = 60) + facet_grid(.~.sign)  +
  scale
  theme(
    axis.text.y = element_text(size = 14),      # Increase Y-axis text (GO terms)
    axis.text.x = element_text(size = 14),      # X-axis (GeneRatio, etc.)
    axis.title = element_text(size = 20),       # Axis titles
    strip.text = element_text(size = 20),       # Facet strip text (".sign")
    legend.text = element_text(size = 14),      # Legend labels
    legend.title = element_text(size = 14),      # Legend title 
    plot.margin = unit(c(1, 1, 1, 1), "cm")  # top, right, bottom, **left**
  )
  

 dotplothybrid


ggsave(filename = "gse.dotplot.hybrid.goBP.tiff", plot = dotplothybrid, width = 7, height = 6, dpi = 600)


library(enrichplot)
gse.hybrid.goBP <- pairwise_termsim(gse.hybrid.goBP)
emapplot(gse.hybrid.goBP, showCategory = 30, cex_label_category = 0.7)


#GSEA for hybrid vs epi
hybrid.v.epi.markers$gene <- rownames(hybrid.v.epi.markers)
hybrid.v.epi.markers <- hybrid.v.epi.markers %>% arrange(desc(avg_log2FC))
hybrid_foldchange <- hybrid.v.epi.markers$avg_log2FC
names(hybrid_foldchange) <- hybrid.v.epi.markers$gene

head(hybrid_foldchange)

hybrid_foldchange.df <- as.data.frame(hybrid_foldchange)

hybrid_foldchange.df <- hybrid_foldchange.df[!duplicated(hybrid_foldchange.df$ENSEMBL), ]
hybrid_enslist <- hybrid_foldchange.df$logFC
names(hybrid_enslist) <- hybrid_foldchange.df$ENSEMBL

hybrid_foldchange.df$ENSEMBL <- mapIds(org.Hs.eg.db, keys = rownames(hybrid_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid_enslist <- na.omit(hybrid_foldchange.df)

hybrid_enslist <- hybrid_foldchange.df$hybrid_foldchange
names(hybrid_enslist) <- hybrid_foldchange.df$ENSEMBL
hybrid_enslist

View(hybrid_enslist)

gse.hybrid.v.epi.goBP <- gseGO(hybrid_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Hs.eg.db", eps = 1e-300, maxGSSize = 200)

gse.hybrid.v.epi.df <- as.data.frame(gse.hybrid.goBP)

write.csv(gse.hybrid.v.epi.df, file = "gse.hybrid.v.epi.goBP.csv")



require(DOSE)
dotplothybrid <- dotplot(gse.hybrid.v.epi.goBP, showCategory=10, split=".sign", label_format = 60) + facet_grid(.~.sign)  +
  scale
  theme(
    axis.text.y = element_text(size = 14),      # Increase Y-axis text (GO terms)
    axis.text.x = element_text(size = 14),      # X-axis (GeneRatio, etc.)
    axis.title = element_text(size = 20),       # Axis titles
    strip.text = element_text(size = 20),       # Facet strip text (".sign")
    legend.text = element_text(size = 14),      # Legend labels
    legend.title = element_text(size = 14),      # Legend title 
    plot.margin = unit(c(1, 1, 1, 1), "cm")  # top, right, bottom, **left**
  )
  

 dotplothybrid


ggsave(filename = "gse.dotplot.hybrid.v.epi.goBP.tiff", plot = dotplothybrid, width = 7, height = 6, dpi = 600)


library(enrichplot)
gse.hybrid.v.epi.goBP <- pairwise_termsim(gse.hybrid.v.epi.goBP)
emapplot(gse.hybrid.v.epi.goBP, showCategory = 30, cex_label_category = 0.7)



#new to test
library(dplyr)
library(org.Hs.eg.db)
library(clusterProfiler)

# Start with your markers object
# Assume: hybrid.v.epi.markers is from FindMarkers, with rownames = gene symbols and avg_log2FC column

# 1. Add gene symbol column and map ENSEMBL IDs
hybrid.v.epi.markers$gene <- rownames(hybrid.v.epi.markers)
hybrid.v.epi.markers$ENSEMBL <- mapIds(
  org.Hs.eg.db,
  keys = hybrid.v.epi.markers$gene,
  keytype = "SYMBOL",
  column = "ENSEMBL"
)

# 2. Remove rows without ENSEMBL IDs
hybrid.v.epi.markers <- hybrid.v.epi.markers[!is.na(hybrid.v.epi.markers$ENSEMBL), ]

# 3. Remove duplicate ENSEMBL IDs (keep highest |logFC|)
hybrid.v.epi.markers <- hybrid.v.epi.markers %>%
  group_by(ENSEMBL) %>%
  slice_max(order_by = abs(avg_log2FC), n = 1) %>%
  ungroup()

# 4. Create named logFC vector for GSEA (names = ENSEMBL)
hybrid_enslist <- hybrid.v.epi.markers$avg_log2FC
names(hybrid_enslist) <- hybrid.v.epi.markers$ENSEMBL

# 5. Sort for GSEA
hybrid_enslist <- sort(hybrid_enslist, decreasing = TRUE)

# 6. Run GSEA
gse.hybrid.v.epi.goBP <- gseGO(
  hybrid_enslist,
  ont = "BP",
  keyType = "ENSEMBL",
  OrgDb = org.Hs.eg.db,
  eps = 1e-300,
  maxGSSize = 200
)

# 7. Save results
gse.hybrid.v.epi.df <- as.data.frame(gse.hybrid.v.epi.goBP)
write.csv(gse.hybrid.v.epi.df, file = "gse.hybrid.v.epi.goBP.csv")




# Find all ENSEMBL IDs that are duplicated
duplicated_ids <- hybrid.v.epi.markers$ENSEMBL[duplicated(hybrid.v.epi.markers$ENSEMBL)]
unique_duplicated_ids <- unique(duplicated_ids)
print(unique_duplicated_ids)
```





#subsetting data to look only at hybrid cells in matched tumor and lymph node patient samples (n=7)
```{r}

CRCcombined_T_LN <- subset(CRCcombined_T, subset = patient.ID %in% c('CRC2783', 'CRC2801', 'CRC2816', 'CRC2817', 'CRC2841', 'SC035', 'SC040'))

table(CRCcombined_T_LN@meta.data$patient.ID)
table(CRCcombined_T_LN@meta.data$sample.origin)

#Count number of hybrid cells after removing doublets
table(CRCcombined_T_LN@meta.data$hybridcells)
#Count number of hybrid cells per cell type category
table(CRCcombined_T_LN@meta.data$hybridcells, CRCcombined_T_LN@meta.data$cell.type)
#Count number of hybrid cells per cell type category
table(CRCcombined_T_LN@meta.data$hybridcells, CRCcombined_T_LN@meta.data$sample.origin)

DimPlot(CRCcombined_T_LN, reduction = 'umap', group.by = c('cell.type'))
DimPlot(CRCcombined_T_LN, reduction = 'umap', group.by = c('hybridcells'), cols = c("gray", "red"), pt.size = 1)

FeaturePlot(CRCcombined_T, features = c('CD68', 'ITGAM', 'PTPRC', 'CD14'))

```


#evaluating Runx1 pathway in hybrid cells compared to other cell types and in tumor vs LN
```{r}
VlnPlot(CRCcombined_T_LN, features = 'RUNX1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T_LN, features = 'THBS1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T_LN, features = 'LGALS3', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T_LN, features = 'SPI1', group.by = 'cell.type', split.by = 'hybridcells', pt.size = 0)

VlnPlot(CRCcombined_T_LN, features = 'RUNX1', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T_LN, features = 'THBS1', group.by = 'sample.origin', split.by = 'hybridcells')
VlnPlot(CRCcombined_T_LN, features = 'LGALS3', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)
VlnPlot(CRCcombined_T_LN, features = 'SPI1', group.by = 'sample.origin', split.by = 'hybridcells', pt.size = 0)

```


#hybrid DEG and GSEA for tumor vs LN hybrids 
```{r}

library(presto)

# Combine tumor and tumor2 into one label
CRCcombined$sample.origin[CRCcombined$sample.origin %in% c("Tumor", "Tumor-2")] <- "tumor"
table(CRCcombined$sample.origin)
CRCcombined_T_LN$sample.origin[CRCcombined_T_LN$sample.origin %in% c("Tumor", "Tumor-2")] <- "tumor"
table(CRCcombined_T_LN$sample.origin)

#subset seurat object to only evaluate hybrid cells between sample types (tumor hybrids vs lymph node hybrids)
CRCcombined_T_LN_hybrids <- subset(CRCcombined_T_LN, subset = cell.type == "Hybrid")

Idents(CRCcombined_T_LN_hybrids) <- "sample.origin"

hybrid.markers.T_LN <- FindMarkers(CRCcombined_T_LN_hybrids, ident.1 = 'tumor', ident.2 = "LymphNode", min.pct = 0.25)
write.csv(hybrid.markers.T_LN, file = "hybrid_markers_T_LN")
View(hybrid.markers.T_LN)


markers_df <- hybrid.markers.T_LN %>% rownames_to_column(var = "gene")


# pick top up and down
top_up   <- markers_df %>% 
  filter(avg_log2FC > 0) %>% 
  arrange(desc(avg_log2FC)) %>% 
  slice_head(n = 20) 


top_down <- markers_df %>% 
  filter(avg_log2FC < 0) %>% 
  arrange(avg_log2FC) %>% 
  slice_head(n = 20)

genes_to_plot <- c(top_up$gene, top_down$gene)

# Scale the DEGs so they will appear in scale data slot, they were not originally scaled in seurat object because it only scales the top HVGs
CRCcombined_T_LN_hybrids <- ScaleData(
  CRCcombined_T_LN_hybrids, 
  features = genes_to_plot,    # your top-up & top-down list
  verbose  = FALSE
)

# DotPlot
DotPlot(
  object   = CRCcombined_T_LN_hybrids,
  features = genes_to_plot,
  group.by = "sample.origin"
) + coord_flip() +
  scale_colour_gradient2(
    low      = "blue",
    mid      = "grey90",
    high     = "red",
    midpoint = 0
  ) 


#heatmap
DoHeatmap(
  CRCcombined_T_LN_hybrids, 
  features = genes_to_plot, 
  slot     = "scale.data") + 
  ggtitle("Top DEGs Tumor vs LN Hybrids") +
  theme(
    plot.title     = element_text(size = 14, face = "bold"),
    axis.text.y    = element_text(size = 14, face = "italic"),        # gene labels
    axis.title     = element_blank(),
    legend.text    = element_text(size = 14),
    legend.title   = element_text(size = 14)
  )



#find shared genes between hybrids in tumor and lymph node
avg_exp <- AggregateExpression(
  CRCcombined_T_LN_hybrids, 
  assays    = "RNA", 
  slot      = "data", 
  group.by  = "sample.origin"
)$RNA

#pick genes with avg expression > 0.5 in both
shared_genes <- rownames(avg_exp)[
  avg_exp[,"tumor"] >  0.5 &
  avg_exp[,"LymphNode"]    >  0.5
]

#save file of shared genes 
write.csv(avg_exp., file = "sharedgeneexpression-tumor-LN-matched-hybrids.csv")



library(org.Hs.eg.db)
library(clusterProfiler)
library(DOSE)
library(fgsea)
library(enrichplot)


#GSEA
hybrid.markers.T_LN$gene <- rownames(hybrid.markers.T_LN)
hybrid.markers.T_LN <- hybrid.markers.T_LN %>% arrange(desc(avg_log2FC))
hybrid_foldchange <- hybrid.markers.T_LN$avg_log2FC
names(hybrid_foldchange) <- hybrid.markers.T_LN$gene

head(hybrid_foldchange)

hybrid_foldchange.df <- as.data.frame(hybrid_foldchange)

hybrid_foldchange.df$ENSEMBL <- mapIds(org.Hs.eg.db, keys = rownames(hybrid_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid_enslist <- na.omit(hybrid_foldchange.df)

hybrid_enslist <- hybrid_foldchange.df$hybrid_foldchange
names(hybrid_enslist) <- hybrid_foldchange.df$ENSEMBL
hybrid_enslist

View(hybrid_enslist)

gse.hybrid.goBP <- gseGO(hybrid_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = org.Hs.eg.db, eps = 1e-300, maxGSSize = 200)

gse.hybrid.df <- as.data.frame(gse.hybrid.goBP)

write.csv(gse.hybrid.df, file = "gse.hybrids-matched-tumor-LN.goBP.csv")

require(DOSE)
dotplothybrid <- dotplot(gse.hybrid.goBP, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplothybrid

ggsave(filename = "gse.dotplot.hybrid.goBP.tiff", plot = dotplothybrid, width = 7, height = 6, dpi = 600)

```




