---
title: "cellfusionscRNAseq"
author: "Ashley Anderson & Konstantin Quietsch"
date: "2024-11-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#loading packages
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
set.seed(77892)
```

```{r}
#Load the datasets
tumor.data <- Read10X(data.dir = "~/Documents/2024-09-17_cellfusionscRNAseq/data/MC38/")

macrophage.data <- Read10X(data.dir = "~/Documents/2024-09-17_cellfusionscRNAseq/data/Macrophage/")

hybrid.m.data <- Read10X(data.dir = "~/Documents/2024-09-17_cellfusionscRNAseq/data/HFF/")

hybrid.t.data <- Read10X(data.dir = "~/Documents/2024-09-17_cellfusionscRNAseq/data/HFP/")



#Initialize the Seurat object with the raw (non-normalized data).
tumor <- CreateSeuratObject(counts = tumor.data, project = "Tumor", min.cells = 3, min.features = 200)
tumor

mac <- CreateSeuratObject(counts = macrophage.data, project = "Macrophage", min.cells = 3, min.features = 200)
mac

hybrid.m <- CreateSeuratObject(counts = hybrid.m.data, project = "Hybrid-M", min.cells = 3, min.features = 200)
hybrid.m

hybrid.t <- CreateSeuratObject(counts = hybrid.t.data, project = "Hybrid-T", min.cells = 3, min.features = 200)
hybrid.t


```

#merge all seurat objects into one 
```{r}
allcells3 <- merge(tumor, y = list(mac, hybrid.m, hybrid.t), add.cell.ids = c('Tumor', 'Mac', 'Hybrid-M', 'Hybrid-T'), project = "allcells3")
allcells3

head(colnames(allcells3))
tail(colnames(allcells3))
unique(sapply(X = strsplit(colnames(allcells3), split = "_"), FUN = "[", 1))
table(allcells3$orig.ident)

saveRDS(allcells3, file = "HybridM-HybridT-scRNAseq-Sept2024-v4-new.RDS")
```


#QC and selecting cells for analysis
```{r}
# The [[ operator can add columns to object metadata. This is a great place to stash QC stats
allcells3[["percent.mt"]] <- PercentageFeatureSet(allcells3, pattern = "^mt-")

# Visualize QC metrics as a violin plot
VlnPlot(allcells3, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)

mt <- VlnPlot(allcells3, "percent.mt")
mt + geom_hline(yintercept = 12.5, color = "red", linetype = "dashed")

nfeature <- VlnPlot(allcells3, "nFeature_RNA")
nfeature + geom_hline(yintercept = 600, color = "red", linetype = "dashed")

allcells3 <- subset(allcells3, subset = nFeature_RNA > 600 & percent.mt < 12.5)
```


```{r}
allcells3 <- NormalizeData(allcells3, normalization.method = "LogNormalize", scale.factor = 10000)
allcells3 <- FindVariableFeatures(allcells3, selection.method = "vst", nfeatures = 2000)
all.genes <- rownames(allcells3)
allcells3 <- ScaleData(allcells3, features = all.genes)
allcells3 <- RunPCA(allcells3, features = VariableFeatures(object = allcells3))
ElbowPlot(allcells3)
allcells3 <- FindNeighbors(allcells3, dims = 1:10)
allcells3 <- FindClusters(allcells3, resolution = 0.5)
head(Idents(allcells3), 5)
allcells3 <- RunUMAP(allcells3, dims = 1:10)
DimPlot(allcells3, reduction = "umap", pt.size=1)

DimPlot(allcells3, group.by = 'orig.ident', pt.size = 1, cols = c("blue4", "red3", "cadetblue3", "gray45", "gold")) +
  theme(element_text(size = 20))

```


#finding differentially expressed features 
```{r}
# install presto for faster implementation of wilcoxon rank sum test 

install.packages('devtools')
devtools::install_github('immunogenomics/presto')

allcells3 <- JoinLayers(allcells3)

# find all markers for each individual cluster 0-11
allcells3.cluster0.markers <- FindMarkers(allcells3, ident.1 = 0, min.pct = 0.25)
head(allcells3.cluster0.markers, n = 20)
write.csv(allcells3.cluster0.markers, file = "cluster0_tumor_markers")
View(allcells3.cluster0.markers)

allcells3.cluster1.markers <- FindMarkers(allcells3, ident.1 = 1, min.pct = 0.25)
write.csv(allcells3.cluster1.markers, file = "cluster1_mac_markers")
head(allcells3.cluster1.markers, n = 20)


allcells3.cluster4.markers <- FindMarkers(allcells3, ident.1 = 2, min.pct = 0.25)
write.csv(allcells3.cluster4.markers, file = "cluster4_hybrid.t_markers")
head(allcells3.cluster4.markers, n = 20)


allcells3.cluster3.markers <- FindMarkers(allcells3, ident.1 = 3, min.pct = 0.25)
write.csv(allcells3.cluster3.markers, file = "cluster3_hybrid.t_markers")
head(allcells3.cluster3.markers, n = 20)


allcells3.cluster4.markers <- FindMarkers(allcells3, ident.1 = 4, min.pct = 0.25)
write.csv(allcells3.cluster4.markers, file = "cluster4_tumor_markers")
head(allcells3.cluster4.markers, n = 20)


allcells3.cluster5.markers <- FindMarkers(allcells3, ident.1 = 5, min.pct = 0.25)
write.csv(allcells3.cluster5.markers, file = "cluster5_hybrid.t_markers")
head(allcells3.cluster5.markers, n = 20)


allcells3.cluster6.markers <- FindMarkers(allcells3, ident.1 = 6, min.pct = 0.25)
write.csv(allcells3.cluster6.markers, file = "cluster6_mac_markers")
head(allcells3.cluster6.markers, n = 20)


allcells3.cluster7.markers <- FindMarkers(allcells3, ident.1 = 7, min.pct = 0.25)
write.csv(allcells3.cluster7.markers, file = "cluster7_hybrid.m_markers")
head(allcells3.cluster7.markers, n = 20)


allcells3.cluster8.markers <- FindMarkers(allcells3, ident.1 = 8, min.pct = 0.25)
write.csv(allcells3.cluster8.markers, file = "cluster8_tumor_markers")
head(allcells3.cluster8.markers, n = 20)


allcells3.cluster9.markers <- FindMarkers(allcells3, ident.1 = 9, min.pct = 0.25)
write.csv(allcells3.cluster9.markers, file = "cluster9_hybrid.m_markers")
head(allcells3.cluster9.markers, n = 20) 

allcells3.cluster10.markers <- FindMarkers(allcells3, ident.1 = 10, min.pct = 0.25)
write.csv(allcells3.cluster9.markers, file = "cluster10_hybrid.m_markers")
head(allcells3.cluster9.markers, n = 20) 

allcells3.cluster11.markers <- FindMarkers(allcells3, ident.1 = 11, min.pct = 0.25)
write.csv(allcells3.cluster9.markers, file = "cluster11_tumor-hybrid.t-mix_markers")
head(allcells3.cluster9.markers, n = 20) 



# find markers for every cluster compared to all remaining cells, report only the positive ones
allcells3.markers <- FindAllMarkers(allcells3, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
allcells3.markers %>%
    group_by(cluster) %>%
    slice_max(n = 2, order_by = avg_log2FC)
write.csv(allcells3.markers, file = "allcells3_pos_markers")

allcells3.markers %>%
    group_by(cluster) %>%
    top_n(n = 10, wt = avg_log2FC) -> top10
DoHeatmap(allcells3, features = top10$gene) + NoLegend()
```


#specific plots for genes of interest
```{r}
FeaturePlot(allcells3, features = c("Runx1", "Thbs1", "Spi1", "Grn", "Pf4", "Lgals3"))

FeaturePlot(allcells3, features = c('Cd163'))

FeaturePlot(allcells3, features = c('B2m', 'S100a11','Tmsb10'))

FeaturePlot(allcells3, features = c('Ptprc'))
FeaturePlot(allcells3, features = c('Adgre1'))
FeaturePlot(allcells3, features = c('Cd14'))

FeaturePlot(allcells3, features = c('Tob2'))

#M1
FeaturePlot(allcells3, features = c("Fcgr3", "Cd86", "Nos2", "Il6"))
VlnPlot(allcells3, features = c('Fcgr3'), group.by = 'orig.ident')
VlnPlot(allcells3, features = c('Cd86'), group.by = 'orig.ident')
VlnPlot(allcells3, features = c('Nos2'), group.by = 'orig.ident')
VlnPlot(allcells3, features = c('Il6'), group.by = 'orig.ident')

#M2
FeaturePlot(allcells3, features = c("Retnlb", "Mrc1", "Arg1", "Ccl2", "Clec7a", "Chil3", "Chil4"))
VlnPlot(allcells3, features = c("Retnlb"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Mrc1"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Arg1"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Ccl2"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Clec7a"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Chil3"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Chil4"), group.by = 'orig.ident')


VlnPlot(allcells3, features = c("Lgals3"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Runx1"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Spi1"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Grn"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Pf4"), group.by = 'orig.ident')
VlnPlot(allcells3, features = c("Thbs1"), group.by = 'orig.ident')


FeaturePlot(allcells3, features = c('Runx1'), pt.size = 1, alpha = 0.7)
FeaturePlot(allcells3, features = c('Thbs1'), pt.size = 1)

FeaturePlot(allcells3, features = c('Mmp8'))
FeaturePlot(allcells3, features = c('Mmp9'))
FeaturePlot(allcells3, features = c('Mmp10'))
FeaturePlot(allcells3, features = c('Col1a1'))
FeaturePlot(allcells3, features = c('Fn1'))
FeaturePlot(allcells3, features = c('Vim'))
FeaturePlot(allcells3, features = c('Elk3'))
FeaturePlot(allcells3, features = c('Ptgs2'))
FeaturePlot(allcells3, features = c('Tgfb1'))


FeaturePlot(allcells3, features = c("Ki67", "Pcna"))

FeaturePlot(allcells3, features = c("Muc1", "Muc5ac", "Epcam", "Egfr", "Cdh1", "Krt8", "Krt5", "Krt14", "Krt19", "Krt9", "Krt7", "Krt18", "Krt10"), ncol = 2)

FeaturePlot(allcells3, features = c("Satb", "Cldn1", "Muc4"))

FeaturePlot(allcells3, features = c("Vim", "Snai1", "Tgfb1", "Cdh2", "Zeb1", "Zeb2", "Timp1", "Fn1", "Acta2"), ncol=2)

FeaturePlot(allcells3, features = c("Ptprc", "Cd68", "Cd14","Fcgr3", "Adgre1", "Itgax" , "Csf1r"))
#note Fcgr3 = CD16 and Ptprc = CD45

FeaturePlot(allcells3, features = c("Cd44", "Cdx2", "Cdx1", "Tead4", "Alcam"))
FeaturePlot(allcells3, features = c("Mapk14", "Cdkn2a", "Mapk1", "Cdkn1a", "Yap1", "Trp53"))
#note Mapk14 = p38, Cdkn2a = p16, mapk1 = Erk, cdkn1a = p21, 

FeaturePlot(allcells3, features = c("Pdcd1", "Cd274"))
#note Pdcd1 = PD1 and CD274 = PDL1

FeaturePlot(allcells3, features = c("Cd47"))

FeaturePlot(allcells3, features = c("Hur"))

FeaturePlot(allcells3, features = c("Ctsk", "Ctsb", "Ctsl", "Ctsd", "Ctss", "Ctsc", "Ctsg", "Ctse", "Ctsz", "Ctsa", "Ctsh", "Ctsq", "Ctsj"))

FeaturePlot(allcells3, features = c("Gap1"))

FeaturePlot(allcells3, features = c("Ccl3", "Il1b", "Cxcr4", "Mmp1", "B2m"))

FeaturePlot(allcells3, features = c("Spp1", "Cxcl9", "Il1a"))

FeaturePlot(allcells3, features = c("Cstb"))

FeaturePlot(allcells3, features = c("Mmp1", "Mmp2", "Mmp3", "Mmp4", "Mmp5", "Mmp6", "Mmp7", "Mmp8", "Mmp9", "Mmp10", "Mmp11", "Mmp12", "Mmp13", "Mmp14"))

##look at genes that cleave MMPs 


FeaturePlot(allcells3, features = c("Tm4sf1", "Twist1", "Steap1", "Cdh18", "Dclk1", "Tead2", "Ddr2"))

#hybrid gene signature feature plots 
FeaturePlot(allcells3, features = c("Crabp1", "Arg1", "Ctla2a", "Thbs1", "Ccr1", "Pf4", "Ccl6"))

FeaturePlot(allcells3, features = c("Sorbs2"))

FeaturePlot(allcells3, features = c("Spp1"))


```


#GSEA
```{r}
library(org.Mm.eg.db)
library(clusterProfiler)
library(DOSE)
library(fgsea)
library(enrichplot)
```

#cluster 0
```{r}
allcells3.cluster0.markers$gene <- rownames(allcells3.cluster0.markers)
allcells3.cluster0.markers <- allcells3.cluster0.markers %>% arrange(desc(avg_log2FC))
cluster0_foldchange <- allcells3.cluster0.markers$avg_log2FC
names(cluster0_foldchange) <- allcells3.cluster0.markers$gene

head(cluster0_foldchange)

cluster0_foldchange.df <- as.data.frame(cluster0_foldchange)

cluster0_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster0_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster0_enslist <- na.omit(cluster0_foldchange.df)

cluster0_enslist <- cluster0_foldchange.df$cluster0_foldchange
names(cluster0_enslist) <- cluster0_foldchange.df$ENSEMBL
cluster0_enslist

View(cluster0_enslist)

gse.cluster0 <- gseGO(cluster0_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster0.df <- as.data.frame(gse.cluster0)

write.csv(gse.cluster0.df, file = "gse.cluster0.csv")

require(DOSE)
dotplot0 <- dotplot(gse.cluster0, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot0

ggsave(filename = "gse.dotplot.cluster0.tiff", plot = dotplot0, width = 7, height = 6, dpi = 600)
```

#cluster 1
```{r}
allcells3.cluster1.markers$gene <- rownames(allcells3.cluster1.markers)
allcells3.cluster1.markers <- allcells3.cluster1.markers %>% arrange(desc(avg_log2FC))
cluster1_foldchange <- allcells3.cluster1.markers$avg_log2FC
names(cluster1_foldchange) <- allcells3.cluster1.markers$gene

head(cluster1_foldchange)

cluster1_foldchange.df <- as.data.frame(cluster1_foldchange)

cluster1_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster1_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster1_enslist <- na.omit(cluster1_foldchange.df)

cluster1_enslist <- cluster1_foldchange.df$cluster1_foldchange
names(cluster1_enslist) <- cluster1_foldchange.df$ENSEMBL
cluster1_enslist

View(cluster1_enslist)

gse.cluster1 <- gseGO(cluster1_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster1.df <- as.data.frame(gse.cluster1)

write.csv(gse.cluster1.df, file = "gse.cluster1.csv")

require(DOSE)
dotplot1 <- dotplot(gse.cluster1, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot1

ggsave(filename = "gse.dotplot.cluster1.tiff", plot = dotplot1, width = 7, height = 6, dpi = 600)
```

#cluster2 
```{r}
allcells3.cluster2.markers$gene <- rownames(allcells3.cluster2.markers)
allcells3.cluster2.markers <- allcells3.cluster2.markers %>% arrange(desc(avg_log2FC))
cluster2_foldchange <- allcells3.cluster2.markers$avg_log2FC
names(cluster2_foldchange) <- allcells3.cluster2.markers$gene

head(cluster2_foldchange)

cluster2_foldchange.df <- as.data.frame(cluster2_foldchange)

cluster2_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster2_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster2_enslist <- na.omit(cluster2_foldchange.df)

cluster2_enslist <- cluster2_foldchange.df$cluster2_foldchange
names(cluster2_enslist) <- cluster2_foldchange.df$ENSEMBL
cluster2_enslist

View(cluster2_enslist)

gse.cluster2 <- gseGO(cluster2_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster2.df <- as.data.frame(gse.cluster2)

write.csv(gse.cluster2.df, file = "gse.cluster2.csv")

require(DOSE)
dotplot2 <- dotplot(gse.cluster2, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot2

ggsave(filename = "gse.dotplot.cluster2.tiff", plot = dotplot2, width = 7, height = 6, dpi = 600)
```


#cluster3 
```{r}
allcells3.cluster3.markers$gene <- rownames(allcells3.cluster3.markers)
allcells3.cluster3.markers <- allcells3.cluster3.markers %>% arrange(desc(avg_log2FC))
cluster3_foldchange <- allcells3.cluster3.markers$avg_log2FC
names(cluster3_foldchange) <- allcells3.cluster3.markers$gene

head(cluster3_foldchange)

cluster3_foldchange.df <- as.data.frame(cluster3_foldchange)

cluster3_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster3_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster3_enslist <- na.omit(cluster3_foldchange.df)

cluster3_enslist <- cluster3_foldchange.df$cluster3_foldchange
names(cluster3_enslist) <- cluster3_foldchange.df$ENSEMBL
cluster3_enslist

View(cluster3_enslist)

gse.cluster3 <- gseGO(cluster3_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster3.df <- as.data.frame(gse.cluster3)

write.csv(gse.cluster3.df, file = "gse.cluster3.csv")

require(DOSE)
dotplot3 <- dotplot(gse.cluster3, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot3

ggsave(filename = "gse.dotplot.cluster3.tiff", plot = dotplot3, width = 7, height = 6, dpi = 600)
```

#cluster4 
```{r}
allcells3.cluster4.markers$gene <- rownames(allcells3.cluster4.markers)
allcells3.cluster4.markers <- allcells3.cluster4.markers %>% arrange(desc(avg_log2FC))
cluster4_foldchange <- allcells3.cluster4.markers$avg_log2FC
names(cluster4_foldchange) <- allcells3.cluster4.markers$gene

head(cluster4_foldchange)

cluster4_foldchange.df <- as.data.frame(cluster4_foldchange)

cluster4_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster4_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster4_enslist <- na.omit(cluster4_foldchange.df)

cluster4_enslist <- cluster4_foldchange.df$cluster4_foldchange
names(cluster4_enslist) <- cluster4_foldchange.df$ENSEMBL
cluster4_enslist

View(cluster4_enslist)

gse.cluster4 <- gseGO(cluster4_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster4.df <- as.data.frame(gse.cluster4)

write.csv(gse.cluster4.df, file = "gse.cluster4.csv")

require(DOSE)
dotplot4 <- dotplot(gse.cluster4, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot4

ggsave(filename = "gse.dotplot.cluster4.tiff", plot = dotplot4, width = 7, height = 6, dpi = 600)
```

#cluster5 
```{r}
allcells3.cluster5.markers$gene <- rownames(allcells3.cluster5.markers)
allcells3.cluster5.markers <- allcells3.cluster5.markers %>% arrange(desc(avg_log2FC))
cluster5_foldchange <- allcells3.cluster5.markers$avg_log2FC
names(cluster5_foldchange) <- allcells3.cluster5.markers$gene

head(cluster5_foldchange)

cluster5_foldchange.df <- as.data.frame(cluster5_foldchange)

cluster5_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster5_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster5_enslist <- na.omit(cluster5_foldchange.df)

cluster5_enslist <- cluster5_foldchange.df$cluster5_foldchange
names(cluster5_enslist) <- cluster5_foldchange.df$ENSEMBL
cluster5_enslist

View(cluster5_enslist)

gse.cluster5 <- gseGO(cluster5_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster5.df <- as.data.frame(gse.cluster5)

write.csv(gse.cluster5.df, file = "gse.cluster5.csv")

require(DOSE)
dotplot5 <- dotplot(gse.cluster5, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot5

ggsave(filename = "gse.dotplot.cluster5.tiff", plot = dotplot5, width = 7, height = 6, dpi = 600)
```

#cluster6 
```{r}
allcells3.cluster6.markers$gene <- rownames(allcells3.cluster6.markers)
allcells3.cluster6.markers <- allcells3.cluster6.markers %>% arrange(desc(avg_log2FC))
cluster6_foldchange <- allcells3.cluster6.markers$avg_log2FC
names(cluster6_foldchange) <- allcells3.cluster6.markers$gene

head(cluster6_foldchange)

cluster6_foldchange.df <- as.data.frame(cluster6_foldchange)

cluster6_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster6_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster6_enslist <- na.omit(cluster6_foldchange.df)

cluster6_enslist <- cluster6_foldchange.df$cluster6_foldchange
names(cluster6_enslist) <- cluster6_foldchange.df$ENSEMBL
cluster6_enslist

View(cluster6_enslist)

gse.cluster6 <- gseGO(cluster6_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster6.df <- as.data.frame(gse.cluster6)

write.csv(gse.cluster6.df, file = "gse.cluster6.csv")

require(DOSE)
dotplot6 <- dotplot(gse.cluster6, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot6

ggsave(filename = "gse.dotplot.cluster6.tiff", plot = dotplot6, width = 7, height = 6, dpi = 600)
```

#cluster7 
```{r}
allcells3.cluster7.markers$gene <- rownames(allcells3.cluster7.markers)
allcells3.cluster7.markers <- allcells3.cluster7.markers %>% arrange(desc(avg_log2FC))
cluster7_foldchange <- allcells3.cluster7.markers$avg_log2FC
names(cluster7_foldchange) <- allcells3.cluster7.markers$gene

head(cluster7_foldchange)

cluster7_foldchange.df <- as.data.frame(cluster7_foldchange)

cluster7_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster7_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster7_enslist <- na.omit(cluster7_foldchange.df)

cluster7_enslist <- cluster7_foldchange.df$cluster7_foldchange
names(cluster7_enslist) <- cluster7_foldchange.df$ENSEMBL
cluster7_enslist

View(cluster7_enslist)

gse.cluster7 <- gseGO(cluster7_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster7.df <- as.data.frame(gse.cluster7)

write.csv(gse.cluster7.df, file = "gse.cluster7.csv")

require(DOSE)
dotplot7 <- dotplot(gse.cluster7, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot7

ggsave(filename = "gse.dotplot.cluster7.tiff", plot = dotplot7, width = 7, height = 6, dpi = 600)
```

#cluster8 
```{r}
allcells3.cluster8.markers$gene <- rownames(allcells3.cluster8.markers)
allcells3.cluster8.markers <- allcells3.cluster8.markers %>% arrange(desc(avg_log2FC))
cluster8_foldchange <- allcells3.cluster8.markers$avg_log2FC
names(cluster8_foldchange) <- allcells3.cluster8.markers$gene

head(cluster8_foldchange)

cluster8_foldchange.df <- as.data.frame(cluster8_foldchange)

cluster8_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster8_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster8_enslist <- na.omit(cluster8_foldchange.df)

cluster8_enslist <- cluster8_foldchange.df$cluster8_foldchange
names(cluster8_enslist) <- cluster8_foldchange.df$ENSEMBL
cluster8_enslist

View(cluster8_enslist)

gse.cluster8 <- gseGO(cluster8_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster8.df <- as.data.frame(gse.cluster8)

write.csv(gse.cluster8.df, file = "gse.cluster8.csv")

require(DOSE)
dotplot8 <- dotplot(gse.cluster8, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot8

ggsave(filename = "gse.dotplot.cluster8.tiff", plot = dotplot8, width = 7, height = 6, dpi = 600)
```

#cluster9 
```{r}
allcells3.cluster9.markers$gene <- rownames(allcells3.cluster9.markers)
allcells3.cluster9.markers <- allcells3.cluster9.markers %>% arrange(desc(avg_log2FC))
cluster9_foldchange <- allcells3.cluster9.markers$avg_log2FC
names(cluster9_foldchange) <- allcells3.cluster9.markers$gene

head(cluster9_foldchange)

cluster9_foldchange.df <- as.data.frame(cluster9_foldchange)

cluster9_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster9_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster9_enslist <- na.omit(cluster9_foldchange.df)

cluster9_enslist <- cluster9_foldchange.df$cluster9_foldchange
names(cluster9_enslist) <- cluster9_foldchange.df$ENSEMBL
cluster9_enslist

View(cluster9_enslist)

gse.cluster9 <- gseGO(cluster9_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster9.df <- as.data.frame(gse.cluster9)

write.csv(gse.cluster9.df, file = "gse.cluster9.csv")

require(DOSE)
dotplot9 <- dotplot(gse.cluster9, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot9

ggsave(filename = "gse.dotplot.cluster9.tiff", plot = dotplot9, width = 7, height = 6, dpi = 600)
```


#cluster10 
```{r}
allcells3.cluster10.markers$gene <- rownames(allcells3.cluster10.markers)
allcells3.cluster10.markers <- allcells3.cluster10.markers %>% arrange(desc(avg_log2FC))
cluster10_foldchange <- allcells3.cluster10.markers$avg_log2FC
names(cluster10_foldchange) <- allcells3.cluster10.markers$gene

head(cluster10_foldchange)

cluster10_foldchange.df <- as.data.frame(cluster10_foldchange)

cluster10_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster10_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster10_enslist <- na.omit(cluster10_foldchange.df)

cluster10_enslist <- cluster10_foldchange.df$cluster10_foldchange
names(cluster10_enslist) <- cluster10_foldchange.df$ENSEMBL
cluster10_enslist

View(cluster10_enslist)

gse.cluster10 <- gseGO(cluster10_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster10.df <- as.data.frame(gse.cluster10)

write.csv(gse.cluster10.df, file = "gse.cluster10.csv")

require(DOSE)
dotplot10 <- dotplot(gse.cluster10, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot10

ggsave(filename = "gse.dotplot.cluster10.tiff", plot = dotplot10, width = 7, height = 6, dpi = 600)
```

#cluster11 
```{r}
allcells3.cluster11.markers$gene <- rownames(allcells3.cluster11.markers)
allcells3.cluster11.markers <- allcells3.cluster11.markers %>% arrange(desc(avg_log2FC))
cluster11_foldchange <- allcells3.cluster11.markers$avg_log2FC
names(cluster11_foldchange) <- allcells3.cluster11.markers$gene

head(cluster11_foldchange)

cluster11_foldchange.df <- as.data.frame(cluster11_foldchange)

cluster11_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(cluster11_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
cluster11_enslist <- na.omit(cluster11_foldchange.df)

cluster11_enslist <- cluster11_foldchange.df$cluster11_foldchange
names(cluster11_enslist) <- cluster11_foldchange.df$ENSEMBL
cluster11_enslist

View(cluster11_enslist)

gse.cluster11 <- gseGO(cluster11_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.cluster11.df <- as.data.frame(gse.cluster11)

write.csv(gse.cluster11.df, file = "gse.cluster11.csv")

require(DOSE)
dotplot11 <- dotplot(gse.cluster11, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplot11

ggsave(filename = "gse.dotplot.cluster11.tiff", plot = dotplot11, width = 7, height = 6, dpi = 600)
```


#assign cluster IDs based on cell type 
```{r}
new.cluster.ids <- c('tumor', 'mac', 'hybrid.t', 'hybrid.t', 'tumor', 'hybrid.t', 'mac', 'hybrid.m', 'tumor', 'hybrid.m', 'hybrid.m', 'mixed')
names(new.cluster.ids) <- levels(allcells3)
allcells3_cellID <- RenameIdents(allcells3, new.cluster.ids)
DimPlot(allcells3_cellID, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

#run DEG and GSEA again with cluster id's relabled for each cell type 
```{r}

allcells3_cellID <- JoinLayers(allcells3_cellID)

# find all markers for each cell type 
allcells3.mac.markers <- FindMarkers(allcells3_cellID, ident.1 = 'mac', min.pct = 0.25)
head(allcells3.mac.markers, n = 20)
write.csv(allcells3.mac.markers, file = "mac_markers")
View(allcells3.mac.markers)

allcells3.tumor.markers <- FindMarkers(allcells3_cellID, ident.1 = 'tumor', min.pct = 0.25)
head(allcells3.tumor.markers, n = 20)
write.csv(allcells3.tumor.markers, file = "tumor_markers")
View(allcells3.tumor.markers)

allcells3.hybrid.t.markers <- FindMarkers(allcells3_cellID, ident.1 = 'hybrid.t', min.pct = 0.25)
head(allcells3.hybrid.t.markers, n = 20)
write.csv(allcells3.hybrid.t.markers, file = "hybrid.t_markers")
View(allcells3.hybrid.t.markers)

allcells3.hybrid.m.markers <- FindMarkers(allcells3_cellID, ident.1 = 'hybrid.m', min.pct = 0.25)
head(allcells3.hybrid.m.markers, n = 20)
write.csv(allcells3.hybrid.m.markers, file = "hybrid.m_markers")
View(allcells3.hybrid.m.markers)
```


#mac GSEA
```{r}
allcells3.mac.markers$gene <- rownames(allcells3.mac.markers)
allcells3.mac.markers <- allcells3.mac.markers %>% arrange(desc(avg_log2FC))
mac_foldchange <- allcells3.mac.markers$avg_log2FC
names(mac_foldchange) <- allcells3.mac.markers$gene

head(mac_foldchange)

mac_foldchange.df <- as.data.frame(mac_foldchange)

mac_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(mac_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
mac_enslist <- na.omit(mac_foldchange.df)

mac_enslist <- mac_foldchange.df$mac_foldchange
names(mac_enslist) <- mac_foldchange.df$ENSEMBL
mac_enslist

View(mac_enslist)

gse.mac <- gseGO(mac_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.mac.df <- as.data.frame(gse.mac)

write.csv(gse.mac.df, file = "gse.mac.csv")

require(DOSE)
dotplotmac <- dotplot(gse.mac, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplotmac

ggsave(filename = "gse.dotplot.mac.tiff", plot = dotplotmac, width = 7, height = 6, dpi = 600)
```

#tumor GSEA
```{r}
allcells3.tumor.markers$gene <- rownames(allcells3.tumor.markers)
allcells3.tumor.markers <- allcells3.tumor.markers %>% arrange(desc(avg_log2FC))
tumor_foldchange <- allcells3.tumor.markers$avg_log2FC
names(tumor_foldchange) <- allcells3.tumor.markers$gene

head(tumor_foldchange)

tumor_foldchange.df <- as.data.frame(tumor_foldchange)

tumor_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(tumor_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
tumor_enslist <- na.omit(tumor_foldchange.df)

tumor_enslist <- tumor_foldchange.df$tumor_foldchange
names(tumor_enslist) <- tumor_foldchange.df$ENSEMBL
tumor_enslist

View(tumor_enslist)

gse.tumor <- gseGO(tumor_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.tumor.df <- as.data.frame(gse.tumor)

write.csv(gse.tumor.df, file = "gse.tumor.csv")

require(DOSE)
dotplottumor <- dotplot(gse.tumor, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplottumor

ggsave(filename = "gse.dotplot.tumor.tiff", plot = dotplottumor, width = 7, height = 6, dpi = 600)
```


#hybrid.t GSEA
```{r}
allcells3.hybrid.t.markers$gene <- rownames(allcells3.hybrid.t.markers)
allcells3.hybrid.t.markers <- allcells3.hybrid.t.markers %>% arrange(desc(avg_log2FC))
hybrid.t_foldchange <- allcells3.hybrid.t.markers$avg_log2FC
names(hybrid.t_foldchange) <- allcells3.hybrid.t.markers$gene

head(hybrid.t_foldchange)

hybrid.t_foldchange.df <- as.data.frame(hybrid.t_foldchange)

hybrid.t_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(hybrid.t_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid.t_enslist <- na.omit(hybrid.t_foldchange.df)

hybrid.t_enslist <- hybrid.t_foldchange.df$hybrid.t_foldchange
names(hybrid.t_enslist) <- hybrid.t_foldchange.df$ENSEMBL
hybrid.t_enslist

View(hybrid.t_enslist)

gse.hybrid.t <- gseGO(hybrid.t_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300, maxGSSize = 200)

gse.hybrid.t.df <- as.data.frame(gse.hybrid.t)

write.csv(gse.hybrid.t.df, file = "gse.hybrid.t.csv")

require(DOSE)
dotplothybrid.t <- dotplot(gse.hybrid.t, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplothybrid.t

ggsave(filename = "gse.dotplot.hybrid.t.tiff", plot = dotplothybrid.t, width = 7, height = 6, dpi = 600)
```

#hybrid.m GSEA
```{r}
allcells3.hybrid.m.markers$gene <- rownames(allcells3.hybrid.m.markers)
allcells3.hybrid.m.markers <- allcells3.hybrid.m.markers %>% arrange(desc(avg_log2FC))
hybrid.m_foldchange <- allcells3.hybrid.m.markers$avg_log2FC
names(hybrid.m_foldchange) <- allcells3.hybrid.m.markers$gene

head(hybrid.m_foldchange)

hybrid.m_foldchange.df <- as.data.frame(hybrid.m_foldchange)

hybrid.m_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(hybrid.m_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid.m_enslist <- na.omit(hybrid.m_foldchange.df)

hybrid.m_enslist <- hybrid.m_foldchange.df$hybrid.m_foldchange
names(hybrid.m_enslist) <- hybrid.m_foldchange.df$ENSEMBL
hybrid.m_enslist

View(hybrid.m_enslist)

gse.hybrid.m <- gseGO(hybrid.m_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300, maxGSSize = 200)

gse.hybrid.m.df <- as.data.frame(gse.hybrid.m)

write.csv(gse.hybrid.m.df, file = "gse.hybrid.m.csv")

require(DOSE)
dotplothybrid.m <- dotplot(gse.hybrid.m, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplothybrid.m

ggsave(filename = "gse.dotplot.hybrid.m.tiff", plot = dotplothybrid.m, width = 7, height = 6, dpi = 600)
```

```{r}
hybrid.m_vs_hybrid.t_markers <- FindMarkers(allcells3_cellID, ident.1 = 'hybrid.m', ident.2 = 'hybrid.t', min.pct = 0.25)
head(hybrid.m_vs_hybrid.t_markers, n = 20)
#write.csv(hybrid.m_vs_hybrid.t_markers, file = "hybrid.m_v_t_markers")
#View(hybrid.m_vs_hybrid.t_markers)

hybrid.m_vs_hybrid.t_markers$gene <- rownames(hybrid.m_vs_hybrid.t_markers)
hybrid.m_vs_hybrid.t_markers <- hybrid.m_vs_hybrid.t_markers %>% arrange(desc(avg_log2FC))
hybrid.m_v_t_foldchange <- hybrid.m_vs_hybrid.t_markers$avg_log2FC
names(hybrid.m_v_t_foldchange) <- hybrid.m_vs_hybrid.t_markers$gene

head(hybrid.m_v_t_foldchange)

hybrid.m_v_t_foldchange.df <- as.data.frame(hybrid.m_v_t_foldchange)

hybrid.m_v_t_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(hybrid.m_v_t_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid.m_v_t_enslist <- na.omit(hybrid.m_v_t_foldchange.df)

hybrid.m_v_t_enslist <- hybrid.m_v_t_foldchange.df$hybrid.m_v_t_foldchange
names(hybrid.m_v_t_enslist) <- hybrid.m_v_t_foldchange.df$ENSEMBL
hybrid.m_v_t_enslist

#View(hybrid.m_v_t_enslist)

gse.hybrid.m_v_t <- gseGO(hybrid.m_v_t_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300, maxGSSize = 200)

gse.hybrid.m_v_t.df <- as.data.frame(gse.hybrid.m_v_t)

#write.csv(gse.hybrid.m_v_t.df, file = "gse.hybrid.m_v_t.csv")

require(DOSE)
dotplothybrid.m_v_t <- dotplot(gse.hybrid.m_v_t, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplothybrid.m_v_t

#ggsave(filename = "gse.dotplot.hybrid.m_v_t.tiff", plot = dotplothybrid.m, width = 7, height = 6, dpi = 600)

```

```{r}
library(ggplot2)
library(ggrepel)

hybrid.m_vs_hybrid.t_markers

# Add a column for -log10(p-value)
hybrid.m_vs_hybrid.t_markers$log10_pval <- -log10(hybrid.m_vs_hybrid.t_markers$p_val_adj + 1e-300) # Avoid log(0)

# Add a column to categorize genes (significant up/down or not)
hybrid.m_vs_hybrid.t_markers$category <- ifelse(
  hybrid.m_vs_hybrid.t_markers$p_val_adj < 0.000000000000000000000001 & hybrid.m_vs_hybrid.t_markers$avg_log2FC > 4.0, "Upregulated",
  ifelse(hybrid.m_vs_hybrid.t_markers$p_val_adj < 0.000000000000000000000001 & hybrid.m_vs_hybrid.t_markers$avg_log2FC < -4.0, "Downregulated", "Not Significant")
)



ggplot(hybrid.m_vs_hybrid.t_markers, aes(x = avg_log2FC, y = -log10_pval, color = category, label = ifelse(category != "Not Significant", rownames(hybrid.m_vs_hybrid.t_markers), ""))) +
  geom_point(alpha = 0.7) +
  geom_text_repel(size = 3) +
  scale_color_manual(values = c("Upregulated" = "red", "Downregulated" = "blue", "Not Significant" = "grey")) +
  theme_minimal() +
  labs(
    title = "Volcano Plot of DEGs",
    x = "Log2 Fold Change",
    y = "-Log10 Adjusted P-Value"
  ) +
  theme(
    legend.position = "top",
    text = element_text(size = 12)
  )

```



```{r}
library(tibble)

# Add rownames as a column
hybrid.m_vs_hybrid.t_markers$gene <- rownames(hybrid.m_vs_hybrid.t_markers)
View(hybrid.m_vs_hybrid.t_markers)

# Top 10 DEGs for hybrid.m (positive log2FC indicates higher in hybrid.m)
top_genes_hybrid_m <- hybrid.m_vs_hybrid.t_markers %>%
  filter(avg_log2FC > 0) %>%
  arrange(p_val_adj) %>%
  head(10) %>%
  pull(gene)

# Top 10 DEGs for hybrid.t (negative log2FC indicates higher in hybrid.t)
top_genes_hybrid_t <- hybrid.m_vs_hybrid.t_markers %>%
  filter(avg_log2FC < 0) %>%
  arrange(p_val_adj) %>%
  head(10) %>%
  pull(gene)

# Combine both sets of genes
top_genes <- unique(c(top_genes_hybrid_f, top_genes_hybrid_p))

# Dot plot for combined top genes
DotPlot(
  allcells3_cellID,
  features = top_genes,
  group.by = "orig.ident", 
  dot.scale = 10
) +
  scale_color_gradient(low = "cyan3", high = "magenta3") +
  theme_minimal() +
  labs(title = "Dot Plot of Top DEGs from hybrid.m and hybrid.t") +
  coord_flip() +
   theme(
    axis.text.y = element_text(size = 16, color = "black"), # Smaller text for y-axis
    axis.text.x = element_text(size = 16, color = "black"), # Adjust x-axis text
    axis.ticks.length.y = unit(0.00001, "cm"), # Reduce space between rows
    axis.ticks.length.x = unit(0.00001, "mm"), # Reduce space between columns
    axis.ticks.y = element_blank(), # Optionally hide y-axis ticks
    panel.spacing = unit(0.001, "mm") # Reduce spacing between grid panels
  )
```

#collapse all hybrid cells into one cell ID for DEG and GSEA 
```{r}
new.cluster.ids <- c('tumor', 'mac', 'hybrid', 'hybrid', 'tumor', 'hybrid', 'mac', 'hybrid', 'tumor', 'hybrid', 'hybrid', 'mixed')
names(new.cluster.ids) <- levels(allcells3)
allcells3_collapsed <- RenameIdents(allcells3, new.cluster.ids)
DimPlot(allcells3_collapsed, reduction = "umap", label = TRUE, pt.size = 0.5) + NoLegend()
```

#find DEGs and GSEA for all hybrid cells 
```{r}

#DEGs
allcells3_collapsed <- JoinLayers(allcells3_collapsed)

#all hybrids compared to mac and tumor and mixed populations
allcells3.hybrid.markers <- FindMarkers(allcells3_collapsed, ident.1 = 'hybrid', min.pct = 0.25)
head(allcells3.hybrid.markers, n = 20)
write.csv(allcells3.hybrid.markers, file = "allhybrid_markers")
View(allcells3.hybrid.markers)

#all hybrids vs tumor
hybrid.v.tum.markers <- FindMarkers(allcells3_collapsed, ident.1 = 'hybrid', ident.2 = "tumor", min.pct = 0.25)
write.csv(hybrid.v.tum.markers, file = "allhybrid_v_tum_markers")

#all hybrids vs tumor
hybrid.v.mac.markers <- FindMarkers(allcells3_collapsed, ident.1 = 'hybrid', ident.2 = "mac", min.pct = 0.25)
write.csv(hybrid.v.mac.markers, file = "allhybrid_v_mac_markers")


#GSEA
allcells3.hybrid.markers$gene <- rownames(allcells3.hybrid.markers)
allcells3.hybrid.markers <- allcells3.hybrid.markers %>% arrange(desc(avg_log2FC))
hybrid_foldchange <- allcells3.hybrid.markers$avg_log2FC
names(hybrid_foldchange) <- allcells3.hybrid.markers$gene

head(hybrid_foldchange)

hybrid_foldchange.df <- as.data.frame(hybrid_foldchange)

hybrid_foldchange.df$ENSEMBL <- mapIds(org.Mm.eg.db, keys = rownames(hybrid_foldchange.df), keytype = "SYMBOL", column = "ENSEMBL")

# omit any NA values 
hybrid_enslist <- na.omit(hybrid_foldchange.df)

hybrid_enslist <- hybrid_foldchange.df$hybrid_foldchange
names(hybrid_enslist) <- hybrid_foldchange.df$ENSEMBL
hybrid_enslist

View(hybrid_enslist)

gse.hybrid.goBP <- gseGO(hybrid_enslist, ont = "BP", keyType = "ENSEMBL", OrgDb = "org.Mm.eg.db", eps = 1e-300)

gse.hybrid.df <- as.data.frame(gse.hybrid.goBP)

write.csv(gse.hybrid.df, file = "gse.hybrid.goBP.csv")

require(DOSE)
dotplothybrid <- dotplot(gse.hybrid.goBP, showCategory=10, split=".sign") + facet_grid(.~.sign)
dotplothybrid

ggsave(filename = "gse.dotplot.hybrid.goBP.tiff", plot = dotplothybrid, width = 7, height = 6, dpi = 600)


```

